### 2024-10-22 latex/paper.tex
latex: added report template

\documentclass[12pt]{article}

\usepackage{sectsty}    % set section font
\usepackage{graphicx}
\usepackage{blindtext}  % generate random text
\usepackage{multicol}   % multiple columns
\usepackage[a4paper,
            bindingoffset=.2in,
            left=.45in,
            right=.65in,
            top=.8in,
            bottom=1in,
            footskip=.5in]  % page number
            {geometry}

% set font family
\usepackage{times}
\usepackage{anyfontsize}

% set roman numbers in sections
\renewcommand{\thesection}{\small\Roman{section}}
% set section and subsection font
\sectionfont{\fontsize{9}{1}\selectfont\sffamily}
\subsectionfont{\fontsize{9}{1}\selectfont\itshape}

% TODO:
% - change the title font
% - change the section font, center, roman numbers

% begin ---------------------------------------
\begin{document}

\fontsize{9}{10}\selectfont  % set font size and spacing

\begin{center}
\section*{\huge{\sffamily The potential of high-performance computing
for the Internet of Sounds}}
Giovanni Santini \\
\textit{Dept.\ of Information Engineering \\
and Computer Science} \\
\textit{University of Trento} \\
Tento, Italy \\
giovanni.santini@unitn.it
\end{center}

\bigskip

\begin{multicols}{2}

\textbf{\textit{Abstract}--
High-Performance Computing (HPC) technology is
impacting several industries, including the creative industries
and those operating in the Internet of Things. However, thus
far, scarce attention has been devoted by researchers to the use
of such technology in Internet of Sounds (IoS) settings. The
IoS is a new area emerging in industry and academia as an
extension of the IoT to the sonic domain, both in musical and
non-musical contexts. The IoS relates to a distributed network
of sound things, which are devices capable of sensing, acquiring,
processing, actuating, and exchanging data serving the purpose
of communicating sound-related information. In this position
paper, we investigate the integration of HPC technology with
the IoS. We illustrate such integration via a set of application
scenarios showing how the IoS can benefit from HPC.\ Finally, we
discuss the open challenges ahead of us to implement a successful
integration of these two kinds of technology. \\
\indent \textit{Index Terms}--Internet of Sounds, High-Performance
Computing, audio processing
}

\fontsize{9}{11}\selectfont  % set font size and spacing

\begingroup\centering\section{INTRODUCTION}\endgroup

% lorem ipsum
\blindtext\blindtext\blindtext

\begingroup\centering\section{RELATED WORK}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{BODY}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{EVALUATION}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{CONCLUSIONS}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{APPENDIX}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{ACKNOWLEDGMENTS}\endgroup

\blindtext\blindtext\blindtext

% end -----------------------------------------
\end{multicols}
\end{document}

### 2024-10-23 benchmarks/pragma_benchmark.cpp
first_assignment: setup

@@ -1,74 +1,0 @@
-/*
- * MIT License
- *
- * Copyright (c) 2024 Giovanni Santini
- *
- * Permission is hereby granted, free of charge, to any person obtaining a copy
- * of this software and associated documentation files (the "Software"), to deal
- * in the Software without restriction, including without limitation the rights
- * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
- * copies of the Software, and to permit persons to whom the Software is
- * furnished to do so, subject to the following conditions:
- *
- * The above copyright notice and this permission notice shall be included in
- all
- * copies or substantial portions of the Software.
- *
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
- * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
- * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
- * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
- * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
- * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
- * SOFTWARE.
- *
- */
-
-#include <pc/pragma.hpp>
-#include <valfuzz/valfuzz.hpp>
-
-BENCHMARK(pragma_mebchmark, "loop")
-{
-    size_t SIZE = 100000;
-    float *a = new float[SIZE];
-    float *b = new float[SIZE];
-
-	auto run_original_loop = [a, b](size_t size) -> void {
-        for (size_t j = 0; j < size; ++j)
-            a[j] = b[j] * 2.0f;
-
-        pc::original_loop(a, b, size);
-        return;
-    };
-
-    RUN_BENCHMARK(100, run_original_loop(100));
-    RUN_BENCHMARK(1000, run_original_loop(1000));
-    RUN_BENCHMARK(10000, run_original_loop(10000));
-    RUN_BENCHMARK(100000, run_original_loop(100000));
-
-    delete[] a;
-    delete[] b;
-}
-
[…]

### 2024-10-23 benchmarks/transpose_benchmark.cpp

@@ -1,0 +1,61 @@
+/*
+ * MIT License
+ *
+ * Copyright (c) 2024 Giovanni Santini
+ *
+ * Permission is hereby granted, free of charge, to any person obtaining a copy
+ * of this software and associated documentation files (the "Software"), to deal
+ * in the Software without restriction, including without limitation the rights
+ * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+ * copies of the Software, and to permit persons to whom the Software is
+ * furnished to do so, subject to the following conditions:
+ *
+ * The above copyright notice and this permission notice shall be included in
+ all
+ * copies or substantial portions of the Software.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+ * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+ * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+ * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+ * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+ * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+ * SOFTWARE.
+ *
+ */
+
+#include <pc/transpose.hpp>
+#include <valfuzz/valfuzz.hpp>
+
+BENCHMARK(transpose_benchmark, "matrix transpose base")
+{
+    std::size_t N = 100;
+    std::size_t M = 100;
+    int **matrix = new int *[N];
+    int **out = new int *[M];
+    for (std::size_t i = 0; i < N; i++)
+    {
+        matrix[i] = new int[N];
+        for (std::size_t j = 0; j < N; j++)
+        {
+            matrix[i][j] = valfuzz::get_random<int>();
+        }
+    }
+    for (std::size_t i = 0; i < M; i++)
+    {
+        out[i] = new int[N];
+    }
+
+    RUN_BENCHMARK(2*2*sizeof(int), pc::matrix_transpose(matrix, out, 2, 2));
[…]

### 2024-10-23 fuzz/transpose_fuzz.cpp

@@ -24,42 +24,5 @@
  *
  */
 
-#include <pc/ilp.hpp>
+#include <pc/transpose.hpp>
 #include <valfuzz/valfuzz.hpp>
-
-/*
-BENCHMARK(array_operations_benchmark, "array_operations")
-{
-    int *a = new (std::nothrow) int[SIZE];
-    int *b = new (std::nothrow) int[SIZE];
-    int *c = new (std::nothrow) int[SIZE];
-
-    if (a == nullptr || b == nullptr || c == nullptr)
-    {
-        std::cerr << "Memory allocation failed!" << std::endl;
-        std::exit(EXIT_FAILURE);
-    }
-
-    // Unoptimized code
-    RUN_BENCHMARK(SIZE, [a, b, c]() -> void {
-        for (int j = 0; j < SIZE; ++j)
-            a[j] = b[j] = c[j] = j;
-
-        pc::array_operations(a, b, c);
-        return;
-    }());
-
-    // Optimized code
-    RUN_BENCHMARK(SIZE, [a, b, c]() -> void {
-        for (int j = 0; j < SIZE; ++j)
-            a[j] = b[j] = c[j] = j;
-
-        pc::array_operations_optimized(a, b, c);
-        return;
-    }());
-
-    delete[] a;
-    delete[] b;
-    delete[] c;
-}
-*/

### 2024-10-23 include/pc/bandwidth.hpp

@@ -1,42 +1,0 @@
-/*
- * MIT License
- *
- * Copyright (c) 2024 Giovanni Santini
-
- * Permission is hereby granted, free of charge, to any person obtaining a copy
- * of this software and associated documentation files (the "Software"), to deal
- * in the Software without restriction, including without limitation the rights
- * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
- * copies of the Software, and to permit persons to whom the Software is
- * furnished to do so, subject to the following conditions:
- *
- * The above copyright notice and this permission notice shall be included in
- all
- * copies or substantial portions of the Software.
- *
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
- * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
- * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
- * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
- * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
- * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
- * SOFTWARE.
- *
- */
-
-#pragma once
-
-#include <cstddef>
-
-namespace pc
-{
-
-void write_and_read(int *a, int size);
-void float_sum(float *vector1, float *vector2, float *result, int N);
-void write_serial(int *a, int size);
-void write_random(int *a, int *indexes, int size);
-void matrix_multiply(float *matrix1, float *matrix2, float *result, int N);
-void bubble_sort(int *a, int *out, int size);
-void merge_sort(int *a, int* out, int size);
-
-} // namespace pc

### 2024-10-23 include/pc/pragma.hpp

@@ -1,37 +1,0 @@
-/*
- * MIT License
- *
- * Copyright (c) 2024 Giovanni Santini
-
- * Permission is hereby granted, free of charge, to any person obtaining a copy
- * of this software and associated documentation files (the "Software"), to deal
- * in the Software without restriction, including without limitation the rights
- * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
- * copies of the Software, and to permit persons to whom the Software is
- * furnished to do so, subject to the following conditions:
- *
- * The above copyright notice and this permission notice shall be included in
- all
- * copies or substantial portions of the Software.
- *
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
- * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
- * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
- * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
- * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
- * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
- * SOFTWARE.
- *
- */
-
-#pragma once
-
-#include <cstddef>
-
-namespace pc
-{
-
-void original_loop(float *a, float *b, size_t size);
-void vectorized_loop(float *a, float *b, size_t size);
-
-} // namespace pc

### 2024-10-23 include/pc/transpose.hpp

@@ -2,7 +2,7 @@
  * MIT License
  *
  * Copyright (c) 2024 Giovanni Santini
- *
+
  * Permission is hereby granted, free of charge, to any person obtaining a copy
  * of this software and associated documentation files (the "Software"), to deal
  * in the Software without restriction, including without limitation the rights
@@ -24,13 +24,13 @@
  *
  */
 
-#include <pc/pc.hpp>
-#include <valfuzz/valfuzz.hpp>
+#pragma once
 
-/*
-// Too slow
-BENCHMARK(nth_fibonacci_benchmark, "Benchmarking fibonacci function")
+#include <cstddef>
+
+namespace pc
 {
-    RUN_BENCHMARK(10000, []()->void{pc::nth_fibonacci(10000);return;});
-}
-*/
+
+void matrix_transpose(int **matrix, int **out, std::size_t N, std::size_t M);
+
+} // namespace pc

### 2024-10-23 src/ilp.cpp

@@ -1,100 +1,0 @@
-/**
- * # Implicit ILP
- *
- * ILP is a measure of how many individual instructions in a program can be
- * processed simultaneously.
- *
- * ## Data dependencies
- *
- * ### True Dependency
- * ```cpp
- * a = b + c;
- * d = a * 2;
- * ```
- *
- * ### Anti dependency
- * ```cpp
- * b = a + 1;
- * a = 5;
- * ```
- *
- * ### Output dependency
- * ```cpp
- * a = 10;
- * a = 20;
- * ```
- *
- * ## Control dependencies
- * ```cpp
- * if (condition)
- *    a = 10;
- * ```
- *
- * # Loop unrolling
- * Original loop:
- * ```cpp
- * for (int i = 0; i < n; ++i)
- *   a[i] = b[i] + c[i];
- * ```
- * Unrolled loop:
- * ```cpp
- * for ( int i = 0; i < n ; i += 4) {
- *   a [ i ] = b [ i ] + c [ i ];
- *   a [ i +1] = b [ i +1] + c [ i +1];
- *   a [ i +2] = b [ i +2] + c [ i +2];
- *   a [ i +3] = b [ i +3] + c [ i +3];
- * }
- * ```
- *
- * # Instruction Reordering
- * Original code:
- * ```cpp
- * a = x + y;
- * b = a + z;
- * c = w * k;
- * ```
- * Reordered code:
- * ```cpp
- * a = x + y;
- * c = w * k;
- * b = a + z;
- * ```
- *
- * # Exercise:
- */
-#include <pc/ilp.hpp>
-void pc::array_operations(int *a, int *b, int *c)
-{
-    for (int i = 1; i < SIZE; i++)
-    {
-        a[i] = b[i] + b[i - 1]; // Instruction 1: True dependency on b[i - 1]
-        b[i] = a[i]
-               * 2; // Instruction 2: Anti dependency (Write After Read on b[i])
-        c[i] = a[i] + b[i - 1]; // Instruction 3: Output dependency (Write After
-                                // Write on a[i])
-    }
-}
-/**
- * Can we unroll the loop?
- */
-
-void pc::array_operations_optimized(int *a, int *b, int *c)
-{
-    for (int i = 1; i < SIZE - 4; i+=4)
-    {
-        a[i] = b[i] + b[i-1];
[…]

### 2024-10-23 src/pc.cpp

@@ -1,40 +1,0 @@
-/*
- * MIT License
- *
- * Copyright (c) 2024 Giovanni Santini
- *
- * Permission is hereby granted, free of charge, to any person obtaining a copy
- * of this software and associated documentation files (the "Software"), to deal
- * in the Software without restriction, including without limitation the rights
- * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
- * copies of the Software, and to permit persons to whom the Software is
- * furnished to do so, subject to the following conditions:
- *
- * The above copyright notice and this permission notice shall be included in
- all
- * copies or substantial portions of the Software.
- *
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
- * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
- * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
- * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
- * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
- * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
- * SOFTWARE.
- *
- */
-
-#include <pc/pc.hpp>
-
-unsigned long long int pc::nth_fibonacci(unsigned int n)
-{
-    using ull = unsigned long long int;
-    ull a = 0, b = 1, c;
-    for (ull i = 0; i < n; i++)
-    {
-        c = a + b;
-        a = b;
-        b = c;
-    }
-    return a;
-}

### 2024-10-23 src/pragma.cpp

@@ -1,23 +1,0 @@
-#include <pc/pragma.hpp>
-
-/**
- * Implicit parallelism
- */
-
-void pc::original_loop(float *a, float *b, size_t size) {
-    for (size_t i = 0; i < size; i++) {
-        a[i] = b[i] * 2.5f + 10.0f;  // Simple linear transformation on each element
-    }
-}
-
-// Optimized loop function using pragmas and loop unrolling
-void pc::vectorized_loop(float *a, float *b, size_t size) {
-	// Use vectorization and loop unrolling to optimize the loop
-    #pragma GCC ivdep  // Ignore vector dependencies, carful with this pragma
-    for (size_t i = 0; i < size; i += 4) {  // Unroll the loop by a factor of 4
-        a[i] = b[i] * 2.5f + 10.0f;
-        a[i + 1] = b[i + 1] * 2.5f + 10.0f;
-        a[i + 2] = b[i + 2] * 2.5f + 10.0f;
-        a[i + 3] = b[i + 3] * 2.5f + 10.0f;
-    }
-}

### 2024-10-23 src/transpose.cpp

@@ -1,0 +1,14 @@
+#include <pc/transpose.hpp>
+
+void pc::matrix_transpose(int **matrix, int **out, std::size_t N, std::size_t M)
+{
+    for (std::size_t i = 0; i < N; i++)
+    {
+        for (std::size_t j = i; j < M; j++)
+        {
+            out[i][j] = matrix[j][i];
+            out[j][i] = matrix[i][j];
+        }
+    }
+    return;
+}

### 2024-10-23 tests/pc_test.cpp

@@ -1,37 +1,0 @@
-/*
- * MIT License
- *
- * Copyright (c) 2024 Giovanni Santini
- *
- * Permission is hereby granted, free of charge, to any person obtaining a copy
- * of this software and associated documentation files (the "Software"), to deal
- * in the Software without restriction, including without limitation the rights
- * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
- * copies of the Software, and to permit persons to whom the Software is
- * furnished to do so, subject to the following conditions:
- *
- * The above copyright notice and this permission notice shall be included in
- all
- * copies or substantial portions of the Software.
- *
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
- * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
- * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
- * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
- * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
- * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
- * SOFTWARE.
- *
- */
-
-#include <pc/pc.hpp>
-#include <valfuzz/valfuzz.hpp>
-
-TEST(nth_fibonacci_test, "Getting fibonacci numbers")
-{
-    ASSERT_EQ(pc::nth_fibonacci(0), 0);
-    ASSERT_EQ(pc::nth_fibonacci(1), 1);
-    ASSERT_EQ(pc::nth_fibonacci(2), 1);
-    ASSERT_EQ(pc::nth_fibonacci(3), 2);
-    ASSERT_EQ(pc::nth_fibonacci(10), 55);
-}

### 2024-10-23 tests/transpose_test.cpp

@@ -2,7 +2,7 @@
  * MIT License
  *
  * Copyright (c) 2024 Giovanni Santini
-
+ *
  * Permission is hereby granted, free of charge, to any person obtaining a copy
  * of this software and associated documentation files (the "Software"), to deal
  * in the Software without restriction, including without limitation the rights
@@ -24,11 +24,34 @@
  *
  */
 
-#pragma once
+#include <pc/transpose.hpp>
+#include <valfuzz/valfuzz.hpp>
 
-namespace pc
+TEST(transpose_matrix, "Matrix Transpose")
 {
+    std::size_t N = 10;
+    std::size_t M = 10;
+    int** matrix = new int*[N];
+    int** out = new int*[M];
+    for (std::size_t i = 0; i < N; i++)
+    {
+        matrix[i] = new int[N];
+        for (std::size_t j = 0; j < N; j++)
+        {
+            matrix[i][j] = valfuzz::get_random<int>();
+        }
+    }
+    for (std::size_t i = 0; i < M; i++)
+    {
+        out[i] = new int[N];
+    }
 
-unsigned long long int nth_fibonacci(unsigned int n);
-
-} // namespace pc
+    pc::matrix_transpose(matrix, out, N, M);
+    for (std::size_t i = 0; i < N; i++)
+    {
+        for (std::size_t j = 0; j < N; j++)
+        {
+            ASSERT(matrix[i][j] == out[j][i]);
+        }
+    }
+}

### 2024-10-23 latex/paper.tex
first_assignment: sequencial matTranspose

@@
-\documentclass[12pt]{article}
-
-\usepackage{sectsty}    % set section font
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{blindtext}  % generate random text
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
 \usepackage{graphicx}
-\usepackage{blindtext}  % generate random text
-\usepackage{multicol}   % multiple columns
-\usepackage[a4paper,
-            bindingoffset=.2in,
-            left=.45in,
-            right=.65in,
-            top=.8in,
-            bottom=1in,
-            footskip=.5in]  % page number
-            {geometry}
-
-% set font family
-\usepackage{times}
-\usepackage{anyfontsize}
-
-% set roman numbers in sections
-\renewcommand{\thesection}{\small\Roman{section}}
-% set section and subsection font
-\sectionfont{\fontsize{9}{1}\selectfont\sffamily}
-\subsectionfont{\fontsize{9}{1}\selectfont\itshape}
-
-% TODO:
-% - change the title font
-% - change the section font, center, roman numbers
-
-% begin ---------------------------------------
+\usepackage{textcomp}
+\usepackage{xcolor}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \begin{document}
 
-\fontsize{9}{10}\selectfont  % set font size and spacing
-
-\begin{center}
-\section*{\huge{\sffamily The potential of high-performance computing
-for the Internet of Sounds}}
-Giovanni Santini \\
-\textit{Dept.\ of Information Engineering \\
-and Computer Science} \\
-\textit{University of Trento} \\
-Tento, Italy \\
-giovanni.santini@unitn.it
-\end{center}
-
-\bigskip
-
-\begin{multicols}{2}
-
-\textbf{\textit{Abstract}--
-High-Performance Computing (HPC) technology is
-impacting several industries, including the creative industries
-and those operating in the Internet of Things. However, thus
-far, scarce attention has been devoted by researchers to the use
-of such technology in Internet of Sounds (IoS) settings. The
-IoS is a new area emerging in industry and academia as an
-extension of the IoT to the sonic domain, both in musical and
-non-musical contexts. The IoS relates to a distributed network
-of sound things, which are devices capable of sensing, acquiring,
-processing, actuating, and exchanging data serving the purpose
-of communicating sound-related information. In this position
-paper, we investigate the integration of HPC technology with
-the IoS. We illustrate such integration via a set of application
-scenarios showing how the IoS can benefit from HPC.\ Finally, we
-discuss the open challenges ahead of us to implement a successful
-integration of these two kinds of technology. \\
-\indent \textit{Index Terms}--Internet of Sounds, High-Performance
-Computing, audio processing
+\title{Fast matrix transpose on Parallel Architectures\\
+%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
+%should not be used}
+%\thanks{Identify applicable funding agency here. If none, delete this.}
 }
 
-\fontsize{9}{11}\selectfont  % set font size and spacing
+\author{\IEEEauthorblockN{1\textsuperscript{st} Giovanni Santini}
+\IEEEauthorblockA{\textit{Dept.\ of Information Engineering and Computer Science} \\
+\textit{University of Trento}\\
+Trento, Italy \\
+giovanni.santini@unitn.it \\
+MAT. 235441}
+}
 
-\begingroup\centering\section{INTRODUCTION}\endgroup
+\maketitle
 
-% lorem ipsum
-\blindtext\blindtext\blindtext
+\begin{abstract}
+This document is a model and instructions for \LaTeX.
+This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
+or Math in Paper Title or Abstract.
+\end{abstract}
 
-\begingroup\centering\section{RELATED WORK}\endgroup
+\begin{IEEEkeywords}
+component, formatting, style, styling, insert
+\end{IEEEkeywords}
 
-\blindtext\blindtext\blindtext
+\section{Introduction}
 
-\begingroup\centering\section{BODY}\endgroup
+Introduction of the problem and importance. \\
+- Describe the background and importance of the problem \\
+- Outline the main objectives of your project \\
 
-\blindtext\blindtext\blindtext
+\blindtext\blindtext
 
-\begingroup\centering\section{EVALUATION}\endgroup
+\section{State of the Art}
 
-\blindtext\blindtext\blindtext
+- Review current research and developments related to your project \\
+- Discuss existing solutions and their limitations \\
+- Identift the gap your project aims to fill \\
 
-\begingroup\centering\section{CONCLUSIONS}\endgroup
+\blindtext\blindtext
 
-\blindtext\blindtext\blindtext
+\section{Contribution and Methodology}
 
-\begingroup\centering\section{APPENDIX}\endgroup
+- Elaborate on the unique contributions of your project \\
+- Describe the methodology proposed, including algorithms (pseudo-code),
+data structures, and parallelization techniques. \\
+- Discuss the challenges faced and how they were addressed \\
 
-\blindtext\blindtext\blindtext
+\blindtext\blindtext
 
-\begingroup\centering\section{ACKNOWLEDGMENTS}\endgroup
+\section{Experiments and System Description}
 
-\blindtext\blindtext\blindtext
+- Detailed description of the computing system and platform. \\
+- Relevant specifications or configurations (e.g., libraries
+and programming toolchains). \\
+- Description of the experimental setup, procedures, and
+methodologies used in the project. \\
+- Discussion on how experiments are designed to test the hypotheses
+or achieve the objectives \\
 
-% end -----------------------------------------
-\end{multicols}
+\blindtext\blindtext
+
+\section{Results and Discussion}
+
+- Presentation of results \\
+- Analysis and interpretation in context \\
+- Comparison with the state-of-the-art \\
+
+\blindtext\blindtext
+
+\section{Conclusions}
+
+- Summary of the key findings and contributions
+
+\section*{References}
+
+Please number citations consecutively within brackets \cite{b1}. The 
+sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
+number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
+the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
+
+Number footnotes separately in superscripts. Place the actual footnote at 
+the bottom of the column in which it was cited. Do not put footnotes in the 
+abstract or reference list. Use letters for table footnotes.
+
+Unless there are six authors or more give all authors' names; do not use 
+``et al.''. Papers that have not been published, even if they have been 
+submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
+that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
+Capitalize only the first word in a paper title, except for proper nouns and 
+element symbols.
+
+For papers published in translation journals, please give the English 
+citation first, followed by the original foreign-language citation \cite{b6}.
+
+\begin{thebibliography}{00}
+\bibitem{b1} G.
+Eason, B.
+Noble, and I.
+N.
+Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
+Trans.
+Roy.
+Soc.
+London, vol.
+A247, pp.
+529--551, April 1955.
+\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
+\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
+\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
+\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
+\bibitem{b6} Y.
+Yorozu, M.
+Hirano, K.
+Oka, and Y.
+Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
+J.
+Magn.
+Japan, vol.
+2, pp.
+740--741, August 1987 [Digests 9th Annual Conf.
+Magnetics Japan, p.
+301, 1982].
+\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+\end{thebibliography}
+\vspace{12pt}
+
 \end{document}

### 2024-10-23 benchmarks/transpose_benchmark.cpp

@@ -30,32 +30,30 @@
 BENCHMARK(transpose_benchmark, "matrix transpose base")
 {
     std::size_t N = 100;
-    std::size_t M = 100;
-    int **matrix = new int *[N];
-    int **out = new int *[M];
+    float **M = new float *[N];
+    float **T = new float *[N];
     for (std::size_t i = 0; i < N; i++)
     {
-        matrix[i] = new int[N];
+        M[i] = new float[N];
         for (std::size_t j = 0; j < N; j++)
         {
-            matrix[i][j] = valfuzz::get_random<int>();
+            M[i][j] = valfuzz::get_random<float>();
         }
     }
-    for (std::size_t i = 0; i < M; i++)
+    for (std::size_t i = 0; i < N; i++)
     {
-        out[i] = new int[N];
+        T[i] = new float[N];
     }
 
-    RUN_BENCHMARK(2*2*sizeof(int), pc::matrix_transpose(matrix, out, 2, 2));
-    RUN_BENCHMARK(4*4*sizeof(int), pc::matrix_transpose(matrix, out, 4, 4));
-    RUN_BENCHMARK(8*8*sizeof(int), pc::matrix_transpose(matrix, out, 8, 8));
-    RUN_BENCHMARK(16*16*sizeof(int), pc::matrix_transpose(matrix, out, 16, 16));
-    RUN_BENCHMARK(32*32*sizeof(int), pc::matrix_transpose(matrix, out, 32, 32));
-    RUN_BENCHMARK(32*32*sizeof(int), pc::matrix_transpose(matrix, out, 32, 32));
-    RUN_BENCHMARK(64*64*sizeof(int), pc::matrix_transpose(matrix, out, 64, 64));
-    RUN_BENCHMARK(80*80*sizeof(int), pc::matrix_transpose(matrix, out, 80, 80));
-    RUN_BENCHMARK(100*100*sizeof(int), pc::matrix_transpose(matrix, out, 100, 100));
+    RUN_BENCHMARK(2 * 2 * sizeof(float), pc::matTranspose(M, T, 2));
+    RUN_BENCHMARK(4 * 4 * sizeof(float), pc::matTranspose(M, T, 4));
+    RUN_BENCHMARK(8 * 8 * sizeof(float), pc::matTranspose(M, T, 8));
+    RUN_BENCHMARK(16 * 16 * sizeof(float), pc::matTranspose(M, T, 16));
[…]

### 2024-10-23 include/pc/transpose.hpp

@@ -31,6 +31,6 @@
 namespace pc
 {
 
-void matrix_transpose(int **matrix, int **out, std::size_t N, std::size_t M);
+void matTranspose(float **M, float **T, std::size_t N);
 
 } // namespace pc

### 2024-10-23 src/transpose.cpp

@@ -1,13 +1,13 @@
 #include <pc/transpose.hpp>
 
-void pc::matrix_transpose(int **matrix, int **out, std::size_t N, std::size_t M)
+void pc::matTranspose(float **M, float **T, std::size_t N)
 {
     for (std::size_t i = 0; i < N; i++)
     {
-        for (std::size_t j = i; j < M; j++)
+        for (std::size_t j = i; j < N; j++)
         {
-            out[i][j] = matrix[j][i];
-            out[j][i] = matrix[i][j];
+            T[i][j] = M[j][i];
+            T[j][i] = M[i][j];
         }
     }
     return;

### 2024-10-23 tests/transpose_test.cpp

@@ -30,28 +30,28 @@
 TEST(transpose_matrix, "Matrix Transpose")
 {
     std::size_t N = 10;
-    std::size_t M = 10;
-    int** matrix = new int*[N];
-    int** out = new int*[M];
+    float **M = new float *[N];
+    float **T = new float *[N];
     for (std::size_t i = 0; i < N; i++)
     {
-        matrix[i] = new int[N];
+        M[i] = new float[N];
         for (std::size_t j = 0; j < N; j++)
         {
-            matrix[i][j] = valfuzz::get_random<int>();
+            M[i][j] = valfuzz::get_random<float>();
         }
     }
-    for (std::size_t i = 0; i < M; i++)
+    for (std::size_t i = 0; i < N; i++)
     {
-        out[i] = new int[N];
+        T[i] = new float[N];
     }
 
-    pc::matrix_transpose(matrix, out, N, M);
+    pc::matTranspose(M, T, N);
+
     for (std::size_t i = 0; i < N; i++)
     {
         for (std::size_t j = 0; j < N; j++)
         {
-            ASSERT(matrix[i][j] == out[j][i]);
+            ASSERT(M[i][j] == T[j][i]);
         }
     }
 }

### 2024-10-23 benchmarks/transpose_benchmark.cpp
first_assignment: sequencial checkSym

@@ -57,3 +57,36 @@ BENCHMARK(transpose_benchmark, "matrix transpose base")
     delete[] M;
     delete[] T;
 }
+
+BENCHMARK(check_sym_benchmark, "check symmetry base")
+{
+    std::size_t N = 100;
+    float **M = new float *[N];
+    for (std::size_t i = 0; i < N; i++)
+    {
+        M[i] = new float[N];
+        for (std::size_t j = 0; j < N; j++)
+        {
+            M[i][j] = valfuzz::get_random<float>();
+        }
+    }
+    // Make the matrix symmetric (worst case scenario)
+    for (std::size_t i = 0; i < N; i++)
+    {
+        for (std::size_t j = i; j < N; j++)
+        {
+            M[i][j] = M[j][i];
+        }
+    }
+
+    RUN_BENCHMARK(2 * 2 * sizeof(float), pc::checkSym(M, 2));
+    RUN_BENCHMARK(4 * 4 * sizeof(float), pc::checkSym(M, 4));
+    RUN_BENCHMARK(8 * 8 * sizeof(float), pc::checkSym(M, 8));
+    RUN_BENCHMARK(16 * 16 * sizeof(float), pc::checkSym(M, 16));
+    RUN_BENCHMARK(32 * 32 * sizeof(float), pc::checkSym(M, 32));
+    RUN_BENCHMARK(64 * 64 * sizeof(float), pc::checkSym(M, 64));
+    RUN_BENCHMARK(80 * 80 * sizeof(float), pc::checkSym(M, 80));
+    RUN_BENCHMARK(100 * 100 * sizeof(float), pc::checkSym(M, 100));
+
+    delete[] M;
+}

### 2024-10-23 include/pc/transpose.hpp

@@ -32,5 +32,6 @@ namespace pc
 {
 
 void matTranspose(float **M, float **T, std::size_t N);
+bool checkSym(float **M, std::size_t N);
 
 } // namespace pc

### 2024-10-23 src/transpose.cpp

@@ -12,3 +12,18 @@ void pc::matTranspose(float **M, float **T, std::size_t N)
     }
     return;
 }
+
+bool pc::checkSym(float **M, std::size_t N)
+{
+    for (std::size_t i = 0; i < N; i++)
+    {
+        for (std::size_t j = 0; j < N; j++)
+        {
+            if (M[i][j] != M[j][i])
+            {
+                return false;
+            }
+        }
+    }
+    return true;
+}

### 2024-10-23 tests/transpose_test.cpp

@@ -27,7 +27,7 @@
 #include <pc/transpose.hpp>
 #include <valfuzz/valfuzz.hpp>
 
-TEST(transpose_matrix, "Matrix Transpose")
+TEST(transpose_matrix_test, "Matrix Transpose")
 {
     std::size_t N = 10;
     float **M = new float *[N];
@@ -54,4 +54,30 @@ TEST(transpose_matrix, "Matrix Transpose")
             ASSERT(M[i][j] == T[j][i]);
         }
     }
+}
+
+TEST(check_sym_test, "Check Symmetry")
+{
+    std::size_t N = 10;
+    float **M = new float *[N];
+    for (std::size_t i = 0; i < N; i++)
+    {
+        M[i] = new float[N];
+        for (std::size_t j = 0; j < N; j++)
+        {
+            M[i][j] = valfuzz::get_random<float>();
+        }
+    }
+
+    ASSERT(pc::checkSym(M, N) == false);
+
+    for (std::size_t i = 0; i < N; i++)
+    {
+        for (std::size_t j = 0; j < N; j++)
+        {
+            M[i][j] = M[j][i];
+        }
+    }
+
+    ASSERT(pc::checkSym(M, N) == true);
 }

### 2024-10-23 benchmarks/transpose_benchmark.cpp
project: now using tenno

@@ -26,21 +26,22 @@
 
 #include <pc/transpose.hpp>
 #include <valfuzz/valfuzz.hpp>
+#include <tenno/ranges.hpp>
 
 BENCHMARK(transpose_benchmark, "matrix transpose base")
 {
-    std::size_t N = 100;
+    tenno::size N = 100;
     float **M = new float *[N];
     float **T = new float *[N];
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
         M[i] = new float[N];
-        for (std::size_t j = 0; j < N; j++)
+        for (auto j : tenno::range(N))
         {
             M[i][j] = valfuzz::get_random<float>();
         }
     }
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
         T[i] = new float[N];
     }
@@ -60,20 +61,20 @@ BENCHMARK(transpose_benchmark, "matrix transpose base")
 
 BENCHMARK(check_sym_benchmark, "check symmetry base")
 {
-    std::size_t N = 100;
+    tenno::size N = 100;
     float **M = new float *[N];
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
         M[i] = new float[N];
-        for (std::size_t j = 0; j < N; j++)
+        for (auto j : tenno::range(N))
         {
             M[i][j] = valfuzz::get_random<float>();
         }
     }
     // Make the matrix symmetric (worst case scenario)
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
-        for (std::size_t j = i; j < N; j++)
+        for (auto j : tenno::range(i, N))
         {
             M[i][j] = M[j][i];
         }

### 2024-10-23 include/pc/transpose.hpp

@@ -26,12 +26,12 @@
 
 #pragma once
 
-#include <cstddef>
+#include <tenno/types.hpp>
 
 namespace pc
 {
 
-void matTranspose(float **M, float **T, std::size_t N);
-bool checkSym(float **M, std::size_t N);
+void matTranspose(float **M, float **T, tenno::size N);
+bool checkSym(float **M, tenno::size N);
 
 } // namespace pc

### 2024-10-23 src/transpose.cpp

@@ -1,10 +1,11 @@
 #include <pc/transpose.hpp>
+#include <tenno/ranges.hpp>
 
-void pc::matTranspose(float **M, float **T, std::size_t N)
+void pc::matTranspose(float **M, float **T, tenno::size N)
 {
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
-        for (std::size_t j = i; j < N; j++)
+        for (auto j : tenno::range(i, N))
         {
             T[i][j] = M[j][i];
             T[j][i] = M[i][j];
@@ -13,11 +14,11 @@ void pc::matTranspose(float **M, float **T, std::size_t N)
     return;
 }
 
-bool pc::checkSym(float **M, std::size_t N)
+bool pc::checkSym(float **M, tenno::size N)
 {
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
-        for (std::size_t j = 0; j < N; j++)
+        for (auto j : tenno::range(i, N))
         {
             if (M[i][j] != M[j][i])
             {

### 2024-10-23 tests/transpose_test.cpp

@@ -26,30 +26,31 @@
 
 #include <pc/transpose.hpp>
 #include <valfuzz/valfuzz.hpp>
+#include <tenno/ranges.hpp>
 
 TEST(transpose_matrix_test, "Matrix Transpose")
 {
-    std::size_t N = 10;
+    tenno::size N = 10;
     float **M = new float *[N];
     float **T = new float *[N];
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
         M[i] = new float[N];
-        for (std::size_t j = 0; j < N; j++)
+        for (auto j : tenno::range(N))
         {
             M[i][j] = valfuzz::get_random<float>();
         }
     }
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
         T[i] = new float[N];
     }
 
     pc::matTranspose(M, T, N);
 
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
-        for (std::size_t j = 0; j < N; j++)
+        for (auto j : tenno::range(N))
         {
             ASSERT(M[i][j] == T[j][i]);
         }
@@ -58,12 +59,12 @@ TEST(transpose_matrix_test, "Matrix Transpose")
 
 TEST(check_sym_test, "Check Symmetry")
 {
-    std::size_t N = 10;
+    tenno::size N = 10;
     float **M = new float *[N];
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
         M[i] = new float[N];
-        for (std::size_t j = 0; j < N; j++)
+        for (auto j : tenno::range(N))
         {
             M[i][j] = valfuzz::get_random<float>();
         }
@@ -71,9 +72,9 @@ TEST(check_sym_test, "Check Symmetry")
 
     ASSERT(pc::checkSym(M, N) == false);
 
-    for (std::size_t i = 0; i < N; i++)
+    for (auto i : tenno::range(N))
     {
-        for (std::size_t j = 0; j < N; j++)
+        for (auto j : tenno::range(N))
         {
             M[i][j] = M[j][i];
         }

### 2024-10-24 include/pc/transpose.hpp
project: added remote tenno dependency using cmake / meson

@@ -31,6 +31,7 @@
 namespace pc
 {
 
+// TODO: __restrict__
 void matTranspose(float **M, float **T, tenno::size N);
 bool checkSym(float **M, tenno::size N);
 

### 2024-10-24 latex/paper_old.tex

\documentclass[12pt]{article}

\usepackage{sectsty}    % set section font
\usepackage{graphicx}
\usepackage{blindtext}  % generate random text
\usepackage{multicol}   % multiple columns
\usepackage[a4paper,
            bindingoffset=.2in,
            left=.45in,
            right=.65in,
            top=.8in,
            bottom=1in,
            footskip=.5in]  % page number
            {geometry}

% set font family
\usepackage{times}
\usepackage{anyfontsize}

% set roman numbers in sections
\renewcommand{\thesection}{\small\Roman{section}}
% set section and subsection font
\sectionfont{\fontsize{9}{1}\selectfont\sffamily}
\subsectionfont{\fontsize{9}{1}\selectfont\itshape}

% begin ---------------------------------------
\begin{document}

\fontsize{9}{10}\selectfont  % set font size and spacing

\begin{center}
% title
\section*{\huge{\sffamily Efficient string matching algorithms for large datasets}}
Giovanni Santini \\
\textit{Dept.\ of Information Engineering \\
and Computer Science} \\
\textit{University of Trento} \\
Tento, Italy \\
giovanni.santini@unitn.it \\
MAT. 325441
\end{center}

\bigskip

\begin{multicols}{2}

\textbf{\textit{Abstract}--
High-Performance Computing (HPC) technology is
impacting several industries, including the creative industries
and those operating in the Internet of Things. However, thus
far, scarce attention has been devoted by researchers to the use
of such technology in Internet of Sounds (IoS) settings. The
IoS is a new area emerging in industry and academia as an
extension of the IoT to the sonic domain, both in musical and
non-musical contexts. The IoS relates to a distributed network
of sound things, which are devices capable of sensing, acquiring,
processing, actuating, and exchanging data serving the purpose
of communicating sound-related information. In this position
paper, we investigate the integration of HPC technology with
the IoS. We illustrate such integration via a set of application
scenarios showing how the IoS can benefit from HPC.\ Finally, we
discuss the open challenges ahead of us to implement a successful
integration of these two kinds of technology. \\
\indent \textit{Index Terms}--Internet of Sounds, High-Performance
Computing, audio processing
}

\fontsize{9}{11}\selectfont  % set font size and spacing

\begingroup\centering\section{INTRODUCTION}\endgroup

% lorem ipsum
\blindtext\blindtext\blindtext

\begingroup\centering\section{RELATED WORK}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{BODY}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{EVALUATION}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{CONCLUSIONS}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{APPENDIX}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{REFERENCES}\endgroup

\blindtext\blindtext\blindtext

% end -----------------------------------------
\end{multicols}
\end{document}

### 2024-10-25 run-benchmarks.sh
project: added script to benchmark multiple targets

@@ -21,35 +21,36 @@
 # LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
 # OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
 # SOFTWARE.
-#
-# ---------------------------------------------------------------------------- #
-# 
-# This script is part of the modern cpp template project. It is used to set the
-# name of the project and update all the files that contain the old name.
-#
-# ---------------------------------------------------------------------------- #
 
-if [ $# -ne 1 ]; then
-    echo "Usage: $0 <project-name>"
-    exit 1
+NUM_ITERATIONS=1000
+BUILD_DIR="build"
+OUTPUT_DIR=benchmarks/plotting/reports
+
+if [ -d $BUILD_DIR ]; then
+    if [ -f "$BUILD_DIR/tests" ]; then
+        echo "Running regular benchmarks"
+        ./build/tests \
+                --benchmark \
+                --num-iterations $NUM_ITERATIONS \
+                --no-multithread \
+                --report $OUTPUT_DIR/base.txt
+    fi
+    if [ -f "$BUILD_DIR/tests_opt" ]; then
+        echo "Running optimized benchmarks"
+        ./build/tests_opt \
+                --benchmark \
+                --num-iterations $NUM_ITERATIONS \
+                --no-multithread \
+                --report $OUTPUT_DIR/opt.txt
+    fi
+    if [ -f "$BUILD_DIR/tests_opt_aggressive" ]; then
+        echo "Running aggressive optimized benchmarks"
+        ./build/tests_opt_aggressive \
+                --benchmark \
+                --num-iterations $NUM_ITERATIONS \
+                --no-multithread \
+                --report $OUTPUT_DIR/opt_aggressive.txt
+    fi
+else
+    echo "Please run the build script first"
 fi
-
-PROJECT_NAME=$1
-echo "Setting project name to $PROJECT_NAME"
-
-# ${PROJECT_NAME,,} Lowercase
-# ${PROJECT_NAME^^} Uppercase
-
-echo "Substituting names..."
-grep -r "pc" -l --exclude-dir=".git/" | tr '\n' ' ' | xargs sed -i "s/pc/${PROJECT_NAME,,}/g"
[…]

### 2024-10-25 run-benchmarks.sh
fix: run benchmark script executes the corrext target based on build dir

@@ -28,24 +28,24 @@ OUTPUT_DIR=benchmarks/plotting/reports
 
 if [ -d $BUILD_DIR ]; then
     if [ -f "$BUILD_DIR/tests" ]; then
-        echo "Running regular benchmarks"
-        ./build/tests \
+        echo "Running regular benchmarks..."
+        ./$BUILD_DIR/tests \
                 --benchmark \
                 --num-iterations $NUM_ITERATIONS \
                 --no-multithread \
                 --report $OUTPUT_DIR/base.txt
     fi
     if [ -f "$BUILD_DIR/tests_opt" ]; then
-        echo "Running optimized benchmarks"
-        ./build/tests_opt \
+        echo "Running optimized benchmarks..."
+        ./$BUILD_DIR/tests_opt \
                 --benchmark \
                 --num-iterations $NUM_ITERATIONS \
                 --no-multithread \
                 --report $OUTPUT_DIR/opt.txt
     fi
     if [ -f "$BUILD_DIR/tests_opt_aggressive" ]; then
-        echo "Running aggressive optimized benchmarks"
-        ./build/tests_opt_aggressive \
+        echo "Running aggressive optimized benchmarks..."
+        ./$BUILD_DIR/tests_opt_aggressive \
                 --benchmark \
                 --num-iterations $NUM_ITERATIONS \
                 --no-multithread \

### 2024-10-25 benchmarks/transpose_benchmark.cpp
benchmark: added correct init and delete functions, corrected size of matrix

@@ -25,14 +25,14 @@
  */
 
 #include <pc/transpose.hpp>
-#include <valfuzz/valfuzz.hpp>
 #include <tenno/ranges.hpp>
+#include <valfuzz/valfuzz.hpp>
 
-BENCHMARK(transpose_benchmark, "matrix transpose base")
+typedef float **matrix;
+
+float **init_matrix(tenno::size N)
 {
-    tenno::size N = 100;
     float **M = new float *[N];
-    float **T = new float *[N];
     for (auto i : tenno::range(N))
     {
         M[i] = new float[N];
@@ -41,36 +41,35 @@ BENCHMARK(transpose_benchmark, "matrix transpose base")
             M[i][j] = valfuzz::get_random<float>();
         }
     }
+    return M;
+}
+
+void delete_matrix(matrix M, tenno::size N)
+{
     for (auto i : tenno::range(N))
     {
-        T[i] = new float[N];
+        delete[] M[i];
     }
-
-    RUN_BENCHMARK(2 * 2 * sizeof(float), pc::matTranspose(M, T, 2));
-    RUN_BENCHMARK(4 * 4 * sizeof(float), pc::matTranspose(M, T, 4));
-    RUN_BENCHMARK(8 * 8 * sizeof(float), pc::matTranspose(M, T, 8));
-    RUN_BENCHMARK(16 * 16 * sizeof(float), pc::matTranspose(M, T, 16));
-    RUN_BENCHMARK(32 * 32 * sizeof(float), pc::matTranspose(M, T, 32));
-    RUN_BENCHMARK(64 * 64 * sizeof(float), pc::matTranspose(M, T, 64));
-    RUN_BENCHMARK(80 * 80 * sizeof(float), pc::matTranspose(M, T, 80));
-    RUN_BENCHMARK(100 * 100 * sizeof(float), pc::matTranspose(M, T, 100));
-
     delete[] M;
-    delete[] T;
+}
+
+BENCHMARK(transpose_benchmark, "matrix transpose base")
+{
+    tenno::size N = 32;
+    matrix M = init_matrix(N);
+    matrix T = init_matrix(N);
+
+    RUN_BENCHMARK(N * N * sizeof(float), pc::matTranspose(M, T, N));
+
+    delete_matrix(M, N);
+    delete_matrix(T, N);
 }
 
 BENCHMARK(check_sym_benchmark, "check symmetry base")
 {
-    tenno::size N = 100;
[…]

### 2024-10-25 tests/transpose_test.cpp

@@ -25,8 +25,8 @@
  */
 
 #include <pc/transpose.hpp>
-#include <valfuzz/valfuzz.hpp>
 #include <tenno/ranges.hpp>
+#include <valfuzz/valfuzz.hpp>
 
 TEST(transpose_matrix_test, "Matrix Transpose")
 {

### 2024-11-22 benchmarks/benchmarks.cpp
matTranspose: added benchmarks

@@ -110,11 +110,13 @@ AFTER()
   matrix_free(matrix_out, PC_MATRIX_MAX_SIZE);
 }
 
+
 /*============================================*\
 |                   TRANSPOSE                  |
 \*============================================*/
 
-BENCHMARK(transpose_benchmark, "matrix transpose base")
+BENCHMARK(transpose_benchmark,
+	  "matTranspose base")
 {
   for (size_t N = 2; N < 12; ++N)
     {
@@ -123,7 +125,8 @@ BENCHMARK(transpose_benchmark, "matrix transpose base")
     }
 }
 
-BENCHMARK(transpose_half_benchmark, "matrix transpose half")
+BENCHMARK(transpose_half_benchmark,
+	  "matTranspose half")
 {
     for (size_t N = 2; N < 12; ++N)
     {
@@ -131,6 +134,176 @@ BENCHMARK(transpose_half_benchmark, "matrix transpose half")
 		    pc::matTransposeHalf(matrix_in, matrix_out, (1<<N)));
     }
 }
+
+BENCHMARK(transpose_columns_benchmark,
+	  "matTranspose columns")
+{
+    for (size_t N = 2; N < 12; ++N)
+    {
+      RUN_BENCHMARK((1<<N) * (1<<N) * sizeof(float),
+		    pc::matTransposeColumns(matrix_in, matrix_out, (1<<N)));
+    }
+}
+
+BENCHMARK(transpose_vectorization_benchmark,
+	  "matTranspose vectorization")
+{
+    for (size_t N = 2; N < 12; ++N)
+    {
+      RUN_BENCHMARK((1<<N) * (1<<N) * sizeof(float),
+		    pc::matTransposeVectorization(matrix_in, matrix_out, (1<<N)));
+    }
+}
+
+BENCHMARK(transpose_unrolling_outer_benchmark,
+	  "matTranspose unrolling outer")
+{
+    for (size_t N = 2; N < 12; ++N)
+    {
+      RUN_BENCHMARK((1<<N) * (1<<N) * sizeof(float),
+		    pc::matTransposeOmp2(matrix_in, matrix_out, (1<<N)));
+    }
+}
+
+BENCHMARK(transpose_omp_4_benchmark,
+	  "matTranspose omp 4")
+{
[…]

### 2024-11-22 src/check_symm.cpp

@@ -30,6 +30,16 @@ bool pc::checkSym(float **M, tenno::size N)
   return symm;
 }
 
+bool pc::checkSymColumns(float **M, tenno::size N)
+{
+  bool symm = true;
+  for (size_t i = 0; i < N; ++i)
+    for (size_t j = i; j < N; ++j)
+      if (M[j][i] != M[i][j])
+	symm = false;
+  return symm;
+}
+
 /*============================================*\
 |              IMPLICIT PARALLELISM            |
 \*============================================*/

### 2024-11-23 src/transpose.cpp
docs: written some notes

@@ -102,6 +102,36 @@ void pc::matTransposeColumns(float **M, float **T, tenno::size N)
 }
 
 
+/* linux kernel 6.12, lib/bch.c 562 */
+/* assuming a 16x16 matrix */
+void pc::matTransposeLinux(float **M, float **T, tenno::size N)
+{
+  for (j = 8; j != 0; j >>= 1, mask ^= (mask << j)) {
+    for (k = 0; k < 16; k = (k+j+1) & ~j) {
+      t = ((M[k] >> j)^M[k+j]) & mask;
+      M[k] ^= (t << j);
+      M[k+j] ^= t;
+    }
+  }
+}
+
+/* Linux strikes again arch/x86/crypto/aria-aesni-avc2-asm_64.S */
+/*
+  
+#define transpose_4x4(x0, x1, x2, x3, t1, t2)		\
+	vpunpckhdq x1, x0, t2;				\
+	vpunpckldq x1, x0, x0;				\
+							\
+	vpunpckldq x3, x2, t1;				\
+	vpunpckhdq x3, x2, x2;				\
+							\
+	vpunpckhqdq t1, x0, x1;				\
+	vpunpcklqdq t1, x0, x0;				\
+							\
+	vpunpckhqdq x2, t2, x3;				\
+	vpunpcklqdq x2, t2, x2;
+*/
+
 /*============================================*\
 |              IMPLICIT PARALLELISM            |
 \*============================================*/

### 2024-11-23 benchmarks/benchmarks.cpp
plotting: checkSymm graphs

@@ -30,7 +30,7 @@
 #include <tenno/random.hpp>
 #include <valfuzz/valfuzz.hpp>
 
-#define PC_MATRIX_MAX_SIZE 1<<12
+#define PC_MATRIX_MAX_SIZE (1<<12) + 4 // 1 is required to prevent segfault
 #define PC_RANDOM_MATRIX_SIZE 256
 
 /*============================================*\
@@ -118,9 +118,9 @@ AFTER()
 BENCHMARK(transpose_benchmark,
 	  "matTranspose base")
 {
-  for (size_t N = 2; N < 12; ++N)
+  for (size_t N = 2; N <= 12; ++N)
     {
-      RUN_BENCHMARK((1<<N) * (1<<N) * sizeof(float),
+      RUN_BENCHMARK((1<<N),
 		    pc::matTranspose(matrix_in, matrix_out, (1<<N)));
     }
 }
@@ -128,9 +128,9 @@ BENCHMARK(transpose_benchmark,
 BENCHMARK(transpose_half_benchmark,
 	  "matTranspose half")
 {
-    for (size_t N = 2; N < 12; ++N)
+    for (size_t N = 2; N <= 12; ++N)
     {
-      RUN_BENCHMARK((1<<N) * (1<<N) * sizeof(float),
+      RUN_BENCHMARK((1<<N),
 		    pc::matTransposeHalf(matrix_in, matrix_out, (1<<N)));
     }
 }
@@ -138,9 +138,9 @@ BENCHMARK(transpose_half_benchmark,
 BENCHMARK(transpose_columns_benchmark,
 	  "matTranspose columns")
 {
-    for (size_t N = 2; N < 12; ++N)
+    for (size_t N = 2; N <= 12; ++N)
     {
-      RUN_BENCHMARK((1<<N) * (1<<N) * sizeof(float),
+      RUN_BENCHMARK((1<<N),
 		    pc::matTransposeColumns(matrix_in, matrix_out, (1<<N)));
     }
 }
@@ -148,9 +148,9 @@ BENCHMARK(transpose_columns_benchmark,
 BENCHMARK(transpose_vectorization_benchmark,
 	  "matTranspose vectorization")
 {
-    for (size_t N = 2; N < 12; ++N)
+    for (size_t N = 2; N <= 12; ++N)
     {
[…]

### 2024-11-23 benchmarks/plotting/data/cluster_data.py

@@ -13,3 +13,7 @@ class cluster_data:
 
     peak_theoretical_bandwidth = \
         memory_clock_rate * bus_width * transfer_rate / 10e9  # GB/s
+
+    core_count = 4
+    clock_speed = 3.2 # GHz
+    flops_per_cycle = 8  # 8 for AVX

### 2024-11-23 src/check_symm.cpp

@@ -26,7 +26,7 @@ bool pc::checkSym(float **M, tenno::size N)
   for (size_t i = 0; i < N; ++i)
     for (size_t j = i; j < N; ++j)
       if (M[i][j] != M[j][i])
-	symm = false;
+	    symm = false;
   return symm;
 }
 
@@ -36,7 +36,7 @@ bool pc::checkSymColumns(float **M, tenno::size N)
   for (size_t i = 0; i < N; ++i)
     for (size_t j = i; j < N; ++j)
       if (M[j][i] != M[i][j])
-	symm = false;
+	    symm = false;
   return symm;
 }
 

### 2024-11-23 src/transpose.cpp

@@ -102,19 +102,6 @@ void pc::matTransposeColumns(float **M, float **T, tenno::size N)
 }
 
 
-/* linux kernel 6.12, lib/bch.c 562 */
-/* assuming a 16x16 matrix */
-void pc::matTransposeLinux(float **M, float **T, tenno::size N)
-{
-  for (j = 8; j != 0; j >>= 1, mask ^= (mask << j)) {
-    for (k = 0; k < 16; k = (k+j+1) & ~j) {
-      t = ((M[k] >> j)^M[k+j]) & mask;
-      M[k] ^= (t << j);
-      M[k+j] ^= t;
-    }
-  }
-}
-
 /* Linux strikes again arch/x86/crypto/aria-aesni-avc2-asm_64.S */
 /*
   

### 2024-11-24 latex/paper.tex
report: introduction

@@
 
-\title{Fast matrix transpose on Parallel Architectures\\
+\title{Matrix Transposition and Symmetry: a case study on Superscalar Architectures\\
 %{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Giovanni Santini}
+\author{\IEEEauthorblockN{Giovanni Santini}
 \IEEEauthorblockA{\textit{Dept.\ of Information Engineering and Computer Science} \\
@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
+TODO *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
 or Math in Paper Title or Abstract.
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+parallel computing, matrix transpose, symmetrix matrix. benchmarking
 \end{IEEEkeywords}
@@
 
-Introduction of the problem and importance. \\
-- Describe the background and importance of the problem \\
-- Outline the main objectives of your project \\
+Contemporary CPU architectures leverage superscalar designs to enhance performance, incorporating multiple cores and advanced Single Instruction Multiple Data (SIMD) extensions, such as Intel's Streaming SIMD Extensions (SSE) and Advanced Vector Extensions (AVX).
+Although SSE is not a new technology, being first introduced in the Pentium III in 1999 and subsequently expanded in later years \cite{b1}, modern algorithms leveraging SIMD instructions continue to be a topic of active research in recent literature.
+For example, the FreeBSD's SIMD-enhanced libc is actively reworking the standard library's code. \cite{b2}.
+In 2023, most of the string functions were reimplemented by Robert Clausecker et al on many architectures \cite{b3}, exploiting vector instructions and improving the overall performance of the standard library \cite{b4}.
+Similarly, efficient multi-threaded algorithms are becoming more critical as the demand of computation increases worldwide due to AI training \cite{b5} \cite{b6}.
+This case study examines the implementations and performance of two algorithms—out-of-place square matrix transposition (matTranspose) and square matrix symmetry checking (checkSymm)—that leverage superscalar architectures.
+Both algorithms are used pervasively in performance critical applications.
+For example, matrix transposition was used in NASA's Apollo 11 Guidance Computer (AGC) to convert vectors in platenary coordinate system to base reference system \cite{b7}.
+Additionally, matrix transposition is used by Fourier Transform in the West (FFTW) \cite{b8}, a popular library to calculate Fast Furier Transofrms, and in signal processing \cite{b9}.
+Furthermore, the current C++26 draft proposes basic linear algebra algorithms in the "linalg" header, including matrix transposition \cite{b10}.
 
-\blindtext\blindtext
+In section 2, we discuss the state of the art of the two algorithms.
+In section 3, we elaborate on the algorithms analyzed, discuss performance evaluation and we present a new benchmarking framework.
+In section 4, we analyze experimental data and graphs.
 
@@
 
-\blindtext\blindtext
+\textit{We define the transpose of a matrix $A$, and we denote it by $A^t$, as the matrix obtained from $A$ by interchanging rows and columns.
+Specifically, if $A$ is an $m$-by-$n$ matrix, then $A^t$ is the $n$-by-$m$ matrix whose entries are given by the equation (1)}.
+\cite{b11}
+
+\begin{equation}
+	(A^t)_{k,j} = A_{j,k}
+\end{equation}
+
+Computing the transpose of a matrix is a well discussed problem.
+
+Older fortran implementations were difficult due to fortran.
+
+# a lot of In place and rectangular
+
+Analysis of In-Situ Transposition
+"Algorithms for in-place matrix transposition make use of the cyclic structure of
+the transposition mapping. A clear introduction to the process was given in
+1958 by Windley [6]. In 1968 Boothroyd presented ACM Algorithm 302 [1]. Then
+in 1970 Laflin and Brefner presented ACM Algorithm 380 which they showed to
+be significantly faster than Algorithm 302 [4]"
+
+	[1] Transpose vector stored array.
+	BOOTHROYD,Z. Algorithm 302: Transpose vector stored array. Go~r~m.A~M I0, ~ (1967),
+	292-293
+
+	anche 380 important, revised by 513
+
+Naive algorithm does not exploit locality of reference. Recursive solutions were proposed.
+
+## out-of-core - where I/O needs to be minimized
+
+ The oldest context in which the spatial locality of transposition seems to have been studied is for out-of-core operation (by Alltop, 1975), where the matrix is too large to fit into main memory ("core").
+ Reviews of out-of-core algorithms, especially as applied to parallel computing, can be found in e.g. Suh & Prasanna (2002) and Krishnamoorth et al. (2004).
+
+ An Efficient Algorithm for
+ Out-of-Core Matrix Transposition
+ Jinwoo Suh, Member, IEEE, and Viktor K. Prasanna, Fellow, IEEE
+
+ 	"These efforts have focused on reducing the number
+ 	of I/O operations. However, in state-of-the-art architectures, memory-memory data transfer time and index computation time are also
+ 	significant components of the overall time. In this paper, we propose an algorithm that considers the index computation time and the
+ 	I/O time and reduces the overall execution time."
+
+FFTW - MPI shit
+/* Recursive "radix-r" distributed transpose, which breaks a transpose
+over p processes into p/r transposes over r processes plus r
+transposes over p/r processes.  If performed recursively, this
+produces a total of O(p log p) messages vs. O(p^2) messages for a
+direct approach.
+*/
+
+## GPUs
+
+In-Place Transposition of Rectangular Matrices on Accelerators
+
+An Efficient Matrix Transpose in CUDA C/C++
+Feb 18, 2013
+By Mark Harris
+
+\textit{A square matrix $A$ is called symmetric if it equals its transpose} \cite{b100}.
 
@@
 \begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+\bibitem{b1} Intel® 64 and IA-32 Architectures Software Developer’s Manual Combined Volumes: 1, 2A, 2B, 2C, 2D, 3A, 3B, 3C, 3D, and 4. Volume 1, Section 2.2.7 "SIMD Instructions"
+\bibitem{b2} SIMD man page "https://man.freebsd.org/cgi/man.cgi?simd(7)"
+\bibitem{b3} Git blame on freeBSD's source code, in "lib/libc/amd64/string", shows recent activity
+\bibitem{b4} https://freebsdfoundation.org/blog/a-sneak-peek-simd-enhanced-string-functions-for-amd64/
+\bibitem{b5} Amodei, D. Hernandez, D. AI and Compute, https://openai.com/blog/ai-and-compute
+\bibitem{b6} Andrew J. Lohn, Micah Musse. AI and Compute How Much Longer Can Computing Power Drive Artificial Intelligence Progress? CSET Issue Brief
+\bibitem{b7} Digitalized Apollo 11 guidance computer by Virtual AGC and MIT Museum.
+Code in "Apollo-11/Comanche055/PLANETERY\_INTERNAL\_ORIENTATION.agc" line 34; "Apollo-11/Luminary099/ATTITUDE\_MANEUVER\_ROUTINE.agc" line 441
+\bibitem{b8} Matteo Frigo and Steven G. Johnson. "The design and implementation of FFTW3". The word "transpose" is mentioned 21 times
+\bibitem{b9} M.
+Henriksson and O.
+Gustafsson, "Streaming Matrix Transposition on FPGAs Using Distributed Memories," 2023 IEEE Nordic Circuits and Systems Conference (NorCAS), Aalborg, Denmark, 2023, pp.
+1-6, doi: 10.1109/NorCAS58970.2023.10305472.
+\bibitem{b10} "https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2023/p1673r12.html" Accessed in november 2024
+\bibitem{b11} Sheldon Axler. Linear Algebra Done Right ourth edition
+16 November 2024. Section 3C Matrices, definition 3.54
+\bibitem{b12} Sheldon Axler. Linear Algebra Done Right ourth edition
+16 November 2024. Section 9A Bilinear Forms and Quadratic Forms, definition 9.11
+
 \end{thebibliography}

### 2024-11-25 benchmarks/benchmarks.cpp
transpose: added cyclic benchmark

@@ -30,7 +30,7 @@
 #include <tenno/random.hpp>
 #include <valfuzz/valfuzz.hpp>
 
-#define PC_MATRIX_MAX_SIZE (1<<12) + 4 // 1 is required to prevent segfault
+#define PC_MATRIX_MAX_SIZE 1<<12
 #define PC_RANDOM_MATRIX_SIZE 256
 
 /*============================================*\
@@ -66,7 +66,7 @@ constexpr tenno::array<float, PC_RANDOM_MATRIX_SIZE> random_arr2()
 matrix matrix_alloc(tenno::size N)
 {
   matrix M = new float*[N];
-  for (const auto i : tenno::range(N))
+  for (unsigned int i = 0; i < N; ++i)
     {
       M[i] = new float[N];
     }
@@ -91,7 +91,7 @@ void matrix_init(matrix M, tenno::size N)
 
 void matrix_free(matrix M, tenno::size N)
 {
-  for (const auto i : tenno::range(N))
+  for (unsigned int i = 0; i < N; ++i)
     delete[] M[i];
   delete[]  M;
 }
@@ -145,6 +145,28 @@ BENCHMARK(transpose_columns_benchmark,
     }
 }
 
+BENCHMARK(transpose_cyclic_benchmark,
+	  "matTranspose cyclic")
+{
+    /* Initialize the vector */
+    float* arr_in = new float[PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE];
+    float* arr_out = new float[PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE];
+
+    constexpr auto arr1 = random_arr1();
+
+    for (size_t i = 0; i < PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE; ++i)
+      arr_in[i] = arr1[i % PC_RANDOM_MATRIX_SIZE];
+      
+    for (size_t N = 2; N <= 12; ++N)
+    {
+      RUN_BENCHMARK((1<<N),
+		    pc::matTransposeCyclic(arr_in, arr_out, (1<<N)));
+    }
+
+    delete[] arr_in;
+    delete[] arr_out;
+}
+
 BENCHMARK(transpose_vectorization_benchmark,
 	  "matTranspose vectorization")
 {
[…]

### 2024-11-25 include/pc/transpose.hpp

@@ -39,6 +39,7 @@ namespace pc
 void matTranspose(float **M, float **T, tenno::size N);
 void matTransposeHalf(float **M, float **T, tenno::size N);
 void matTransposeColumns(float **M, float **T, tenno::size N);
+void matTransposeCyclic(float *M, float *T, tenno::size N);
 
 
 /*============================================*\
@@ -59,6 +60,7 @@ void matTransposeUnrolledOuter(float **M, float **T, tenno::size N);
 void matTransposeHalfVectorization(float **M, float **T, tenno::size N);
 void matTransposeHalfUnrolledInner(float **M, float **T, tenno::size N);
 void matTransposeHalfUnrolledOuter(float **M, float **T, tenno::size N);
+void matTransposeCyclicUnrolled(float *M, float *T, tenno::size N);
 
 
 /*============================================*\

### 2024-11-25 src/check_symm.cpp

@@ -58,7 +58,7 @@ bool pc::checkSymVectorization(float **M, tenno::size N)
 bool pc::checkSymUnrollingOuter(float **M, tenno::size N)
 {
   bool symm = true;
-  for (size_t i = 0; i < N; i = i + 4)
+  for (size_t i = 0; i < N - 4; i += 4)
     for (size_t j = i; j < N; ++j)
       {
 	if (M[i][j] != M[j][i]) symm = false;
@@ -73,7 +73,7 @@ bool pc::checkSymUnrollingInner(float **M, tenno::size N)
 {
   bool symm = true;
   for (size_t i = 0; i < N; ++i)
-    for (size_t j = i; j < N; j = j + 4)
+    for (size_t j = i; j < N - 4; j += 4)
       {
 	if (M[i][j] != M[j][i]) symm = false;
 	if (M[i][j+1] != M[j+1][i]) symm = false;
@@ -86,7 +86,7 @@ bool pc::checkSymUnrollingInner(float **M, tenno::size N)
 bool pc::checkSymUnrollingOuterNoBranch(float **M, tenno::size N)
 {
   bool symm = true;
-  for (size_t i = 0; i < N; i = i + 4)
+  for (size_t i = 0; i < N - 4; i += 4)
     for (size_t j = i; j < N; ++j)
       {
 	symm &=
@@ -99,11 +99,12 @@ bool pc::checkSymUnrollingOuterNoBranch(float **M, tenno::size N)
       }
   return symm;
 }
+
 bool pc::checkSymUnrollingInnerNoBranch(float **M, tenno::size N)
 {
   bool symm = true;
   for (size_t i = 0; i < N; ++i)
-    for (size_t j = i; j < N; j = j + 4)
+    for (size_t j = i; j < N - 4; j += 4)
       {
 	symm &=
 	        ((

### 2024-11-25 src/transpose.cpp

@@ -13,15 +13,6 @@
  */ 
 
 /*
-void transpose(float *src, float *dst, const int N, const int M) {
-    #pragma omp parallel for
-    for(int n = 0; n<N*M; n++) {
-        int i = n/N;
-        int j = n%N;
-        dst[n] = src[M*j + i];
-    }
-}
-
 // With blocks
 inline void transpose_block_SSE4x4(float *A, float *B, const int n, const int m, const int lda, const int ldb ,const int block_size) {
     #pragma omp parallel for
@@ -47,16 +38,6 @@ inline void transpose_block_SSE4x4(float *A, float *B, const int n, const int m,
 |                   BASELINE                   |
 \*============================================*/
 
-/*
- * matTranspose
- *
- * For each element of the matrix with column i and row j,
- * the algorithm copies the value in the output matrix
- * in column j and row i.
- * Time Complexity: O(n^2)
- * Space Complexity: O(n^2)
- * Assumption: square matrix
- */
 void pc::matTranspose(float **M, float **T, tenno::size N)
 {
   for (tenno::size i = 0; i < N; ++i)
@@ -65,20 +46,6 @@ void pc::matTranspose(float **M, float **T, tenno::size N)
   return;
 }
 
-/*
- * matTransposeHalf
- *
- * An alternative algorithm to matTranspose is to iterate
- * over only the upper half of the matrix with respect
- * to the diagonal. This cuts in half the number of iterations
- * but the number of memory accesses (both read and writes)
- * do not change, as well as the time and space complexity.
- * Despite this, this algorithm improves from the previous
- * one by about 33%
- * Time Complexity: O(n^2)
- * Space Complexity: O(n^2)
- * Assumption: square matrix
- */
 void pc::matTransposeHalf(float **M, float **T, tenno::size N)
 {
   for (tenno::size i = 0; i < N; ++i)
@@ -101,6 +68,13 @@ void pc::matTransposeColumns(float **M, float **T, tenno::size N)
 
 }
 
+void pc::matTransposeCyclic(float *M, float *T, tenno::size N) {
+    for(long unsigned int n = 0; n<N*N; n++) {
+        long unsigned int i = n/N;
+        long unsigned int j = n%N;
[…]

### 2024-11-25 latex/paper.tex
latex: state of the art draft

@@
 
-Computing the transpose of a matrix is a well discussed problem.
+Computing the transpose of a matrix is a well discussed problem. Earlier
+work focus on in-situ transporision, that is transposing a matrix without
+relying on additional space. The implementations often exploit the acyclic
+structure of the transposition mapping since values of a matrix of size $N$ can be stored
+in a single contiguous array or size $N$*$N$ and indexed with $row*#rows+ column$.
+Remarcable contributions include Boothroyd's ACM Algorithm 302
+presented in 1966\cite{b?}, the 1970 revision by Laflin and Brefner
+in ACS Algorithm 380 and a later revision with ACM Algorithm 513 /* TODO SOURCES and names */.
 
-Older fortran implementations were difficult due to fortran.
+SIMD implementations have also been studied
+
+Another class of algorithms include out-of-core matrix transposition, that
+is computing the transpose of a matrix that is too big to fit in memory.
+This problem focuses on minimizing the number of I/O operations since memory-memory data transfer
+is notably slower than math operations. Work in out-of-core transposition include /*TODO*/
+
+GPU /* TODO */
 

### 2024-11-25 src/transpose.cpp

@@ -27,6 +27,8 @@ inline void transpose_block_SSE4x4(float *A, float *B, const int n, const int m,
             }
         }
     }
+
+    // we may directly store the squares sequencially
 }
 */
 

### 2024-11-25 latex/paper.tex
transpose: asm implementation

--- 2024-11-24 latex/paper.tex
@@
 
-Contemporary CPU architectures leverage superscalar designs to enhance performance, incorporating multiple cores and advanced Single Instruction Multiple Data (SIMD) extensions, such as Intel's Streaming SIMD Extensions (SSE) and Advanced Vector Extensions (AVX).
+Modern CPUs use superscalar architectures to improve performance, combining multiple cores with advanced Single Instruction Multiple Data (SIMD) technologies.
+Examples include Intel's Streaming SIMD Extensions (SSE) and Advanced Vector Extensions (AVX).
 Although SSE is not a new technology, being first introduced in the Pentium III in 1999 and subsequently expanded in later years \cite{b1}, modern algorithms leveraging SIMD instructions continue to be a topic of active research in recent literature.
@@
 Similarly, efficient multi-threaded algorithms are becoming more critical as the demand of computation increases worldwide due to AI training \cite{b5} \cite{b6}.
-This case study examines the implementations and performance of two algorithms—out-of-place square matrix transposition (matTranspose) and square matrix symmetry checking (checkSymm)—that leverage superscalar architectures.
-Both algorithms are used pervasively in performance critical applications.
+This case study examines the implementations and performance of two algorithms—out-of-place square matrix transposition (matTranspose) and square matrix symmetry checking (checkSymm)—in the context of a superscalar architecture.
+Both algorithms have been used pervasively in performance critical applications.
 For example, matrix transposition was used in NASA's Apollo 11 Guidance Computer (AGC) to convert vectors in platenary coordinate system to base reference system \cite{b7}.
@@
 In section 3, we elaborate on the algorithms analyzed, discuss performance evaluation and we present a new benchmarking framework.
-In section 4, we analyze experimental data and graphs.
+In section 4, we show experimental data and graphs that is analyzed
+in section 5. The conclusion is in section 6.
 
 \section{State of the Art}
-
-- Review current research and developments related to your project \\
-- Discuss existing solutions and their limitations \\
-- Identift the gap your project aims to fill \\
 
@@
 
-Computing the transpose of a matrix is a well discussed problem.
+Computing the transpose of a matrix is a well discussed computational problem. Early
+work focuses on in-situ transporision, that is transposing a matrix without
+relying on additional space. Implementations were often based on the cyclic
+structure of the transposition mapping: values of a matrix of size $N$ can be stored
+in a single contiguous array or size $N$*$N$ and indexed with $row*\#rows+ column$.
+Remarcable contributions include Boothroyd's ACM Algorithm 302 (1966) using cyclic mapping \cite{b13}, Laflin and Brefner ACM Algorithm 380
+adding fixed points arithmetic \cite{b14} and
+a later revision of the two by Esko G. Cate and David W. Twigg in ACM Algorithm 513 \cite{b15} .
+SIMD implementations for specific architectures are also widely adopted
+in performance critical applications like cryptographic libraries using instructions like Intel's vpunpckldq and vpunpckhdq \cite{b16} reporting improvements up to 70\% \cite{b17}.
+However, matrix operations are not efficient as vector registers access data only linearly, motivating research on new vector register file (VFS) designs such as diagonal registers \cite{b23}.
+Another area of research includes out-of-core matrix transposition, that
+is computing the transpose of a matrix that exceeds the size of the available in-core memory.
+This problem focuses on optimizing the number of I/O operations and file layout since memory-memory data transfer is notably slower than math operations.
+Work in out-of-core transposition include W.
+0.
+Alltop (1975) \cite{b18}, Suh \& Prasanna (2002) \cite{b19} and Krishnamoorth et al.
+(2004) \cite{b20}.
+Recent work explores parallelization and the use of GPUs to offload algebra operations. Examples are I-Jui Sung et al \cite{b21} and Mark Harris \cite{b22}.
 
-Older fortran implementations were difficult due to fortran.
+\textit{A square matrix $A$ is called symmetric if it equals its transpose} \cite{b12}.
 
-# a lot of In place and rectangular
-
-Analysis of In-Situ Transposition
-"Algorithms for in-place matrix transposition make use of the cyclic structure of
-the transposition mapping. A clear introduction to the process was given in
-1958 by Windley [6]. In 1968 Boothroyd presented ACM Algorithm 302 [1]. Then
-in 1970 Laflin and Brefner presented ACM Algorithm 380 which they showed to
-be significantly faster than Algorithm 302 [4]"
-
-	[1] Transpose vector stored array.
-	BOOTHROYD,Z. Algorithm 302: Transpose vector stored array. Go~r~m.A~M I0, ~ (1967),
-	292-293
-
-	anche 380 important, revised by 513
-
-Naive algorithm does not exploit locality of reference. Recursive solutions were proposed.
-
-## out-of-core - where I/O needs to be minimized
-
- The oldest context in which the spatial locality of transposition seems to have been studied is for out-of-core operation (by Alltop, 1975), where the matrix is too large to fit into main memory ("core").
- Reviews of out-of-core algorithms, especially as applied to parallel computing, can be found in e.g. Suh & Prasanna (2002) and Krishnamoorth et al. (2004).
-
- An Efficient Algorithm for
- Out-of-Core Matrix Transposition
- Jinwoo Suh, Member, IEEE, and Viktor K. Prasanna, Fellow, IEEE
-
- 	"These efforts have focused on reducing the number
- 	of I/O operations. However, in state-of-the-art architectures, memory-memory data transfer time and index computation time are also
- 	significant components of the overall time. In this paper, we propose an algorithm that considers the index computation time and the
- 	I/O time and reduces the overall execution time."
-
-FFTW - MPI shit
-/* Recursive "radix-r" distributed transpose, which breaks a transpose
-over p processes into p/r transposes over r processes plus r
-transposes over p/r processes.  If performed recursively, this
-produces a total of O(p log p) messages vs. O(p^2) messages for a
-direct approach.
-*/
-
-## GPUs
-
-In-Place Transposition of Rectangular Matrices on Accelerators
-
-An Efficient Matrix Transpose in CUDA C/C++
-Feb 18, 2013
-By Mark Harris
-
-\textit{A square matrix $A$ is called symmetric if it equals its transpose} \cite{b100}.
+While symmetric matrces are actively used in various areas of research as their
+properties allow for several optimizations in various problems, an in depth case study of symmetry
+checking on modern architectures is yet to be published.
 
@@
 
-\blindtext\blindtext
+\iffalse
+Remember: write
+- what you did (contribution): high level overview like "hey, i want to test different vectorization ond parallelization implementations"
+- how you did it (methodology): we benchmarked algorithm A ... (all the pseudocodes). Whay we mesured
+
+\fi
+
+Quiesti qui bellissimi e stra fighi
+- uno (pseudocodice)
+- due (pseudocodice)
+- tre (pseudocodice)
+- ...
+Ho usato implicit ed esplicit parallelism con vectorization e openmp
+Voglio fare questo per comparare quest'altro cos' pel meme
+Benchmarkare è difficle perchè... quindi ho fatto ciò... (aplia qua)
 
@@
 
-\blindtext\blindtext
+All relevant cluster info, compiler, flags, c++ standard. Test flags and data. Benchmarks metrics collected
+Example:
+System Description:
+The experiments were conducted on a system equipped with an Intel Core i9-12900K CPU (16 cores, 24 threads, base clock 3.2 GHz), 32 GB of DDR5 RAM, and an NVIDIA RTX 3080 GPU (10 GB GDDR6X).
+The operating system was Ubuntu 22.04, and all code was compiled using GCC 12.2 with the -O3 optimization flag.
+Benchmarks were run in an isolated environment with minimal background processes to ensure consistency.
+
+Experimental Setup:
+The study benchmarked three sorting algorithms: QuickSort, MergeSort, and HeapSort.
+Input datasets ranged from 1,000 to 1,000,000 elements and included random, sorted, and reverse-sorted orders.
+Each experiment was repeated 10 times, with execution time and memory usage recorded.
+Performance metrics were averaged across runs to account for variability.
+
+Mamma mia che bel framework che ho
+
+Ho runanto tutto sul pbs con questi cores bellissimi con le ottimizzazioni diverse.
+Ho collezionato i dativi in un basino e l'ho grafato con quella mmerda di pitone.
+Ho fatto i test così perchè mi piacevano.
 
@@
 
-\blindtext\blindtext
+cinquantamila grafici qui
+
+Mamma mia ma quanto e' veloce sto asesmbly,
+buu al compilatore che non ci azzecca per bene. 
+Anche ciclico è nammerda, invece l'unrolling
+è figo.
+In generale abbiamo capito
+cose ed il cluster è pazzesco e le differenze
+sono notevoli, evviva i soldi.
 
@@
 
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
+Ciao ciao
 
@@
 16 November 2024. Section 9A Bilinear Forms and Quadratic Forms, definition 9.11
-
+\bibitem{b13} BOOTHROYD,Z. Algorithm 302: Transpose vector stored array. Go~r~m.A~M I0, ~ (1967),
+292-293
+\bibitem{b14} LAFLIN, S., AND BREFNER, M.A. Algorithm 380: In-situ transposition of a rectangular
+matrix, Comm. ACM IS, 5 (1970), 324-326
+\bibitem{b15} ESKO G. CATE and DAVID W. TWIGG ALGORITHM 513: Analysis of In-Situ Transposition (1977)
+\bibitem{b16} most of the crypto functions are written in assembly usin SIMD instructions in the
+linux kernel "linux/arch/x86/crypto" for exanple transpose\_4x4 in "serpent-avx2-asm\_64.S", "sm4-aesni-avx-asm\_64.S",
+"cast6-avx-x86\_64-asm\_64.S" and others
+\bibitem{b17} https://github.com/torvalds/linux/commit/d34a460092d857f1616e39eed7eac6f40cea2225
+\bibitem{b18} W. O. Alltop, "A Computer Algorithm for Transposing Nonsquare Matrices," in IEEE Transactions on Computers, vol. C-24, no. 10, pp. 1038-1040, Oct. 1975, doi: 10.1109/T-C.1975.224124.
+\bibitem{b19} Jinwoo Suh and V.
+K.
+Prasanna, "An efficient algorithm for out-of-core matrix transposition," in IEEE Transactions on Computers, vol.
+51, no.
+4, pp.
+420-438, April 2002, doi: 10.1109/12.995452.
+\bibitem{b20} Kandemir, M.
+\& Choudhary, A.
+\& Ramanujam, J.
+\& Arunachalam, Meenakshi.
+(2000).
+A unified framework for optimizing locality, parallelism, and communication in out-of-core computations.
+Parallel and Distributed Systems, IEEE Transactions on.
+11.
+648 - 668.
+10.1109/71.877759.
+\bibitem{b21} I-Jui Sung, Juan G´omez-Luna, Josè Marìa Gonzàlez-Linares
+Nicolàs Gui, Wen-Mei W. Hwu. In-Place Transposition of Rectangular Matrices on Accelerators
+\bibitem{b22} Post: An Efficient Matrix Transpose in CUDA C/C++. Feb 18, 2013 By Mark Harris, https://developer.nvidia.com/blog/efficient-matrix-transpose-cuda-cc/
+\bibitem{b23} Hanounik, Bedros. (2000). Diagonal Registers: Novel Vector Register File Design for High Performance and Multimedia Computing. 
 \end{thebibliography}

### 2024-11-25 benchmarks/benchmarks.cpp

@@ -33,6 +33,7 @@
 #define PC_MATRIX_MAX_SIZE 1<<12
 #define PC_RANDOM_MATRIX_SIZE 256
 
+
 /*============================================*\
 |                     SETUP                    |
 \*============================================*/
@@ -165,6 +166,38 @@ BENCHMARK(transpose_cyclic_benchmark,
 
     delete[] arr_in;
     delete[] arr_out;
+}
+
+BENCHMARK(transpose_4x4_intrinsic_cyclic_benchmark,
+	  "matTranspose intrinsic cyclic")
+{
+    /* Initialize the vector */
+    float* arr_in = new float[PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE];
+    float* arr_out = new float[PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE];
+
+    constexpr auto arr1 = random_arr1();
+
+    for (size_t i = 0; i < PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE; ++i)
+      arr_in[i] = arr1[i % PC_RANDOM_MATRIX_SIZE];
+      
+    for (size_t N = 2; N <= 12; ++N)
+    {
+      RUN_BENCHMARK((1<<N),
+		    pc::matTransposeIntrinsicCyclic(arr_in, arr_out, (1<<N)));
+    }
+
+    delete[] arr_in;
+    delete[] arr_out;
+}
+
+BENCHMARK(transpose_4x4_intrinsic_benchmark,
+	  "matTranspose intrinsic")
+{
+    for (size_t N = 2; N <= 12; ++N)
+    {
+      RUN_BENCHMARK((1<<N),
+		    pc::matTransposeIntrinsic(matrix_in, matrix_out, (1<<N)));
+    }
 }
 
 BENCHMARK(transpose_vectorization_benchmark,

### 2024-11-25 include/pc/transpose.hpp

@@ -40,6 +40,8 @@ void matTranspose(float **M, float **T, tenno::size N);
 void matTransposeHalf(float **M, float **T, tenno::size N);
 void matTransposeColumns(float **M, float **T, tenno::size N);
 void matTransposeCyclic(float *M, float *T, tenno::size N);
+void matTransposeIntrinsic(float **mat_in, float **mat_out, size_t N);
+void matTransposeIntrinsicCyclic(float *mat_in, float *mat_out, size_t N);
 
 
 /*============================================*\

### 2024-11-25 src/transpose.cpp

@@ -12,29 +12,10 @@
  * - explicit optimizations: OMP
  */ 
 
-/*
-// With blocks
-inline void transpose_block_SSE4x4(float *A, float *B, const int n, const int m, const int lda, const int ldb ,const int block_size) {
-    #pragma omp parallel for
-    for(int i=0; i<n; i+=block_size) {
-        for(int j=0; j<m; j+=block_size) {
-            int max_i2 = i+block_size < n ? i + block_size : n;
-            int max_j2 = j+block_size < m ? j + block_size : m;
-            for(int i2=i; i2<max_i2; i2+=4) {
-                for(int j2=j; j2<max_j2; j2+=4) {
-                    transpose4x4_SSE(&A[i2*lda +j2], &B[j2*ldb + i2], lda, ldb);
-                }
-            }
-        }
-    }
-
-    // we may directly store the squares sequencially
-}
-*/
-
 #include <pc/transpose.hpp>
 #include <tenno/ranges.hpp>
 #include <omp.h>
+#include <immintrin.h>         /* For AVX intrinsics */
 
 /*============================================*\
 |                   BASELINE                   |
@@ -78,9 +59,8 @@ void pc::matTransposeCyclic(float *M, float *T, tenno::size N) {
     }
 }
 
-/* Linux strikes again arch/x86/crypto/aria-aesni-avc2-asm_64.S */
 /*
-  
+Linux strikes again arch/x86/crypto/aria-aesni-avc2-asm_64.S
 #define transpose_4x4(x0, x1, x2, x3, t1, t2)		\
 	vpunpckhdq x1, x0, t2;				\
 	vpunpckldq x1, x0, x0;				\
@@ -93,7 +73,75 @@ void pc::matTransposeCyclic(float *M, float *T, tenno::size N) {
 							\
 	vpunpckhqdq x2, t2, x3;				\
 	vpunpcklqdq x2, t2, x2;
+
+transpose_4x4_f32_intrinsic is the intrinsic version of the above code
 */
+void transpose_4x4_f32_intrinsic(const float *src0,
+                                 const float *src1, 
+                                 const float *src2,
+                                 const float *src3,
+                                 float *dst0,
+                                 float *dst1,
+                                 float *dst2,
+                                 float *dst3) {
+
+    __m128 row0 = _mm_loadu_ps(src0); // Load Rows
[…]

### 2024-11-26 latex/paper.tex
latex: mothodologie

@@
 - Discuss the challenges faced and how they were addressed \\
+
+In this report we benchmark different implementations of matTranspose
+and checkSymm, comparing implicit and explicit parallelism
+techniques. In particular, the benchmarks will indicate changes in cache and SIMD performance,
+speedup with multiple threads and compile flags optimizations. Data will be compared through tables
+and visually via graphs such as the roofline model. All benchmarks
+will be tested against every matrix sixe from $2^2$ to $2^{12}$
+We will now explain the methodology for matTranspose.
+The algorithms being benchmarked take a $N \times N$ square matrix and transpose it to another $N \times N$ output matrix.
+The following will be benchmarked:
+\begin{itemize}
+\item matTranspose: naive implementation of matTranspose, iterating over rows and columns and
+  swapping the items. (TODO: Code)
+\item matTranspose columns: similar to matTranspose, but iterating first by columns and then by rows.
+  The purpose of this implementation is to test different access patterns.
+  swapping the items. (TODO: Code)
+\item matTranspose half:
+\item matTranspose cyclic:   old algorithms
+\item matTranspose vectorization:
+\item matTranspose unrolling outer:
+\item matTranspose intrinsic:
+\item matTranspose cyclic unrolled:
+\item matTranspose intrinsic unrolled:
+\item matTranspose omp:  all types
+\item matTranspose ched: 3 types
+\end{itemize}
+
+We will now explain the methodology for checkSymm.
 

### 2024-11-26 get-info.sh

@@ -20,4 +20,7 @@ lscpu >> $OUTPUT
 echo "==== uname =====" >> $OUTPUT
 uname -a >> $OUTPUT 
 
+echo "==== coherency_line_size =====" >> $OUTPUT
+cat /sys/devices/system/cpu/cpu0/cache/index0/coherency_line_size >> $OUTPUT
+
 echo "Done"

### 2024-11-27 latex/paper.tex
latex: pseudocode

--- 2024-11-25 latex/paper.tex
@@
 \usepackage{xcolor}
+\usepackage[noend]{algpseudocode}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
+\iffalse % pseudocode
+  \begin{algorithm}
+    \caption{My Algorithm}\label{euclid}
+    \begin{algoritmic}[1]
+      \procedure{MyProcedure}{}
+      % use \state and \textit{} for variables
+      % use \text for text
+      % use \gets for assignments
+      % use \If \EndIf \Return \Else
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+\endif
+
 - Elaborate on the unique contributions of your project \\
@@
 - Discuss the challenges faced and how they were addressed \\
+
+In this report we benchmark different implementations of matTranspose
+and checkSymm, comparing implicit and explicit parallelism
+techniques. In particular, the benchmarks aim to increase cache and SIMD performance,
+speedup with multiple threads and the impact of compile flags optimizations. Data will be compared through tables
+and visually via graphs such as the roofline model. All benchmarks
+will be tested against every matrix sixe from $2^2$ to $2^{12}$, we will assume the size is a multiple of 4 for ease of implementation.
+We will now explain the methodology for matTranspose.
+The algorithms being benchmarked take a $N \times N$ input square matrix, an output matrix with the same shape and the size of the matrix.
+The following will be benchmarked:
+\begin{itemize}
+\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix and storing them in the output matrix with rows and columns indexes swapped.
+
+  \begin{algorithm}
+    \caption{matTranspose}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{matTranspose}{$Matrix input, Matrix output, int size$}
+        \For{\textit{r} \textbf{ in } \textit{0..size}}
+            \State \For{\textit{c} \textbf{ in } \textit{0..size}}
+            \State $output[c][r] = input[r][c]$
+            \EndFor
+        \EndFor
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+    \item \textit{matTransposeColumns}: similar to \textit{matTranspose}, but iterating on the input matrix first by columns and then by rows.
+  The purpose of this implementation is to test a different access pattern.
+
+  \begin{algorithm}
+    \caption{matTranspose columns}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{matTransposeColumns}{$Matrix input, Matrix output, int size$}
+        \For{\textit{r} \textbf{ in } \textit{0..size}}
+            \State \For{\textit{c} \textbf{ in } \textit{0..size}}
+            \State $output[r][c] = input[c][r]$
+            \EndFor
+        \EndFor
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+\item \item{matTransposeHalf}: similar to \textit{matTranspose}, but the algorithm
+  iterates only on the upper half of the input matrix. For each
+  pair $(i,j)$ of indices for rows and columns in the upper half of the input matrix $M$, the output matrix O will
+  store store $O[i][j]=M[j][i]$ and $O[j][i] = M[i][j]$.
+
+  \begin{algorithm}
+    \caption{matTranspose half}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{matTransposeHalf}{$Matrix input, Matrix output, int size$}
+        \For{\textit{r} \textbf{ in } \textit{0..size}}
+            \State \For{\textit{c} \textbf{ in } \textit{r..size}}
+            \State $output[r][c] = input[c][r]$
+            \State $output[c][r] = input[r][c]$
+            \EndFor
+        \EndFor
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+\item \item{matTransposeCyclic}: implementation of \cite{b13}, storing the matrix as a contiguous one-dimensional array of size $N*N$.
+
+  \begin{algorithm}
+    \caption{matTranspose cyclic}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{matTransposeCyclic}{$Matrix input, Matrix output, int size$}
+        \For{\textit{n} \textbf{ in } \textit{0..(size*size)}}
+        \State int \textit{i} = $\frac{n}{size}$
+        \State int \textit{j} = $n \mod N$
+        $T[n] = M[size * j + i]$
+        \EndFor
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+\item \textit{matTransposeIntrinsic}: based on \textit{matrix\_transpose4x4} from the linux kernel \cite{b16}, revised to work for out-of-place float matrices on x86 architectures and implemented using intrinsics.
+\textit{transpose\_4x4\_f32\_intrinsic} transposes a 4 by 4 matrix using vector registers of 128 bits by doing the following steps: load 4 rows of 4 floats from the the input matrix, combine pairs of rows in an alternating pattern (interleaving), shuffle to get the transpose and store them in the output matrix.
+\textit{matTransposeIntrinsic} iterates over 4x4 blockw in the input matrix and saves the computed transposed block in the output matrix with rows and columns swapped.
+
+  \begin{algorithm}
+    \caption{matTranspose intrinsic}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{transpose\_4x4\_f32\_intrinsic}{float* src0, float* src1, float* src2, float* src3, float* dst0, float* dst1, float *dst2, float* dst3}
+    \textbf{comment} Load rows  
+    \_\_m128 row0 = \_mm\_loadu\_ps(src0);
+    \_\_m128 row1 = \_mm\_loadu\_ps(src1);
+    \_\_m128 row2 = \_mm\_loadu\_ps(src2);
+    \_\_m128 row3 = \_mm\_loadu\_ps(src3);
+
+    \textbf{comment} unpack and interleave rows
+    \_\_m128 t0 = \_mm\_unpacklo\_ps(row0, row1);
+    \_\_m128 t1 = \_mm\_unpackhi\_ps(row0, row1);
+    \_\_m128 t2 = \_mm\_unpacklo\_ps(row2, row3);
+    \_\_m128 t3 = \_mm\_unpackhi\_ps(row2, row3);
+
+    \textbf{comment} Shuffle to form transposed rows
+    \_\_m128 d0 = \_mm\_movelh\_ps(t0, t2);
+    \_\_m128 d1 = \_mm\_movehl\_ps(t2, t0);
+    \_\_m128 d2 = \_mm\_movelh\_ps(t1, t3);
+    \_\_m128 d3 = \_mm\_movehl\_ps(t3, t1);
+
+    \textbf{comment} Store transposed rows
+    \_mm\_storeu\_ps(dst0, d0);
+    \_mm\_storeu\_ps(dst1, d1);
+    \_mm\_storeu\_ps(dst2, d2);
+    \_mm\_storeu\_ps(dst3, d3);
+
+      \EndProcedure
+      \procedure{matTransposeIntrinsic}{$Matrix input, Matrix output, int size$}
+        \For{int i = 0, i < size, i += 4}
+        \For{int j = 0, j < size, j += 4}
+        \State transpose\_4x4\_f32\_intrinsic(input[i][j],
+        input[i+1][j], input[i+2][j], input[i+3][j], output[j][i], output[j+1][i], output[j+2][i], output[j+3][i])
+        \EndFor
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+\item \textit{matTransposeIntrinsicCyclic}: same as \textit{matTransposeIntrinsic} but the matrix is saved as a one dimensional array similarly to \textit{matTransposeCyclic}
+%\item matTranspose vectorization:
+%\item matTranspose half vectorization:
+\item \textit{matTransposeUnrollingOuter}: similar to \textit{matTranspose} but the outer loop is manually unrolled
+
+  \begin{algorithm}
+    \caption{matTransposeUnrollingOuter}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{matTransposeUnrollingOuter}{$Matrix input, Matrix output, int size$}
+        \For{$int c = 0, i < size, i += 4$}
+            \State \For{\textit{r} \textbf{ in } \textit{0..size}}
+            \State $output[c][r] = input[r][c]$
+            \State $output[c+1][r] = input[r][c+1]$
+            \State $output[c+2][r] = input[r][c+2]$
+            \State $output[c+3][r] = input[r][c+3]$
+            \EndFor
+        \EndFor
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+\item \textit{matTransposeUnrollingInner}: same as \textit{matTransposeUnrollingOuter} but the inner for loop is manually unrolled
+
+\item \textit{matTransposeHalfUnrollingOuter}: same as \textit{matTransposeHalf} but the outer for loop is manually unrolled
+
+\item \textit{matTransposeHalfUnrollingInner}: same as \textit{matTransposeHalf} but the inner for loop is manually unrolled
+
+\item \textit{matTransposeCyclicUnrolled}: same as \textit{matTransposeCyclic} but the for loop in manually unrolled
+
+  \begin{algorithm}
+    \caption{matTranspose cyclic unrolled}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{matTransposeCyclicUnrolled}{$Matrix input, Matrix output, int size$}
+        \For{\textit{n} \textbf{ in } \textit{0..(size*size)}}
+        $T[n] = M[size * ((n+1) \mod size ) + (\frac{n}{size})]$
+        $T[n+1] = M[size * ((n+1+1) \mod size ) + (\frac{n+1}{size})]$
+        $T[n+2] = M[size * ((n+2+1) \mod size ) + (\frac{n+2}{size})]$
+        $T[n+3] = M[size * ((n+3+1) \mod size ) + (\frac{n+3}{size})]$
+        \EndFor
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+\item \textit{matTransposeOmpX}: this set of algorithms are all derived from \textit{matTranspose} with the outer for loops parallelized using omp with the number of threads \textit{X} being a power of 2 from $2^1$ to $2^6$.
+
+\item \textit{matTransposeOmpXCollapse}: similar to \textit{matTransposeOmpX} but the for loops are parellelized using \textit{collapse(2)}
+
+\item \textit{matTransposeOmp16SchedX}: similar to \textit{matTransposeOmpX} with 16 threads using different schedule algorithms, namely: static, synamic, guided.
+\end{itemize}
+
+We will now explain the methodology for \textit{checkSymm}. The algorithms take an $N \times N$ square matrix and Its size are return true if the matrix is symmetrix, false otherwise.
+The algorithms being benchmarked are the wollowing:
+\being{itemize}
+\item \textit{checkSymm}: naive implementation of checkSymm. Iterates over the upper half of the matrix and testing if every value is equal to the symmetric value.
+
+  \begin{algorithm}
+    \caption{checkSymm}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{checkSymm}{$Matrix input, int size$}
+        \State $isSymm = true$
+        \For{\textit{i} \textbf{ in } \textit{0..size}}
+            \For{\textit{j} \textbf{ in } \textit{i..size}}
+            \If $input[i][j] != input[j][i]$
+            \State $isSymm = false$
+        \EndFor
+        \Return $isSymm$
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+% \item \textit{checkSymColumns}:
+
+  % \item \textit{checkSymVectorization}
+\item \textit{checkSymmUnrolledOuter}: similar to \textit{checkSymm} but the outer loop is manually unrolled.
+
+  \begin{algorithm}
+    \caption{checkSymmUnrolledOuter}\label{matTranspose}
+    \begin{algoritmic}[1]
+      \procedure{checkSymmUnrolledOuter}{$Matrix input, int size$}
+        \State $isSymm = true$
+        \For{\textit{i} \textbf{ in } \textit{0..size}}
+            \For{\textit{j} \textbf{ in } \textit{i..size}}
+            \If $input[i][j] != input[j][i]$
+            \State $isSymm = false$
+            \If $input[i+1][j] != input[j][i+1]$
+            \State $isSymm = false$
+            \If $input[i+2][j] != input[j][i+2]$
+            \State $isSymm = false$
+            \If $input[i+3][j] != input[j][i+3]$
+            \State $isSymm = false$
+        \EndFor
+        \Return $isSymm$
+      \EndProcedure
+    \end{algoritmic}
+  \end{algorithm}
+
+\item \textit{checkSymmUnrollingInner}: similar to \textit{checkSymmUnrolledOuter} but the inner loop is manually unrolled instead of the outer loop.
+\item \textit{checkSymmUnrollingOuterNoBranch}:
+\item \textit{checkSymmUnrollingInnerNoBranch}:
+\item \textit{checkSymOmpX}:
+\item \textit{checkSymOmpXCollapse}:
+\item \textit{checkSymm16SchedX}:
+\end{itemize}
 

### 2024-11-27 src/transpose.cpp

@@ -40,15 +40,12 @@ void pc::matTransposeHalf(float **M, float **T, tenno::size N)
   return;
 }
 
-/* access pattern */
-
 void pc::matTransposeColumns(float **M, float **T, tenno::size N)
 {
   for (tenno::size i = 0; i < N; ++i)
       for (tenno::size j = 0; j < N; ++j)
             T[j][i] = M[i][j];
   return;
-
 }
 
 void pc::matTransposeCyclic(float *M, float *T, tenno::size N) {
@@ -240,6 +237,7 @@ void pc::matTransposeCyclicUnrolled(float *M, float *T, tenno::size N) {
         T[n+3] = M[N*((n+3)%N) + ((n+3)/N)];
     }
 }
+
 
 /*============================================*\
 |              EXPLICIT PARALLELISM            |

### 2024-11-27 latex/paper.tex
report: completed section 3

--- 2024-11-25 latex/paper.tex
@@
 \usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
+\usepackage{algorithm}
+% \usepackage{algorithmic}
 \usepackage{graphicx}
@@
 \usepackage{xcolor}
+\usepackage{algpseudocode}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
+\iffalse
+
 \textit{We define the transpose of a matrix $A$, and we denote it by $A^t$, as the matrix obtained from $A$ by interchanging rows and columns.
@@
 \end{equation}
+
+\fi
 
@@
 SIMD implementations for specific architectures are also widely adopted
-in performance critical applications like cryptographic libraries using instructions like Intel's vpunpckldq and vpunpckhdq \cite{b16} reporting improvements up to 70\% \cite{b17}.
+in performance critical applications like cryptographic libraries, utilizing instructions like Intel's vpunpckldq and vpunpckhdq \cite{b16} and reporting improvements up to 70\% \cite{b17}.
 However, matrix operations are not efficient as vector registers access data only linearly, motivating research on new vector register file (VFS) designs such as diagonal registers \cite{b23}.
@@
 
+\iffalse
 \textit{A square matrix $A$ is called symmetric if it equals its transpose} \cite{b12}.
-
+\fi 
 While symmetric matrces are actively used in various areas of research as their
@@
 
-- Elaborate on the unique contributions of your project \\
-- Describe the methodology proposed, including algorithms (pseudo-code),
-data structures, and parallelization techniques. \\
-- Discuss the challenges faced and how they were addressed \\
+\iffalse % pseudocode
+    \begin{algorithmic}[1]
+      \Procedure{MyProcedure}{}
+      % use \state and \textit{} for variables
+      % use \text for text
+      % use \gets for assignments
+      % use \If \EndIf \Return \Else
+      \EndProcedure
+    \end{algorithmic}
+\fi
+
+In this report different implementations of matTranspose
+and checkSymm are benchmarked, comparing implicit and explicit parallelism
+techniques. In particular, the benchmarks aim to indicate cache and SIMD performance differences,
+speedup with multiple threads and the impact of compile flags optimizations. Data will be compared through tables
+and visually via graphs such as the roofline model. All benchmarks
+will be tested against every matrix size from $2^2$ to $2^{12}$, we will assume the size value is a multiple of 4 for ease of implementation.  \\
+The algorithms for \textit{matTranspose} will now be explained.
+Each algorithm takes a $N \times N$ input square matrix \textit{input}, a matrix of the same shape \textit{ouput}, and the size of the two matrices \textit{size}.
+The analyzed algorithms are the following:
+\begin{itemize}
+\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix and storing them in the output matrix with indexes of rows and columns swapped.
+See Algorithm~\ref{matTranspose} for the pseudocode.
+
+\begin{algorithm}
+	\caption{matTranspose}\label{matTranspose}
+	\begin{algorithmic}[1]
+		\For{\textit{r} \textbf{ in } \textit{0..size-1}}
+	    \For{\textit{c} \textbf{ in } \textit{0..size-1}}
+		\State $output[c][r] \gets input[r][c]$
+		\EndFor
+		\EndFor
+	\end{algorithmic}
+\end{algorithm}
+
+    \item \textit{matTransposeColumns}: similar to \textit{matTranspose}, but iterating on the input matrix first by columns and then by rows.
+  The purpose of this implementation is to test a different access pattern.
+
+\iffalse
+\begin{algorithm}
+	\caption{matTransposeColumns}\label{matTransposeColumns}
+\begin{algorithmic}[1]
+	\For{\textit{r} \textbf{ in } \textit{0..size-1}}
+	 \For{\textit{c} \textbf{ in } \textit{0..size-1}}
+            \State $output[r][c] \gets input[c][r]$
+            \EndFor
+        \EndFor
+\end{algorithmic}
+\end{algorithm}
+\fi
+
+  \item \textit{matTransposeHalf}: similar to \textit{matTranspose}, but the algorithm
+  iterates only on the upper half of the input matrix. For each
+  pair $(i,j)$ of indices of rows and columns in the upper half of the input matrix $M$, with output matrix $O$, $O[i][j]=M[j][i]$ and $O[j][i] = M[i][j]$.
+\iffalse
+\begin{algorithm}
+	\caption{matTransposeHalf}\label{matTransposeHalf}
+\begin{algorithmic}[1]
+	\For{\textit{r} \textbf{ in } \textit{0..size-1}}
+           \For{\textit{c} \textbf{ in } \textit{r..size-1}}
+            \State $output[r][c] \gets input[c][r]$
+            \State $output[c][r] \gets input[r][c]$
+            \EndFor
+        \EndFor
+\end{algorithmic}
+\end{algorithm}
+\fi
+
+\item \textit{matTransposeCyclic}: implementation of \cite{b13}, storing the matrix as a contiguous one-dimensional array of size $N*N$. See Algorithm~\ref{matTransposeCyclic} for the pseudocode.
+  \begin{algorithm}
+  	\caption{matTransposeCyclic}\label{matTransposeCyclic}
+  \begin{algorithmic}[1]
+
+  	  \For{\textit{n} \textbf{ in } \textit{0..size*size-1}}
+        \State $int\ \textit{i} \gets \frac{n}{size}$
+        \State $int\ \textit{j} \gets n \mod N$
+        \State $T[n] \gets M[size * j + i]$
+        \EndFor
+          \end{algorithmic}
+  \end{algorithm}
+
+\item \textit{matTransposeIntrinsic}: based on \textit{matrix\_transpose4x4} from the linux kernel \cite{b16}, revised to work for out-of-place float matrices on x86 architectures and implemented using intrinsics.
+\textit{transpose\_4x4\_f32\_intrinsic} transposes a 4 by 4 matrix using vector registers of 128 bits by doing the following steps: load 4 rows of 4 floats from the the input matrix, combine pairs of rows in an alternating pattern (interleaving), shuffle to get the transpose and store them in the output matrix.
+\textit{matTransposeIntrinsic} iterates over 4x4 blocks in the input matrix, calling \textit{transpose\_4x4\_f32\_intrinsic} and saving the computed transposed block in the output matrix with rows and columns swapped.
+See Algorithm~\ref{matTransposeIntrinsic} for the pseudocode.
+  \begin{algorithm}
+	\caption{matTransposeIntrinsic}\label{matTransposeIntrinsic}
+	    \begin{algorithmic}[1]
+	\iffalse
+
+      \Procedure{transpose\_4x4\_f32\_intrinsic}{Matrix4x4 in, Matrix4x4 out}
+    \State \textbf{comment} Load rows  
+    \State \textbf{comment} unpack and interleave rows
+    \State \textbf{comment} Shuffle to form transposed rows
+    \State \textbf{comment} Store transposed rows
+      \EndProcedure
+
+      \fi
+
+        \For{$int\ i = 0, i < size, i += 4$}
+        \For{$int\ j = 0, j < size, j += 4$}
+        \State $transpose\_4x4\_f32\_intrinsic(input[i][j],$
+        \State $                               output[j][i])$
+        \EndFor
+        \EndFor
+
+    \end{algorithmic}
+    \end{algorithm}
+
+\item \textit{matTransposeIntrinsicCyclic}: same as \textit{matTransposeIntrinsic} but the matrix is saved as a one dimensional array similarly to \textit{matTransposeCyclic}
+%\item matTranspose vectorization:
+%\item matTranspose half vectorization:
+\item \textit{matTransposeUnrollingOuter}: similar to \textit{matTranspose} but the outer loop is manually unrolled. See Algorithm~\ref{matTransposeUnrollingOuter} for the pseudocode.
+
+  \begin{algorithm}
+    \caption{matTransposeUnrollingOuter}\label{matTransposeUnrollingOuter}
+    \begin{algorithmic}[1]
+        \For{$int\ c = 0, i < size, i += 4$}
+            \For{\textit{r} \textbf{ in } \textit{0..size-1}}
+            \State $output[c][r] \gets input[r][c]$
+            \State $output[c+1][r] \gets input[r][c+1]$
+            \State $output[c+2][r] \gets input[r][c+2]$
+            \State $output[c+3][r] \gets input[r][c+3]$
+            \EndFor
+        \EndFor
+    \end{algorithmic}
+  \end{algorithm}
+
+\item \textit{matTransposeUnrollingInner}: same as \textit{matTransposeUnrollingOuter} but the inner for loop is manually unrolled
+
+\item \textit{matTransposeHalfUnrollingOuter}: same as \textit{matTransposeHalf} but the outer for loop is manually unrolled
+
+\item \textit{matTransposeHalfUnrollingInner}: same as \textit{matTransposeHalf} but the inner for loop is manually unrolled
+
+\item \textit{matTransposeCyclicUnrolled}: same as \textit{matTransposeCyclic} but the for loop in manually unrolled
+\iffalse
+  \begin{algorithm}
+    \caption{matTranspose cyclic unrolled}\label{matTransposeCyclicUnrolled}
+    \begin{algorithmic}[1]
+      \Procedure{matTransposeCyclicUnrolled}{$Matrix input, Matrix output, int size$}
+        \For{\textit{n} \textbf{ in } \textit{0..size*size-1}}
+        \State $T[n] \gets M[size * ((n+1) \mod size ) + (\frac{n}{size})]$
+        \State $T[n+1] \gets M[size * ((n+1+1) \mod size ) + (\frac{n+1}{size})]$
+        \State $T[n+2] \gets M[size * ((n+2+1) \mod size ) + (\frac{n+2}{size})]$
+        \State $T[n+3] \gets M[size * ((n+3+1) \mod size ) + (\frac{n+3}{size})]$
+        \EndFor
+      \EndProcedure
+    \end{algorithmic}
+  \end{algorithm}
+  \fi
+\item \textit{matTransposeOmpX}: this set of algorithms are all derived from \textit{matTranspose} with the outer for loops parallelized using omp with the number of threads \textit{X} being a power of 2 from $2^1$ to $2^6$.
+
+\item \textit{matTransposeOmpXCollapse}: similar to \textit{matTransposeOmpX} but the for loops are parellelized using \textit{collapse(2)}
+
+\item \textit{matTransposeOmp16SchedX}: similar to \textit{matTransposeOmpX} with 16 threads using different schedule algorithms, namely: static, dynamic, guided.
+\end{itemize}
+
+The algorithms for \textit{checkSymm} will now be explained.
+Each algorithm take an $N \times N$ square matrix \textit{input}, a size \textit{size} and return true if the matrix is symmetrix, false otherwise.
+The analyzed algorithms are the following:
+\begin{itemize}
+\item \textit{checkSymm}: naive implementation of checkSymm.
+It Iterates over the upper half of the matrix and testing if every value is equal to the symmetric value.
+See Algorithm~\ref{checkSymm} for the pseudocode.
+
+  \begin{algorithm}
+    \caption{checkSymm}\label{checkSymm}
+    \begin{algorithmic}[1]
+        \State $\textbf{bool}\ isSymm \gets true$
+        \For{\textit{i} \textbf{ in } \textit{0..size-1}}
+            \For{\textit{j} \textbf{ in } \textit{i..size-1}}
+            	\If{$input[i][j] != input[j][i]$}
+            	\State $isSymm \gets \textbf{false}$
+            	\EndIf
+        	\EndFor
+        \EndFor
+        \State \Return $isSymm$
+    \end{algorithmic}
+  \end{algorithm}
+
+% \item \textit{checkSymColumns}:
+
+  % \item \textit{checkSymVectorization}
+\item \textit{checkSymmUnrolledOuter}: similar to \textit{checkSymm} but the outer loop is manually unrolled.
+
+\iffalse
+  \begin{algorithm}
+    \caption{checkSymmUnrolledOuter}\label{checkSymmUnrolledOuter}
+    \begin{algorithmic}[1]
+      \Procedure{checkSymmUnrolledOuter}{$Matrix input, int size$}
+        \State $isSymm = true$
+        \For{\textit{i} \textbf{ in } \textit{0..size-1}}
+            \For{\textit{j} \textbf{ in } \textit{i..size-1}}
+            \If $input[i][j] != input[j][i]$
+            \State $isSymm = false$
+            \EndIf
+            \If $input[i+1][j] != input[j][i+1]$
+            \State $isSymm = false$
+                        \EndIf
+            \If $input[i+2][j] != input[j][i+2]$
+            \State $isSymm = false$
+                        \EndIf
+            \If $input[i+3][j] != input[j][i+3]$
+            \State $isSymm = false$
+                        \EndIf
+        \EndFor
+        \EndFor
+        \Return $isSymm$
+      \EndProcedure
+    \end{algorithmic}
+  \end{algorithm}
+\fi
+
+\item \textit{checkSymmUnrollingInner}: similar to \textit{checkSymmUnrolledOuter} but the inner loop is manually unrolled instead of the outer loop.
+\item \textit{checkSymmUnrollingOuterNoBranch}: similar to \textit{checkSymmUnrollingOuter}, but 4 pairs of values are checked for equality with a single math formula.
+The formula performs a XOR on 4 pairs of values and their transpose, if any of those are different than 0 then the matrix is not symmetric.
+See Arlgoritm~\ref{checkSymmUnrollingOuterNoBranch} for the pseudocode.
+
+\begin{algorithm}
+	\caption{checkSymmUnrollingOuterNoBranch}\label{checkSymmUnrollingOuterNoBranch}
+	\begin{algorithmic}[1]
+		\State $\textbf{bool}\ isSymm \gets true$
+		\For{$int\ i = 0; i < size; i += 4$}
+		\For{$int\ j = i; j < size; j++$}
+		\State $isSymm \gets isSymm \mid (((input[i][j] \oplus input[j][i]) \mid (input[i+1][j] \oplus input[j][i+1]) \mid (input[i+2][j] \oplus input[j][i+2]) \mid (input[i+3][j] \oplus input[j][i+3])) != 0)$
+		\EndFor
+		\EndFor
+		\State \Return $isSymm$
+	\end{algorithmic}
+\end{algorithm}
+
+\item \textit{checkSymmUnrollingInnerNoBranch}: similar to \textit{checkSymmUnrollingOuterNoBranch} but the inner loop is manually unrolled.
+\item \textit{checkSymOmpX}: similar to \textit{matTransposeOmpX} but based on \textit{checkSymm}. After each parallel section, a barrier is needed to synchronize the shared boolean output variable.
+\item \textit{checkSymOmpXCollapse}: similar to \textit{checkSymmOmpX} bur the two loops are parallelized with collapse(2).
+\item \textit{checkSymm16SchedX}: similar to \textit{matTranspose16SchedX} but based on \textit{matTranspose}.
+\end{itemize}
 
@@
 \fi
-
-Quiesti qui bellissimi e stra fighi
-- uno (pseudocodice)
-- due (pseudocodice)
-- tre (pseudocodice)
-- ...
-Ho usato implicit ed esplicit parallelism con vectorization e openmp
-Voglio fare questo per comparare quest'altro cos' pel meme
-Benchmarkare è difficle perchè... quindi ho fatto ciò... (aplia qua)
 
@@
 or achieve the objectives \\
+
+The experiments were conducted on a cluster equipped with 64 x86\_64 cpus, more specifically Intel(R) Xeon(R) Gold 5220 CPU @ 2.20GHz.
 

### 2024-11-27 src/check_symm.cpp

@@ -89,13 +89,13 @@ bool pc::checkSymUnrollingOuterNoBranch(float **M, tenno::size N)
   for (size_t i = 0; i < N - 4; i += 4)
     for (size_t j = i; j < N; ++j)
       {
-	symm &=
+	symm |=
 	        ((
-		  ((uint32_t) M[i][j] - (uint32_t) M[j][i])
-	        | ((uint32_t) M[i+1][j] - (uint32_t) M[j][i+1])
-                | ((uint32_t) M[i+2][j] - (uint32_t) M[j][i+2])
-                | ((uint32_t) M[i+3][j] - (uint32_t) M[j][i+3]))
-         	== 0); // normalize the results
+		  ((uint32_t) M[i][j] ^ (uint32_t) M[j][i])
+	        | ((uint32_t) M[i+1][j] ^ (uint32_t) M[j][i+1])
+                | ((uint32_t) M[i+2][j] ^ (uint32_t) M[j][i+2])
+                | ((uint32_t) M[i+3][j] ^ (uint32_t) M[j][i+3]))
+         	!= 0); // normalize the results
       }
   return symm;
 }

### 2024-11-27 src/transpose.cpp

@@ -49,10 +49,8 @@ void pc::matTransposeColumns(float **M, float **T, tenno::size N)
 }
 
 void pc::matTransposeCyclic(float *M, float *T, tenno::size N) {
-    for(long unsigned int n = 0; n<N*N; n++) {
-        long unsigned int i = n/N;
-        long unsigned int j = n%N;
-        T[n] = M[N*j + i];
+    for(long unsigned int n = 0; n<N*N; ++n) {
+        T[n] = M[N*(n%N) + n/N];
     }
 }
 

### 2024-11-29 latex/paper.tex
report: experiments and system description draft

@@
 
+\iffalse
 - Detailed description of the computing system and platform. \\
@@
 or achieve the objectives \\
+\fi
 
-The experiments were conducted on a cluster equipped with 64 x86\_64 cpus, more specifically Intel(R) Xeon(R) Gold 5220 CPU @ 2.20GHz.
+The experiments were conducted on a cluster equipped with 64 Intel(R) Xeon(R) Gold 5220 CPUs at 2.20GHz, 32K of L1d cache, 32K L1i cache, 1025K L2 cache, 25344K L3 cache with cache lines of 64 bits, 1 thread per core, 4GB of DDR4 RAM running Linux 3.10.
+The algorithms were benchmarked using valFuzz \ref{b?}.
+The set of benchmarks were compiled with flags "-fopenmp", "-fopenmp-simd", 4 set of benchmarks are compiles with either no optimizations, -O1, -O2, -O3 -Ofast -march=native.
+To obtain results with statistical significance, each benchmark has been executed 100 times collecting the following data concerning wall-clock execution time: minimum, maximum, median, arithmetic mean, standard deviation, Q1, Q3.
+To prevent interference between the benchmarks, the cache is cleared by initializing an array twice as big as the L3 cache size.
+To ensure fair comparison, the same input matrix is provided to all algorithms.
+The input matrix is initialized with an uniform random distribution of floats ranging from -1 to 1 generated using lcg algorithm \ref{b?} with the parameters a=1664525 and c = 1013904223.
+To save time, generation of random numbers is done at compile time using tenno-tl \ref{b?} with the initial seed "1337".
+Additional algorithms and data structures were used from tenno-tl including ranges and arrays.
 

### 2024-12-01 src/check_symm.cpp
report: finalized report

@@ -89,13 +89,13 @@ bool pc::checkSymUnrollingOuterNoBranch(float **M, tenno::size N)
   for (size_t i = 0; i < N - 4; i += 4)
     for (size_t j = i; j < N; ++j)
       {
-	symm |=
+	symm &=
 	        ((
 		  ((uint32_t) M[i][j] ^ (uint32_t) M[j][i])
 	        | ((uint32_t) M[i+1][j] ^ (uint32_t) M[j][i+1])
                 | ((uint32_t) M[i+2][j] ^ (uint32_t) M[j][i+2])
                 | ((uint32_t) M[i+3][j] ^ (uint32_t) M[j][i+3]))
-         	!= 0); // normalize the results
+         	== 0); // normalize the results
       }
   return symm;
 }
@@ -108,10 +108,10 @@ bool pc::checkSymUnrollingInnerNoBranch(float **M, tenno::size N)
       {
 	symm &=
 	        ((
-		  ((uint32_t) M[i][j] - (uint32_t) M[j][i])
-	        | ((uint32_t) M[i][j+1] - (uint32_t) M[j+1][i])
-                | ((uint32_t) M[i][j+2] - (uint32_t) M[j+2][i])
-                | ((uint32_t) M[i][j+3] - (uint32_t) M[j+3][i]))
+		  ((uint32_t) M[i][j] ^ (uint32_t) M[j][i])
+	        | ((uint32_t) M[i][j+1] ^ (uint32_t) M[j+1][i])
+                | ((uint32_t) M[i][j+2] ^ (uint32_t) M[j+2][i])
+                | ((uint32_t) M[i][j+3] ^ (uint32_t) M[j+3][i]))
          	== 0); // normalize the results
       }
   return symm;

### 2024-12-01 src/transpose.cpp

@@ -144,7 +144,7 @@ void pc::matTransposeIntrinsic(float **mat_in, float **mat_out, size_t N)
 
 void pc::matTransposeVectorization(float **M, float **T, tenno::size N)
 {
-  #pragma omp simd
+  #pragma omp for simd
   for (tenno::size i = 0; i < N; ++i)
       for (tenno::size j = 0; j < N; ++j)
             T[i][j] = M[j][i];

### 2024-12-01 latex/paper.tex
report: final touches

--- 2024-11-27 latex/paper.tex
@@
 \usepackage{algpseudocode}
+\usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 
-\title{Matrix Transposition and Symmetry: a case study on Superscalar Architectures\\
+\title{Matrix Transposition and Symmetry in Superscalar Architectures: A Case Study\\
 %{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 Trento, Italy \\
-giovanni.santini@unitn.it \\
+giovanni.santini@studenti.unitn.it \\
 MAT. 235441}
@@
 \begin{abstract}
-TODO *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+Contemporary CPU designs achieve higher performance by integrating multicore processing with sophisticated Single Instruction Multiple Data (SIMD) capabilities.
+This study investigates the performance of two algorithms—square matrix transposition and symmetry checking—on modern superscalar CPU architectures.
+Both implicit and explicit parallelism techniques will be discussed, as well as the effects of compiler optimization flags.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-parallel computing, matrix transpose, symmetrix matrix. benchmarking
+parallel computing, matrix transpose, symmetric matrix, benchmark
 \end{IEEEkeywords}
@@
 
-Modern CPUs use superscalar architectures to improve performance, combining multiple cores with advanced Single Instruction Multiple Data (SIMD) technologies.
+Modern CPUs make use of superscalar architectures to improve performance, combining multiple cores with advanced Single Instruction Multiple Data (SIMD) technologies.
 Examples include Intel's Streaming SIMD Extensions (SSE) and Advanced Vector Extensions (AVX).
@@
 In section 2, we discuss the state of the art of the two algorithms.
-In section 3, we elaborate on the algorithms analyzed, discuss performance evaluation and we present a new benchmarking framework.
-In section 4, we show experimental data and graphs that is analyzed
-in section 5. The conclusion is in section 6.
+In section 3, we elaborate on the algorithms analyzed.
+In section 4, we explain the benchmarking procedures and system information.
+In section 5, we present the results. The conclusion is in section 6. Section 7 contains
+the project's github link.
 
@@
 work focuses on in-situ transporision, that is transposing a matrix without
-relying on additional space. Implementations were often based on the cyclic
-structure of the transposition mapping: values of a matrix of size $N$ can be stored
-in a single contiguous array or size $N$*$N$ and indexed with $row*\#rows+ column$.
-Remarcable contributions include Boothroyd's ACM Algorithm 302 (1966) using cyclic mapping \cite{b13}, Laflin and Brefner ACM Algorithm 380
+relying on additional space. Remarcable contributions include Boothroyd's ACM Algorithm 302 (1966) using cyclic mapping \cite{b13}, Laflin and Brefner ACM Algorithm 380
 adding fixed points arithmetic \cite{b14} and
@@
 in performance critical applications like cryptographic libraries, utilizing instructions like Intel's vpunpckldq and vpunpckhdq \cite{b16} and reporting improvements up to 70\% \cite{b17}.
-However, matrix operations are not efficient as vector registers access data only linearly, motivating research on new vector register file (VFS) designs such as diagonal registers \cite{b23}.
+However, vector registers are unoptimal for matrices since registers read data only sequencially, motivating research on new vector register file (VFS) designs such as diagonal registers \cite{b23}.
 Another area of research includes out-of-core matrix transposition, that
@@
 \fi 
-While symmetric matrces are actively used in various areas of research as their
+In constrast, while symmetric matrces are actively used in various areas of research as their
 properties allow for several optimizations in various problems, an in depth case study of symmetry
@@
 
-In this report different implementations of matTranspose
-and checkSymm are benchmarked, comparing implicit and explicit parallelism
-techniques. In particular, the benchmarks aim to indicate cache and SIMD performance differences,
-speedup with multiple threads and the impact of compile flags optimizations. Data will be compared through tables
-and visually via graphs such as the roofline model. All benchmarks
-will be tested against every matrix size from $2^2$ to $2^{12}$, we will assume the size value is a multiple of 4 for ease of implementation.  \\
+In this report, different implementations of \textit{matTranspose}
+and \textit{checkSymm} are benchmarked, comparing implicit and explicit parallelism
+techniques. In particular, the benchmarks aim to indicate cache and SIMD performance,
+speedup with multiple threads and the impact of compile flags. Data will be compared visually via graphs. All benchmarks
+will be tested against every matrix size of power of $2$ from $2^2$ to $2^{12}$.  \\
 The algorithms for \textit{matTranspose} will now be explained.
@@
 \begin{itemize}
-\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix and storing them in the output matrix with indexes of rows and columns swapped.
+\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix by rows and columns and storing them in the output matrix with indexes swapped.
 See Algorithm~\ref{matTranspose} for the pseudocode.
@@
 
-\item \textit{matTransposeCyclic}: implementation of \cite{b13}, storing the matrix as a contiguous one-dimensional array of size $N*N$. See Algorithm~\ref{matTransposeCyclic} for the pseudocode.
+\item \textit{matTransposeCyclic}: implementation of \cite{b28}, storing the matrix as a contiguous one-dimensional array of size $N*N$. See Algorithm~\ref{matTransposeCyclic} for the pseudocode.
   \begin{algorithm}
@@
         \State $int\ \textit{i} \gets \frac{n}{size}$
-        \State $int\ \textit{j} \gets n \mod N$
+        \State $int\ \textit{j} \gets n \mod size$
         \State $T[n] \gets M[size * j + i]$
@@
 \item \textit{matTransposeIntrinsicCyclic}: same as \textit{matTransposeIntrinsic} but the matrix is saved as a one dimensional array similarly to \textit{matTransposeCyclic}
-%\item matTranspose vectorization:
+\item \textit{matTransposeVectorization}: similar to \textit{matTranspose} but the outer loop is decorated with $\# pragma\ simd $
 %\item matTranspose half vectorization:
@@
 
+\iffalse
 \item \textit{matTransposeHalfUnrollingOuter}: same as \textit{matTransposeHalf} but the outer for loop is manually unrolled
@@
 \item \textit{matTransposeHalfUnrollingInner}: same as \textit{matTransposeHalf} but the inner for loop is manually unrolled
+\fi
 
@@
 The algorithms for \textit{checkSymm} will now be explained.
-Each algorithm take an $N \times N$ square matrix \textit{input}, a size \textit{size} and return true if the matrix is symmetrix, false otherwise.
+Each algorithm takes an $N \times N$ square matrix \textit{input}, a size \textit{size} and return \textit{true} if the matrix is symmetrix, \textit{false} otherwise.
 The analyzed algorithms are the following:
 \begin{itemize}
-\item \textit{checkSymm}: naive implementation of checkSymm.
-It Iterates over the upper half of the matrix and testing if every value is equal to the symmetric value.
-See Algorithm~\ref{checkSymm} for the pseudocode.
+\item \textit{checkSymm}: naive implementation of checkSymm. It Iterates over the upper half of the matrix and testing if every value is equal to the symmetric value.
+% See Algorithm~\ref{checkSymm} for the pseudocode.
 
+\iffalse
   \begin{algorithm}
@@
   \end{algorithm}
+\fi
 
@@
 \item \textit{checkSymmUnrollingInner}: similar to \textit{checkSymmUnrolledOuter} but the inner loop is manually unrolled instead of the outer loop.
-\item \textit{checkSymmUnrollingOuterNoBranch}: similar to \textit{checkSymmUnrollingOuter}, but 4 pairs of values are checked for equality with a single math formula.
+\item \textit{checkSymmUnrollingOuterBranchless}: similar to \textit{checkSymmUnrollingOuter}, but 4 pairs of values are checked for equality with a single math formula.
 The formula performs a XOR on 4 pairs of values and their transpose, if any of those are different than 0 then the matrix is not symmetric.
-See Arlgoritm~\ref{checkSymmUnrollingOuterNoBranch} for the pseudocode.
+See Arlgoritm~\ref{checkSymmUnrollingOuterBranchless} for the pseudocode.
 
 \begin{algorithm}
-	\caption{checkSymmUnrollingOuterNoBranch}\label{checkSymmUnrollingOuterNoBranch}
+	\caption{checkSymmUnrollingOuterBranchless}\label{checkSymmUnrollingOuterBranchless}
 	\begin{algorithmic}[1]
 		\State $\textbf{bool}\ isSymm \gets true$
-		\For{$int\ i = 0; i < size; i += 4$}
-		\For{$int\ j = i; j < size; j++$}
-		\State $isSymm \gets isSymm \mid (((input[i][j] \oplus input[j][i]) \mid (input[i+1][j] \oplus input[j][i+1]) \mid (input[i+2][j] \oplus input[j][i+2]) \mid (input[i+3][j] \oplus input[j][i+3])) != 0)$
+		\For{$int\ i = 0;\ i < size;\ i += 4$}
+		\For{$int\ j = i;\ j < size;\ j++$}
+		\State $isSymm \gets isSymm\ \& $
+		\State $(((input[i][j] \oplus input[j][i]) \mid $
+		\State $(input[i+1][j] \oplus input[j][i+1]) \mid $
+		\State $(input[i+2][j] \oplus input[j][i+2]) \mid $
+		\State $(input[i+3][j] \oplus input[j][i+3])) == 0)$
 		\EndFor
@@
 
-\item \textit{checkSymmUnrollingInnerNoBranch}: similar to \textit{checkSymmUnrollingOuterNoBranch} but the inner loop is manually unrolled.
 \item \textit{checkSymOmpX}: similar to \textit{matTransposeOmpX} but based on \textit{checkSymm}. After each parallel section, a barrier is needed to synchronize the shared boolean output variable.
@@
 
-\iffalse
-Remember: write
-- what you did (contribution): high level overview like "hey, i want to test different vectorization ond parallelization implementations"
-- how you did it (methodology): we benchmarked algorithm A ... (all the pseudocodes). Whay we mesured
-
-\fi
-
 \section{Experiments and System Description}
 
+\iffalse
 - Detailed description of the computing system and platform. \\
@@
 or achieve the objectives \\
+\fi
 
-The experiments were conducted on a cluster equipped with 64 x86\_64 cpus, more specifically Intel(R) Xeon(R) Gold 5220 CPU @ 2.20GHz.
+Experiments were conducted on a cluster with 64 Intel(R) Xeon(R) Gold 5220 CPUs (2.20GHz) featuring 32K L1d and L1i caches, 1025K L2 cache, 25344K L3 cache (64-bit cache lines), 1 thread per core, and 4GB RAM, running Linux 3.10.
+The algorithms were benchmarked using valFuzz \cite{b24}, with benchmarks compiled using gcc-9.1.0 and flags "-fopenmp", "-fopenmp-simd".
+Four benchmark sets were compiled with different optimization levels: none, -O1, -O2 and "-O3 -Ofast -march=native".
 
-All relevant cluster info, compiler, flags, c++ standard. Test flags and data. Benchmarks metrics collected
-Example:
-System Description:
-The experiments were conducted on a system equipped with an Intel Core i9-12900K CPU (16 cores, 24 threads, base clock 3.2 GHz), 32 GB of DDR5 RAM, and an NVIDIA RTX 3080 GPU (10 GB GDDR6X).
-The operating system was Ubuntu 22.04, and all code was compiled using GCC 12.2 with the -O3 optimization flag.
-Benchmarks were run in an isolated environment with minimal background processes to ensure consistency.
+To ensure statistical significance, each benchmark was executed 100 times, recording wall-clock execution time statistics: minimum, maximum, median, mean, standard deviation, Q1, and Q3.
+Cache interference was minimized by clearing it before each run by writing to an array twice the L3 cache size.
+Input matrices, consistent across algorithms, were initialized with uniformly random floats (-1 to 1) generated via a Linear Congruential Generator (LCG) \cite{b25} with parameters a=1664525, c=1013904223, m=232 \cite{b26}, and seed 1337.
 
-Experimental Setup:
-The study benchmarked three sorting algorithms: QuickSort, MergeSort, and HeapSort.
-Input datasets ranged from 1,000 to 1,000,000 elements and included random, sorted, and reverse-sorted orders.
-Each experiment was repeated 10 times, with execution time and memory usage recorded.
-Performance metrics were averaged across runs to account for variability.
-
-Mamma mia che bel framework che ho
-
-Ho runanto tutto sul pbs con questi cores bellissimi con le ottimizzazioni diverse.
-Ho collezionato i dativi in un basino e l'ho grafato con quella mmerda di pitone.
-Ho fatto i test così perchè mi piacevano.
+To expedite random number generation, values were precomputed at compile time using tenno-tl \cite{b27}. Additional tenno-tl features, such as ranges and arrays, were also employed.
 
@@
 
+\iffalse
 - Presentation of results \\
@@
 
-Mamma mia ma quanto e' veloce sto asesmbly,
-buu al compilatore che non ci azzecca per bene. 
-Anche ciclico è nammerda, invece l'unrolling
-è figo.
-In generale abbiamo capito
-cose ed il cluster è pazzesco e le differenze
-sono notevoli, evviva i soldi.
+\fi
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.45\textwidth]{"../benchmarks/plotting/images/transpose_base.pdf"}
+	\caption{matTranspose results without parallelism}
+	\label{fig:transpose_base}
+\end{figure}
+\begin{figure}[htb]
+
+	\centering
+	\includegraphics[width=0.45\textwidth]{"../benchmarks/plotting/images/symm_implicit.pdf"}
+	\caption{checkSymm results with implicit parallelism}
+	\label{fig:symm_implicit}
+\end{figure}
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.45\textwidth]{"../benchmarks/plotting/images/transpose_all.pdf"}
+	\caption{matTranspose results}
+	\label{fig:transpose_benchmarks}
+\end{figure}
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.45\textwidth]{"../benchmarks/plotting/images/symm_all.pdf"}
+	\caption{checkSymm explicit parallelism results}
+	\label{fig:symm_benchmarks}
+\end{figure}
+
+The results of the benchmarks are now presented, starting with \textit{matTranspose}.
+
+% matTranspose
+
+The roofline model of the base algorithm in Figure~\ref{fig:transpose_benchmarks}(d) indicates that the base algorithm is memory-bound, meaning improvements in memory access efficiency can significantly enhance performance.
+Among the non-parallel algorithms depicted in Figure~\ref{fig:transpose_base}, \textit{matTransposeCyclic} consistently underperforms compared to others.
+In contrast, \textit{matTransposeColumns} shows negligible difference relative to the base algorithm when identical compiler optimizations are applied.
+Notably, the performance gap among algorithms widens as more aggressive compiler optimizations are introduced.
+Interestingly, \textit{checkSymmHalf} becomes slower than the baseline when these optimizations are enabled.
+
+When evaluating implicit parallelism techniques (Figure~\ref{fig:transpose_benchmarks}(a)), the fastest algorithm for matrices smaller than $10^2$ is \textit{matTransposeIntrinsicCyclic}.
+However, for larger matrices, \textit{matTransposeIntrinsic} and \textit{matTransposeUnrollingOuter} outperform it.
+Throughout all sizes, \textit{matTransposeCyclicUnrolled} remains the slowest.
+
+Explicit parallelism results are illustrated in Figure~\ref{fig:transpose_benchmarks}(b).
+Benchmarks using a higher number of threads initially take longer to execute for matrices smaller than $10^3$.
+Beyond this range, the trend reverses, with higher-threaded configurations becoming more efficient.
+All parallel benchmarks after $10^2$ are faster than the base single-core algorithm.
+The results for \textit{matTransposeOmpXCollpase} show close similarity with the non collapsed version.
+
+Finally, an analysis of scheduling (Figure~\ref{fig:transpose_benchmarks}(c)) shows that the static scheduler is generally slower for most matrix sizes.
+However, as matrix sizes increase, all algorithms converge to similar results.
+
+% checkSymm
+
+The results for \textit{checkSymm} are now presented.
+Among techniques utilizing implicit parallelism (Figure ~\ref{fig:symm_implicit}), there is minimal speed difference between methods, even with compiler optimizations enabled.
+The sole exception is \textit{checkSymmUnrollingOuterBranchless}, which significantly outperforms the others when optimizations are applied, leveraging SIMD instructions on the CPU for maximum efficiency.
+
+Explicit parallelism results (Figure~\ref{fig:symm_benchmarks}) mirror those observed in \textit{matTranspose}, initially showing poorer performance with higher thread counts.
+This trend reverses between input sizes of $10^2$ and $10^3$ when no compiler optimizations are used.
+Interestingly, with optimizations enabled, execution time exhibits a different behavior, scaling linearly with input size.
+The results with \textit{collapse} are similar to the unoptimized explicit parallelism ones.
+
+Scheduling algorithms follow a consistent performance pattern: the static scheduler is the fastest, followed by dynamic and guided scheduling.
+When optimizations are applied, all scheduling approaches demonstrate linear scaling with input size.
 
@@
 
+\iffalse
 - Summary of the key findings and contributions
+\fi
 
-Ciao ciao
+The performance analysis highlights the importance of optimization techniques and parallelism in matrix operations.
+For \textit{matTranspose}, aggressive compiler optimizations widen the performance gap between algorithms, with implicit and explicit parallelism improving efficiency for larger matrices.
+Similarly, \textit{checkSymm} benefits significantly from SIMD-enhanced methods like \textit{checkSymmUnrollingOuterBranchless}.
+Explicit parallelism for both algorithms shows a reversal in performance trends as input sizes grow, particularly with optimizations applied.
+Overall, scheduling and compiler optimizations play a critical role in achieving scalable and efficient implementations.
+
+\section{GIT and Instructions for Reproducibility}
+
+Github link: \href{https://github.com/San7o/parallel-computing-cpp/tree/first-assignment}{https://github.com/San7o/parallel-computing-cpp/tree/first-assignment} \\
+Please read the instructions in the \textit{README.md} for building.
 
@@
 \bibitem{b23} Hanounik, Bedros. (2000). Diagonal Registers: Novel Vector Register File Design for High Performance and Multimedia Computing. 
+\bibitem{b24} https://github.com/San7o/valFuzz
+\bibitem{b25} Rotenberg, A. (1960). "A New Pseudo-Random Number Generator". Journal of the ACM. 7 (1): 75–77
+\bibitem{b26} Numerical Recipes ranqd1, Chapter 7.1, An Even Quicker Generator, Eq. 7.1.6
+parameters from Knuth and H. W. Lewis
+\bibitem{b27} https://github.com/San7o/tenno-tl
+\bibitem{b28} MARTIN F. BERMAN.  A method for Transposing a Matrix. System Developement Corporation, Santa Monica, California.
 \end{thebibliography}

### 2024-12-03 tests/mpi/broadcast_bcast.cpp
mpi: exercises

@@ -1,0 +1,34 @@
+#include <mpi.h>
+#include <cstdlib>
+#include <cstdio>
+
+int main(int argc, char **argv)
+{
+  MPI_Init(&argc, &argv);
+
+  int world_size;
+  MPI_Comm_size(MPI_COMM_WORLD, &world_size);
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+
+  int err;
+  MPI_Status status;
+  int message;
+  if (world_rank == 0)
+    {
+      message = world_rank;
+      err = MPI_Bcast(&message, 1, MPI_INT,
+		      0, MPI_COMM_WORLD);
+      if (err != MPI_SUCCESS)
+	{
+	  fprintf(stderr, "RANK: %d, Error sending message!\n", world_rank);
+	}
+      fprintf(stdout, "RANK: %d, Sent message!\n", world_rank);
+    }
+  else
+    {
+      fprintf(stdout, "RANK: %d, Received message: %d\n", world_rank, message);
+    }
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-03 tests/mpi/broadcast_manual.cpp

@@ -1,0 +1,48 @@
+#include <mpi.h>
+#include <cstdlib>
+#include <cstdio>
+
+int main(int argc, char **argv)
+{
+  MPI_Init(&argc, &argv);
+
+  int world_size;
+  MPI_Comm_size(MPI_COMM_WORLD, &world_size);
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+
+  int err;
+  MPI_Status status;
+  if (world_rank == 0)
+    {
+  for (int i = 0; i < world_size; ++i)
+    {
+      err = MPI_Send((void*) &world_rank, 1, MPI_INT,
+		     (world_rank + 1) % world_size, 0, MPI_COMM_WORLD);
+      if (err != MPI_SUCCESS)
+	{
+	  fprintf(stderr, "RANK: %d, Error sending message!\n", world_rank);
+	}
+      fprintf(stdout, "RANK: %d, Sent message!\n", world_rank);
+    }
+    }
+  else
+    {
+
+  int message;
+  // int MPI_Recv(void *buf, int count, MPI_Datatype datatype,
+  //              int source, int tag, MPI_Comm comm,
+  //              MPI_Status *status)
+  err = MPI_Recv((void*) &message, 1, MPI_INT,
+		 (world_rank - 1) % world_size, 0, MPI_COMM_WORLD,
+		 &status);
+  if (err != MPI_SUCCESS)
+    {
+      fprintf(stderr, "RANK: %d, Error receiving the message!\n", world_rank);
+      exit(1);
+    }
+  fprintf(stdout, "RANK: %d, Received message: %d\n", world_rank, message);
+    }
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-03 tests/mpi/hello.cpp

@@ -1,0 +1,21 @@
+#include <mpi.h>
+#include <cstdio>
+
+int main(void)
+{
+  MPI_Init(NULL, NULL);
+
+  int world_size;
+  MPI_Comm_size(MPI_COMM_WORLD, &world_size);
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+
+  char processor_name[MPI_MAX_PROCESSOR_NAME];
+  int name_len;
+  MPI_Get_processor_name(processor_name, &name_len);
+  printf("Hello world from processor %s, rank %d out of %d processors\n", processor_name, world_rank, world_size);
+
+  MPI_Finalize();
+  
+  return 0;
+}

### 2024-12-03 tests/mpi/min.cpp

@@ -1,0 +1,28 @@
+/*
+ * Write a program that calculates the minimum value across
+ * all ranks using MPI_Allreduce. Each rank prints the results.
+ */
+
+#include <mpi.h>
+#include <stdio.h>
+
+int main(int argc, char** argv)
+{
+  MPI_Init(&argc, &argv);
+
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+  int val = world_rank + 10;
+
+  // int MPI_Allreduce(const void *sendbuf, void *recvbuf, int count,
+  //                   MPI_Datatype datatype, MPI_Op op, MPI_Comm comm)
+  int min;
+  int err = MPI_Allreduce(&val, &min, 1, MPI_INT, MPI_MIN, MPI_COMM_WORLD);
+  if (err != MPI_SUCCESS)
+    {
+      fprintf(stderr, "Error allreduce\n");
+    }
+  fprintf(stdout, "RANK: %d, Minimum value: %d\n", world_rank, min); 
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-03 tests/mpi/reduce.cpp

@@ -1,0 +1,29 @@
+#include <mpi.h>
+#include <stdio.h>
+
+int main(int argc, char** argv)
+{
+  MPI_Init(&argc, &argv);
+
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+
+  int res;
+  int tot;
+  // int MPI_Reduce(const void *sendbuf, void *recvbuf, int count,
+  //               MPI_Datatype datatype, MPI_Op op, int root,
+  //               MPI_Comm comm)
+  res = MPI_Reduce(&world_rank, &tot, 1, MPI_INT, MPI_SUM,
+		   0, MPI_COMM_WORLD);
+  if (res != MPI_SUCCESS)
+    {
+      fprintf(stderr, "Error reducing!\n");
+    }
+  if (world_rank == 0)
+    {
+      fprintf(stdout, "Tot rank: %d\n", tot);
+    }
+
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-03 tests/mpi/round-robin.cpp

@@ -1,0 +1,42 @@
+#include <mpi.h>
+#include <cstdlib>
+#include <cstdio>
+
+int main(int argc, char **argv)
+{
+  MPI_Init(&argc, &argv);
+
+  int world_size;
+  MPI_Comm_size(MPI_COMM_WORLD, &world_size);
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+
+  int err;
+  MPI_Status status;
+  // int MPI_Send(void *buf, int count, MPI_Datatype datatype,
+  //              int destination, int tag, MPI_Comm comm)
+  err = MPI_Send((void*) &world_rank, 1, MPI_INT,
+		 (world_rank + 1) % world_size, 0, MPI_COMM_WORLD);
+  if (err != MPI_SUCCESS)
+    {
+      fprintf(stderr, "RANK: %d, Error sending message!\n", world_rank);
+    }
+  fprintf(stdout, "RANK: %d, Sent message!\n", world_rank);
+
+  int message;
+  // int MPI_Recv(void *buf, int count, MPI_Datatype datatype,
+  //              int source, int tag, MPI_Comm comm,
+  //              MPI_Status *status)
+  err = MPI_Recv((void*) &message, 1, MPI_INT,
+		 (world_rank - 1) % world_size, 0, MPI_COMM_WORLD,
+		 &status);
+  if (err != MPI_SUCCESS)
+    {
+      fprintf(stderr, "RANK: %d, Error receiving the message!\n", world_rank);
+      exit(1);
+    }
+  fprintf(stdout, "RANK: %d, Received message: %d\n", world_rank, message);
+
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-03 tests/mpi/sum.cpp

@@ -1,0 +1,45 @@
+/*
+ * Write a program that calculates the minimum value across
+ * all ranks using MPI_Allreduce. Each rank prints the results.
+ */
+
+#include <mpi.h>
+#include <stdio.h>
+#include <stdlib.h>  /* exit */
+
+int main(int argc, char** argv)
+{
+  MPI_Init(&argc, &argv);
+
+  int world_size;
+  MPI_Comm_size(MPI_COMM_WORLD, &world_size);
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+  fprintf(stdout, "World size: %d\n", world_size);
+  int val = world_rank + 10;
+
+  int buff[] = { 0, 1, 2 };
+  for (int i = 0; i < 3; ++i)
+    buff[i] = buff[i] * (world_rank + 1);
+  int rec[3 * world_size] = {};
+
+  // int MPI_Alltoall(const void *sendbuf, int sendcount,
+  //	   MPI_Datatype sendtype, void *recvbuf, int recvcount,
+  //       MPI_Datatype recvtype, MPI_Comm comm)
+  int err = MPI_Alltoall(buff, 3, MPI_INT,
+			 rec, 3, MPI_INT, MPI_COMM_WORLD);
+  if (err != MPI_SUCCESS)
+    {
+      fprintf(stdout, "Error Alltoall\n");
+      MPI_Finalize();
+      exit(1);
+    }
+
+  fprintf(stdout, "RANK: %d, Output: \n", world_rank);
+  for (int j = 0; j < 3; ++j)
+    fprintf(stdout, "RANK %d, %d ", world_rank, rec[world_rank*3+j]);
+  fprintf(stdout, "\n");
+  
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-09 benchmarks/benchmarks.cpp
project: mpi workers

@@ -30,6 +30,10 @@
 #include <tenno/random.hpp>
 #include <valfuzz/valfuzz.hpp>
 
+#include <iostream>
+#include <cstdlib>    /* exit */
+#include <omp.h>
+
 #define PC_MATRIX_MAX_SIZE 1<<12
 #define PC_RANDOM_MATRIX_SIZE 256
 
@@ -103,12 +107,23 @@ BEFORE()
   matrix_in = matrix_alloc(PC_MATRIX_MAX_SIZE);
   matrix_init(matrix_in, PC_MATRIX_MAX_SIZE);
   matrix_out = matrix_alloc(PC_MATRIX_MAX_SIZE);
+
+  MPI_Init();
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+  if (world_rank != 0)
+    {
+      std::cerr << "Error initializing: Rank is not 0" << endl;
+      std::exit(1);
+    }
 }
 /* Execute this after all benchmarks */
 AFTER()
 {
   matrix_free(matrix_in, PC_MATRIX_MAX_SIZE);
   matrix_free(matrix_out, PC_MATRIX_MAX_SIZE);
+
+  MPI_Finalize();
 }
 
 

### 2024-12-09 include/pc/transpose.hpp

@@ -91,4 +91,6 @@ void matTransposeOmp16SchedStatic(float **M, float **T, tenno::size N);
 void matTransposeOmp16SchedDynamic(float **M, float **T, tenno::size N);
 void matTransposeOmp16SchedGuided(float **M, float **T, tenno::size N);
 
+void matTransposeMPI(float **M, float **T, tenno::size N);
+
 } // namespace pc

### 2024-12-09 src/transpose.cpp

@@ -17,6 +17,7 @@
 #include <omp.h>
 #include <immintrin.h>         /* For AVX intrinsics */
 
+
 /*============================================*\
 |                   BASELINE                   |
 \*============================================*/
@@ -396,3 +397,21 @@ void pc::matTransposeOmp16SchedGuided(float **M, float **T, tenno::size N)
   return;
 }
 
+
+/*============================================*\
+|                     MPI                      |
+\*============================================*/
+
+
+void matTransposeMPI(float **M, float **T, tenno::size N)
+{
+  char message[10] = "Ciaone\0";
+  int err = MPI_Bcast(&message, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
+  if (err != MPI_SUCCESS)
+    return;
+  
+  for (tenno::size i = 0; i < N; ++i)
+      for (tenno::size j = 0; j < N; ++j)
+            T[i][j] = M[j][i];
+  return;
+}

### 2024-12-09 src/workers.cpp

@@ -1,0 +1,30 @@
+#include <mpi.h>
+#include <unistd.h>
+#include <stdio.h>
+
+/* World rank of this process */
+/* should never be 0          */
+int world_rank;
+
+int main(int argc, char** argv)
+{
+  MPI_Init(&argc, &argv);
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+  if (world_rank == 0)
+    {
+      fprintf(stderr, "Error: worker should never be rank 0\n");
+      exit(1);
+    }
+
+  char buff[10] = {};
+  int err = MPI_Recv(&buff, 10, MPI_CHAR, 0, 0, MPI_COMM_WORLD, NULL);
+  if (err != MPI_SUCCESS)
+    {
+      fprintf(stderr, "Error: MPI_Recv");
+      exit(1);
+    }
+  fprintf(stdout, "WORKER %d: Received Message: %s", world_rank, buff);
+  
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-09 benchmarks/benchmarks.cpp
second-assignment: setup build system

@@ -33,6 +33,7 @@
 #include <iostream>
 #include <cstdlib>    /* exit */
 #include <omp.h>
+#include <mpi.h>
 
 #define PC_MATRIX_MAX_SIZE 1<<12
 #define PC_RANDOM_MATRIX_SIZE 256
@@ -108,15 +109,17 @@ BEFORE()
   matrix_init(matrix_in, PC_MATRIX_MAX_SIZE);
   matrix_out = matrix_alloc(PC_MATRIX_MAX_SIZE);
 
-  MPI_Init();
+  MPI_Init(NULL, NULL);
   int world_rank;
   MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
   if (world_rank != 0)
     {
-      std::cerr << "Error initializing: Rank is not 0" << endl;
+      std::cerr << "Error initializing: Rank is not 0" << std::endl;
+      MPI_Finalize();
       std::exit(1);
     }
 }
+
 /* Execute this after all benchmarks */
 AFTER()
 {
@@ -130,6 +133,8 @@ AFTER()
 /*============================================*\
 |                   TRANSPOSE                  |
 \*============================================*/
+
+// Sequencial
 
 BENCHMARK(transpose_benchmark,
 	  "matTranspose base")
@@ -215,185 +220,15 @@ BENCHMARK(transpose_4x4_intrinsic_benchmark,
     }
 }
 
-BENCHMARK(transpose_vectorization_benchmark,
-	  "matTranspose vectorization")
+// MPI
+
+BENCHMARK(transpose_mpi,
+	  "matTranspose MPI")
 {
     for (size_t N = 2; N <= 12; ++N)
     {
       RUN_BENCHMARK((1<<N),
-		    pc::matTransposeVectorization(matrix_in, matrix_out, (1<<N)));
-    }
-}
-
-BENCHMARK(transpose_unrolling_outer_benchmark,
-	  "matTranspose unrolling outer")
-{
-    for (size_t N = 2; N <= 12; ++N)
-    {
-      RUN_BENCHMARK((1<<N),
-		    pc::matTransposeOmp2(matrix_in, matrix_out, (1<<N)));
-    }
-}
-
[…]

### 2024-12-09 include/pc/check_symm.hpp

@@ -40,48 +40,4 @@ bool checkSym(float **M, tenno::size N);
 bool checkSymColumns(float **M, tenno::size N);
 
 
-/*============================================*\
-|              IMPLICIT PARALLELISM            |
-\*============================================*/
-
-// Vectorization (pragma simd)       DONE
-// Loop Unrolling inner (manual)     DONE
-// Loop Unrolling outer (manual)     DONE
-// #pragma GCC ivdep                 ASSUMED IN SIMD
-// branchless inner                  DONE
-// branchless outer                  DONE
-
-bool checkSymVectorization(float **M, tenno::size N);
-bool checkSymUnrollingInner(float **M, tenno::size N);
-bool checkSymUnrollingOuter(float **M, tenno::size N);
-bool checkSymUnrollingOuterNoBranch(float **M, tenno::size N);
-bool checkSymUnrollingInnerNoBranch(float **M, tenno::size N);
-
-
-/*============================================*\
-|              EXPLICIT PARALLELISM            |
-\*============================================*/
-
-// omp                     DONE
-// omp collapse            DONE
-// different schedulers    DONE
-
-bool checkSymOmp2(float **M, tenno::size N);
-bool checkSymOmp4(float **M, tenno::size N);
-bool checkSymOmp8(float **M, tenno::size N);
-bool checkSymOmp16(float **M, tenno::size N);
-bool checkSymOmp32(float **M, tenno::size N);
-bool checkSymOmp64(float **M, tenno::size N);
-
-bool checkSymOmp2Collapse(float **M, tenno::size N);
-bool checkSymOmp4Collapse(float **M, tenno::size N);
-bool checkSymOmp8Collapse(float **M, tenno::size N);
-bool checkSymOmp16Collapse(float **M, tenno::size N);
-bool checkSymOmp32Collapse(float **M, tenno::size N);
-bool checkSymOmp64Collapse(float **M, tenno::size N);
-
-bool checkSymOmp16SchedStatic(float **M, tenno::size N);
-bool checkSymOmp16SchedDynamic(float **M, tenno::size N);
-bool checkSymOmp16SchedGuided(float **M, tenno::size N);
-
 } // pc

### 2024-12-09 include/pc/transpose.hpp

@@ -45,51 +45,8 @@ void matTransposeIntrinsicCyclic(float *mat_in, float *mat_out, size_t N);
 
 
 /*============================================*\
-|              IMPLICIT PARALLELISM            |
+|                     MPI                      |
 \*============================================*/
-
-// Vectorization (pragma simd)       DONE
-// Loop Unrolling inner (manual)     DONE
-// Loop Unrolling outer (manual)     DONE
-// #pragma GCC ivdep                 ASSUMED IN SIMD
-// branchless inner                  DONE
-// branchless outer                  DONE
-
-void matTransposeVectorization(float **M, float **T, tenno::size N);
-void matTransposeUnrolledInner(float **M, float **T, tenno::size N);
-void matTransposeUnrolledOuter(float **M, float **T, tenno::size N);
-
-void matTransposeHalfVectorization(float **M, float **T, tenno::size N);
-void matTransposeHalfUnrolledInner(float **M, float **T, tenno::size N);
-void matTransposeHalfUnrolledOuter(float **M, float **T, tenno::size N);
-void matTransposeCyclicUnrolled(float *M, float *T, tenno::size N);
-
-
-/*============================================*\
-|              EXPLICIT PARALLELISM            |
-\*============================================*/
-
-// omp                     DONE
-// omp collapse            DONE
-// different schedulers    DONE
-
-void matTransposeOmp2(float **M, float **T, tenno::size N);
-void matTransposeOmp4(float **M, float **T, tenno::size N);
-void matTransposeOmp8(float **M, float **T, tenno::size N);
-void matTransposeOmp16(float **M, float **T, tenno::size N);
-void matTransposeOmp32(float **M, float **T, tenno::size N);
-void matTransposeOmp64(float **M, float **T, tenno::size N);
-
-void matTransposeOmp2Collapse(float **M, float **T, tenno::size N);
-void matTransposeOmp4Collapse(float **M, float **T, tenno::size N);
-void matTransposeOmp8Collapse(float **M, float **T, tenno::size N);
-void matTransposeOmp16Collapse(float **M, float **T, tenno::size N);
-void matTransposeOmp32Collapse(float **M, float **T, tenno::size N);
-void matTransposeOmp64Collapse(float **M, float **T, tenno::size N);
-
-void matTransposeOmp16SchedStatic(float **M, float **T, tenno::size N);
-void matTransposeOmp16SchedDynamic(float **M, float **T, tenno::size N);
[…]

### 2024-12-09 run-benchmarks.sh

@@ -32,39 +32,41 @@ fi
 if [ -d $BUILD_DIR ]; then
     if [ -f "$BUILD_DIR/tests" ]; then
         echo "Running regular benchmarks..."
-        ./$BUILD_DIR/tests \
-                --benchmark \
-                --num-iterations $NUM_ITERATIONS \
-                --no-multithread \
-                --report $OUTPUT_DIR/base.txt \
-                --reporter csv
+        mpirun -np 1 ./$BUILD_DIR/tests \
+                        --benchmark \
+                        --num-iterations $NUM_ITERATIONS \
+                        --no-multithread \
+                        --report $OUTPUT_DIR/base.txt \
+                        --reporter csv \
+			--run-one-benchmark "matTranspose MPI" \
+		: -np 1 ./build/worker
     fi
     if [ -f "$BUILD_DIR/tests_opt_o1" ]; then
         echo "Running optimized benchmarks..."
-        ./$BUILD_DIR/tests_opt_o1 \
-                --benchmark \
-                --num-iterations $NUM_ITERATIONS \
-                --no-multithread \
-                --report $OUTPUT_DIR/opt_o1.txt \
-                --reporter csv
+        mpirun -np 1 ./$BUILD_DIR/tests_opt_o1 \
+                        --benchmark \
+                        --num-iterations $NUM_ITERATIONS \
+                        --no-multithread \
+                        --report $OUTPUT_DIR/opt_o1.txt \
+                        --reporter csv
     fi
     if [ -f "$BUILD_DIR/tests_opt_o2" ]; then
         echo "Running optimized benchmarks..."
-        ./$BUILD_DIR/tests_opt_o2 \
-                --benchmark \
-                --num-iterations $NUM_ITERATIONS \
-                --no-multithread \
-                --report $OUTPUT_DIR/opt_o2.txt \
-                --reporter csv
+        mpirun -np 1 ./$BUILD_DIR/tests_opt_o2 \
+                        --benchmark \
+                        --num-iterations $NUM_ITERATIONS \
+                        --no-multithread \
+                        --report $OUTPUT_DIR/opt_o2.txt \
+                        --reporter csv
     fi
     if [ -f "$BUILD_DIR/tests_opt_o3" ]; then
         echo "Running aggressive optimized benchmarks..."
-        ./$BUILD_DIR/tests_opt_o3 \
-                --benchmark \
-                --num-iterations $NUM_ITERATIONS \
-                --no-multithread \
[…]

### 2024-12-09 src/check_symm.cpp

@@ -40,397 +40,4 @@ bool pc::checkSymColumns(float **M, tenno::size N)
   return symm;
 }
 
-/*============================================*\
-|              IMPLICIT PARALLELISM            |
-\*============================================*/
-
-bool pc::checkSymVectorization(float **M, tenno::size N)
-{
-  bool symm = true;
-  #pragma omp simd /* Already assumes ivdep */
-  for (size_t i = 0; i < N; ++i)
-    for (size_t j = i; j < N; ++j)
-      if (M[i][j] != M[j][i])
-	symm = false;
-  return symm;
-}
-
-bool pc::checkSymUnrollingOuter(float **M, tenno::size N)
-{
-  bool symm = true;
-  for (size_t i = 0; i < N - 4; i += 4)
-    for (size_t j = i; j < N; ++j)
-      {
-	if (M[i][j] != M[j][i]) symm = false;
-	if (M[i+1][j] != M[j][i+1]) symm = false;
-	if (M[i+2][j] != M[j][i+2]) symm = false;
-	if (M[i+3][j] != M[j][i+3]) symm = false;
-      }
-  return symm;
-}
-
-bool pc::checkSymUnrollingInner(float **M, tenno::size N)
-{
-  bool symm = true;
-  for (size_t i = 0; i < N; ++i)
-    for (size_t j = i; j < N - 4; j += 4)
-      {
-	if (M[i][j] != M[j][i]) symm = false;
-	if (M[i][j+1] != M[j+1][i]) symm = false;
-	if (M[i][j+2] != M[j+2][i]) symm = false;
-	if (M[i][j+3] != M[j+3][i]) symm = false;
-      }
-  return symm;
-}
-
-bool pc::checkSymUnrollingOuterNoBranch(float **M, tenno::size N)
-{
-  bool symm = true;
-  for (size_t i = 0; i < N - 4; i += 4)
-    for (size_t j = i; j < N; ++j)
-      {
-	symm &=
-	        ((
-		  ((uint32_t) M[i][j] - (uint32_t) M[j][i])
-	        | ((uint32_t) M[i+1][j] - (uint32_t) M[j][i+1])
-                | ((uint32_t) M[i+2][j] - (uint32_t) M[j][i+2])
-                | ((uint32_t) M[i+3][j] - (uint32_t) M[j][i+3]))
-         	== 0); // normalize the results
-      }
-  return symm;
-}
-
[…]

### 2024-12-09 src/transpose.cpp

@@ -16,6 +16,7 @@
 #include <tenno/ranges.hpp>
 #include <omp.h>
 #include <immintrin.h>         /* For AVX intrinsics */
+#include <mpi.h>
 
 
 /*============================================*\
@@ -144,266 +145,13 @@ void pc::matTransposeIntrinsic(float **mat_in, float **mat_out, size_t N)
     }
 }
 
-/*============================================*\
-|              IMPLICIT PARALLELISM            |
-\*============================================*/
-
-void pc::matTransposeVectorization(float **M, float **T, tenno::size N)
-{
-  #pragma omp simd
-  for (tenno::size i = 0; i < N; ++i)
-      for (tenno::size j = 0; j < N; ++j)
-            T[i][j] = M[j][i];
-  return;
-}
-
-void pc::matTransposeUnrolledInner(float **M, float **T, tenno::size N)
-{
-  for (tenno::size i = 0; i < N; ++i)
-      for (tenno::size j = 0; j < N; j += 4)
-	{
-            T[i][j] = M[j][i];
-            T[i][j+1] = M[j+1][i];
-            T[i][j+2] = M[j+2][i];
-            T[i][j+3] = M[j+3][i];
-	}
-  return;
-}
-
-void pc::matTransposeUnrolledOuter(float **M, float **T, tenno::size N)
-{
-  for (tenno::size i = 0; i < N; i += 4)
-      for (tenno::size j = 0; j < N; ++j)
-	{
-            T[i][j] = M[j][i];
-            T[i+1][j] = M[j][i+1];
-            T[i+2][j] = M[j][i+2];
-            T[i+3][j] = M[j][i+3];
-	}
-  return;
-}
-
-/* half */
-
-void pc::matTransposeHalfVectorization(float **M, float **T, tenno::size N)
-{
-  #pragma omp simd
-  for (tenno::size i = 0; i < N; ++i)
-      for (tenno::size j = i; j < N; ++j)
-        {
-            T[i][j] = M[j][i];
-            T[j][i] = M[i][j];
-        }
-  return;
-}
-
-void pc::matTransposeHalfUnrolledInner(float **M, float **T, tenno::size N)
-{
-  for (tenno::size i = 0; i < N; ++i)
-      for (tenno::size j = i; j < N; j += 4)
-        {
-            T[i][j] = M[j][i];
-            T[i][j+1] = M[j+1][i];
[…]

### 2024-12-09 src/workers.cpp

@@ -1,6 +1,7 @@
 #include <mpi.h>
 #include <unistd.h>
 #include <stdio.h>
+#include <cstdlib>
 
 /* World rank of this process */
 /* should never be 0          */
@@ -13,17 +14,23 @@ int main(int argc, char** argv)
   if (world_rank == 0)
     {
       fprintf(stderr, "Error: worker should never be rank 0\n");
+      MPI_Finalize();
       exit(1);
     }
 
+  fprintf(stdout, "WORKER %d is listening...\n", world_rank);
+  
   char buff[10] = {};
   int err = MPI_Recv(&buff, 10, MPI_CHAR, 0, 0, MPI_COMM_WORLD, NULL);
   if (err != MPI_SUCCESS)
     {
       fprintf(stderr, "Error: MPI_Recv");
+      MPI_Finalize();
       exit(1);
     }
   fprintf(stdout, "WORKER %d: Received Message: %s", world_rank, buff);
+
+  // TODO: Receive until STOP message
   
   MPI_Finalize();
   return 0;

### 2024-12-09 src/workers.cpp
src: worker: worker listening

@@ -21,14 +21,14 @@ int main(int argc, char** argv)
   fprintf(stdout, "WORKER %d is listening...\n", world_rank);
   
   char buff[10] = {};
-  int err = MPI_Recv(&buff, 10, MPI_CHAR, 0, 0, MPI_COMM_WORLD, NULL);
+  int err = MPI_Bcast(&buff, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
   if (err != MPI_SUCCESS)
     {
       fprintf(stderr, "Error: MPI_Recv");
       MPI_Finalize();
       exit(1);
     }
-  fprintf(stdout, "WORKER %d: Received Message: %s", world_rank, buff);
+  fprintf(stdout, "WORKER %d: Received Message: %s\n", world_rank, buff);
 
   // TODO: Receive until STOP message
   

### 2024-12-09 benchmarks/benchmarks.cpp
ompi: added include directories to worker targets

@@ -32,7 +32,6 @@
 
 #include <iostream>
 #include <cstdlib>    /* exit */
-#include <omp.h>
 #include <mpi.h>
 
 #define PC_MATRIX_MAX_SIZE 1<<12

### 2024-12-09 include/pc/transpose.hpp

@@ -27,6 +27,7 @@
 #pragma once
 
 #include <tenno/types.hpp>
+#include <omp.h>
 
 namespace pc
 {
@@ -47,6 +48,22 @@ void matTransposeIntrinsicCyclic(float *mat_in, float *mat_out, size_t N);
 /*============================================*\
 |                     MPI                      |
 \*============================================*/
+
+// Blocks Rows            TODO
+// Blocks Columns         TODO
+// Blocks Blocks          TODO
+// Cyclic ColumnWise      TODO
+// Cyclic RowWise         TODO
+
+// NOTE: Use Datatypes. You need this because scatter is so weak and stupid
+// that only sends contiguous data. We can use those datatypes to send
+// data nicely. GG
+// MPI_Datatype
+// MPI_Type_vector / contiguous / subarray (multi dimensional array)
+// MPI_Type_commit
+// MPI_Scatter
+// MPI_Type_free
+// NOTE: Handle cases where the blocks are too large at the end
 
 void matTransposeMPI(float **M, float **T, tenno::size N);
 

### 2024-12-09 src/check_symm.cpp

@@ -1,6 +1,5 @@
 #include <pc/check_symm.hpp>
 #include <cstdint>
-#include <omp.h>
 
 /*============================================*\
 |                     NOTES                    |

### 2024-12-09 src/transpose.cpp

@@ -14,7 +14,6 @@
 
 #include <pc/transpose.hpp>
 #include <tenno/ranges.hpp>
-#include <omp.h>
 #include <immintrin.h>         /* For AVX intrinsics */
 #include <mpi.h>
 
@@ -149,7 +148,6 @@ void pc::matTransposeIntrinsic(float **mat_in, float **mat_out, size_t N)
 /*============================================*\
 |                     MPI                      |
 \*============================================*/
-
 
 void pc::matTransposeMPI(float **M, float **T, tenno::size N)
 {

### 2024-12-10 latex/paper.tex

@@
 
-\title{Matrix Transposition and Symmetry in Superscalar Architectures: A Case Study\\
+\title{Matrix Transposition and Symmetry: a case study on Superscalar Architectures\\
 %{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 In section 5, we present the results. The conclusion is in section 6. Section 7 contains
-the project's github link.
+the project's github link
 
@@
 work focuses on in-situ transporision, that is transposing a matrix without
-relying on additional space. Remarcable contributions include Boothroyd's ACM Algorithm 302 (1966) using cyclic mapping \cite{b13}, Laflin and Brefner ACM Algorithm 380
+relying on additional space. Implementations were often based on the cyclic
+structure of the transposition mapping: values of a matrix of size $N$ can be stored
+in a single contiguous array or size $N$*$N$ and indexed with $row*\#rows+ column$.
+Remarcable contributions include Boothroyd's ACM Algorithm 302 (1966) using cyclic mapping \cite{b13}, Laflin and Brefner ACM Algorithm 380
 adding fixed points arithmetic \cite{b14} and
@@
 \fi 
-In constrast, while symmetric matrces are actively used in various areas of research as their
+While symmetric matrces are actively used in various areas of research as their
 properties allow for several optimizations in various problems, an in depth case study of symmetry
@@
 
-In this report, different implementations of \textit{matTranspose}
+In this report different implementations of \textit{matTranspose}
 and \textit{checkSymm} are benchmarked, comparing implicit and explicit parallelism
-techniques. In particular, the benchmarks aim to indicate cache and SIMD performance,
-speedup with multiple threads and the impact of compile flags. Data will be compared visually via graphs. All benchmarks
-will be tested against every matrix size of power of $2$ from $2^2$ to $2^{12}$.  \\
+techniques. In particular, the benchmarks aim to indicate cache and SIMD performance differences,
+speedup with multiple threads and the impact of compile flags optimizations. Data will be compared visually via graphs. All benchmarks
+will be tested against every matrix size from $2^2$ to $2^{12}$, we will assume the size value is a multiple of 4 for ease of implementation.  \\
 The algorithms for \textit{matTranspose} will now be explained.
@@
 \begin{itemize}
-\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix by rows and columns and storing them in the output matrix with indexes swapped.
+\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix and storing them in the output matrix with indexes of rows and columns swapped.
 See Algorithm~\ref{matTranspose} for the pseudocode.
@@
 
-\item \textit{matTransposeCyclic}: implementation of \cite{b28}, storing the matrix as a contiguous one-dimensional array of size $N*N$. See Algorithm~\ref{matTransposeCyclic} for the pseudocode.
+\item \textit{matTransposeCyclic}: implementation of \cite{b13}, storing the matrix as a contiguous one-dimensional array of size $N*N$. See Algorithm~\ref{matTransposeCyclic} for the pseudocode.
   \begin{algorithm}
@@
         \State $int\ \textit{i} \gets \frac{n}{size}$
-        \State $int\ \textit{j} \gets n \mod size$
+        \State $int\ \textit{j} \gets n \mod N$
         \State $T[n] \gets M[size * j + i]$
@@
 The algorithms for \textit{checkSymm} will now be explained.
-Each algorithm takes an $N \times N$ square matrix \textit{input}, a size \textit{size} and return \textit{true} if the matrix is symmetrix, \textit{false} otherwise.
+Each algorithm take an $N \times N$ square matrix \textit{input}, a size \textit{size} and return true if the matrix is symmetrix, false otherwise.
 The analyzed algorithms are the following:
@@
 
+\iffalse
+Remember: write
+- what you did (contribution): high level overview like "hey, i want to test different vectorization ond parallelization implementations"
+- how you did it (methodology): we benchmarked algorithm A ... (all the pseudocodes). Whay we mesured
+
+\fi
+
 \section{Experiments and System Description}
@@
 or achieve the objectives \\
+\fi
+
+\iffalse
+The experiments were conducted on a cluster equipped with 64 Intel(R) Xeon(R) Gold 6140M CPU with 2.30GHz, 32K of L1d cache, 32K L1i cache, 1025K L2 cache, 25344K L3 cache with cache lines of 64 bits, 1 thread per core, 4GB of DDR4 RAM running Linux 3.10.
+The algorithms were benchmarked using valFuzz \cite{b24}.
+The set of benchmarks were compiled using gcc-9.1.0 with flags "-fopenmp", "-fopenmp-simd", 4 set of benchmarks are compiled with either no optimizations, -O1, -O2 and "-O3 -Ofast -march=native".
+To obtain results with statistical significance, each benchmark has been executed 100 times collecting the following data concerning wall-clock execution time: minimum, maximum, median, arithmetic mean, standard deviation, Q1, Q3.
+To prevent interference between the benchmarks, the cache is cleared by initializing an array twice as big as the L3 cache size.
+To ensure fair comparison, the same input matrix is provided to all algorithms.
+The input matrix is initialized with an uniform random distribution of floats ranging from -1 to 1 generated using a Linear Congruential Generator (LGC) \cite{b25} with the parameters $a=1664525$, $c = 1013904223$, $m=2^{32}$ \cite{b26} and seed $1337$.
+To save time, generation of random numbers is done at compile time using tenno-tl \cite{b27}. Additional algorithms and data structures were used from tenno-tl including ranges and arrays.
 \fi
@@
 
+Mamma mia ma quanto e' veloce sto asesmbly,
+buu al compilatore che non ci azzecca per bene. 
+Anche ciclico è nammerda, invece l'unrolling
+è figo.
+In generale abbiamo capito
+cose ed il cluster è pazzesco e le differenze
+sono notevoli, evviva i soldi.
+
+Ok fatto un po' meglio:
+
+\fi
+
+\iffalse
+The results are now presented, matTranspose will be analyzed first.
+The roofline model in figure~\ref{fig:transpose_benchmarks} (d) shows that the base algorithm is memory bounded so improvement over accessing memory efficiently will be rewarded.
+Starting from the algorithms without any parallelism depicted in figure~\ref{fig:transpose_base}, we notice \textit{matTransposeCyclic} to be constantly
+slower that the others, while \textit{matTransposeColumns} shows little to no difference compared to the base algorithm with the same compiler flags.
+The performance difference between the other algorithms increases as more aggressive compiler optimizations are enabled.
+\textit{checkSymmHalf} becomes slower than the base line with the same optimizations, when they are enabled.
+Taking into consideration implicit parallelism techniques in figure~\ref{fig:transpose_benchmarks} (a), the faster algorithm is \textit{matTransposeIntrinsicCyclic} from size smaller than $10^2$, getting surpassed by \textit{matTransposeIntrinsic} and \textit{matTransposeUnrollingOuter} in larger matrices.
+\textit{matTransposeCyclicUnrolled} is the slower for all sizes.
+Looking at explicit parallelism in figure-\ref{fig:transpose_benchmarks} (b), benchmarks with giher number of threads take more time than benchmarks with fewer threads until around size $10^2$ to $10^3$ where the trend is inverted.
+When analysing schedulers, static is slower for most matrix sixzes but all algorithms converge to about the same timing in the bigger input.
 \fi
@@
 
-Github link: \href{https://github.com/San7o/parallel-computing-cpp/tree/first-assignment}{https://github.com/San7o/parallel-computing-cpp/tree/first-assignment} \\
+Github link: \href{https://github.com/San7o/parallel-computing-cpp}{https://github.com/San7o/parallel-computing-cpp} \\
 Please read the instructions in the \textit{README.md} for building.
@@
 \bibitem{b27} https://github.com/San7o/tenno-tl
-\bibitem{b28} MARTIN F. BERMAN.  A method for Transposing a Matrix. System Developement Corporation, Santa Monica, California.
 \end{thebibliography}

### 2024-12-10 latex/paper2.tex
mpi: setup

\iffalse

References draft:

Jaeyoung Choi, Jack J. Dongarra, David W. Walker.  Parallel Matrix Transpose Algorithms on Distributed Memory Concurrent Computers.  Mathematical Sciences Section Oak Ridge National Laboratory P.O. Box 2008, Bldg. 6012 Oak Ridge, TN 37831-6367

Notes: The communication is based of the greatest commond divisor (GCD) of the number of rows and columns (P and Q). If P and Q are relatively prima, the matrix transpose algorithm involves complete exchange communication. A block is LCM/P x LCM/Q if not relativaly prime and only LCM/GCD communications are required.

Jonathan Eckstein, Gyorgy Matyasfalvi. Efficient Distributed-Memory Parallel Matrix-Vector Multiplication with Wide or Tall Unstructured Sparse Matrices.

John C. Bowman, Malcolm Roberts. Adaptive Matrix Transpose Algorithms for Distributed Multicore Processors.

Notes: "The most popular algorithms for transposing an NxN matrix distributed over P processes are the direct communicatoin (all-to-all, simpler implementation) and recursive binary exchange (butterfly) algorithms."

Jaeyoung Choi, Jack Dongarra, David W. Walker. The design of scalable software libraries for distributed memory concurrent computers. Oak Ridge National Laboartory Mathematical Sciences Section. Oak RIdge, TN 37831-6367.
- they explained (apparently) block cyclic data distribution

So we learned that block cyclic data distribution (round robin) is efficient for data access. You can mix blocks with cycles to get a few combinations to try out.

\fi

### 2024-12-10 benchmarks/benchmarks.cpp

@@ -8,8 +8,7 @@
  * in the Software without restriction, including without limitation the rights
  * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
  * copies of the Software, and to permit persons to whom the Software is
- * furnished to do so, subject to the following conditions:
- *
+ * furnished to do so, subject to the following conditions: *
  * The above copyright notice and this permission notice shall be included in
  all
  * copies or substantial portions of the Software.
@@ -25,14 +24,15 @@
  */
 
 #include <pc/transpose.hpp>
+#include <pc/benchmarks.hpp>
 #include <pc/check_symm.hpp>
+#include <pc/mpi.hpp>
 #include <tenno/ranges.hpp>
 #include <tenno/random.hpp>
 #include <valfuzz/valfuzz.hpp>
 
 #include <iostream>
 #include <cstdlib>    /* exit */
-#include <mpi.h>
 
 #define PC_MATRIX_MAX_SIZE 1<<12
 #define PC_RANDOM_MATRIX_SIZE 256
@@ -42,12 +42,13 @@
 |                     SETUP                    |
 \*============================================*/
 
-typedef float** matrix;
-
 /* Using a global matrix so that it will be
  * initialized only once */
-matrix matrix_in;
-matrix matrix_out;
+pc::matrix pc::matrix_in;
+pc::matrix pc::matrix_out;
+
+/* The MPI world rank */
+int pc::world_rank;
 
 /**
  * Generating two random arrays at compile time for
@@ -68,9 +69,9 @@ constexpr tenno::array<float, PC_RANDOM_MATRIX_SIZE> random_arr2()
     return tenno::random_array<PC_RANDOM_MATRIX_SIZE>(seed2, min, max);
 }
 
-matrix matrix_alloc(tenno::size N)
+pc::matrix matrix_alloc(tenno::size N)
 {
-  matrix M = new float*[N];
+  pc::matrix M = new float*[N];
   for (unsigned int i = 0; i < N; ++i)
     {
       M[i] = new float[N];
[…]

### 2024-12-10 include/pc/benchmarks.hpp

@@ -1,0 +1,38 @@
+/*
+ * MIT License
+ *
+ * Copyright (c) 2024 Giovanni Santini
+ *
+ * Permission is hereby granted, free of charge, to any person obtaining a copy
+ * of this software and associated documentation files (the "Software"), to deal
+ * in the Software without restriction, including without limitation the rights
+ * to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+ * copies of the Software, and to permit persons to whom the Software is
+ * furnished to do so, subject to the following conditions:
+ *
+ * The above copyright notice and this permission notice shall be included in
+ all
+ * copies or substantial portions of the Software.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+ * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+ * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+ * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+ * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+ * OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+ * SOFTWARE.
+ *
+ */
+
+#pragma once
+
+namespace pc
+{
+
+typedef float** matrix;
+
+extern int world_rank;
+extern matrix matrix_in;
+extern matrix matrix_out;
+
+} // namespace pc

### 2024-12-10 include/pc/transpose.hpp

@@ -64,7 +64,15 @@ void matTransposeIntrinsicCyclic(float *mat_in, float *mat_out, size_t N);
 // MPI_Scatter
 // MPI_Type_free
 // NOTE: Handle cases where the blocks are too large at the end
+// NOTE: Use MPI_Scatter to distribute pieces of an array between
+// different processes and MPI_Gather to get the results.
+// In this last function, only the root must have a receive buffer, all
+// the other can pass NULL. recv_count is the count of element received per process.
 
-void matTransposeMPI(float **M, float **T, tenno::size N);
+// Example naive approach:
+// use MPI_Type_contiguous for rows, scatter, reverse order of all rows, gather,
+// use MPI_Type_vector for columns, scatter, reverse order of all columns, gather.
+  
+void matTransposeMPIInvert2(float **M, float **T, tenno::size N);
 
 } // namespace pc

### 2024-12-10 run-debug.sh

@@ -22,51 +22,31 @@
 # OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
 # SOFTWARE.
 
-NUM_ITERATIONS=10
 BUILD_DIR="build"
-OUTPUT_DIR=benchmarks/plotting/reports
 
 if [ ! -d $OUTPUT_DIR ]; then
     mkdir $OUTPUT_DIR
 fi
 if [ -d $BUILD_DIR ]; then
     if [ -f "$BUILD_DIR/tests" ]; then
-        echo "Running regular benchmarks..."
+        echo "Running regular tests..."
         mpirun -np 1 ./$BUILD_DIR/tests \
-                        --benchmark \
-                        --num-iterations $NUM_ITERATIONS \
-                        --no-multithread \
-                        --report $OUTPUT_DIR/base.txt \
-                        --reporter csv \
-			--run-one-benchmark "matTransposeMPIInvert2" \
-		: -np 1 ./build/worker
+		: -np 1 gdb ./build/worker
     fi
     if [ -f "$BUILD_DIR/tests_opt_o1" ]; then
-        echo "Running optimized benchmarks..."
+        echo "Running tests o1..."
         mpirun -np 1 ./$BUILD_DIR/tests_opt_o1 \
-                        --benchmark \
-                        --num-iterations $NUM_ITERATIONS \
-                        --no-multithread \
-                        --report $OUTPUT_DIR/opt_o1.txt \
-                        --reporter csv
+		: -np 1 gdb ./build/worker
     fi
     if [ -f "$BUILD_DIR/tests_opt_o2" ]; then
-        echo "Running optimized benchmarks..."
+        echo "Running tests o2..."
         mpirun -np 1 ./$BUILD_DIR/tests_opt_o2 \
-                        --benchmark \
-                        --num-iterations $NUM_ITERATIONS \
-                        --no-multithread \
-                        --report $OUTPUT_DIR/opt_o2.txt \
-                        --reporter csv
+		: -np 1 gdb ./build/worker
     fi
     if [ -f "$BUILD_DIR/tests_opt_o3" ]; then
-        echo "Running aggressive optimized benchmarks..."
+        echo "Running tests o3..."
         mpirun -np 1 ./$BUILD_DIR/tests_opt_o3 \
-                        --benchmark \
-                        --num-iterations $NUM_ITERATIONS \
-                        --no-multithread \
[…]

### 2024-12-10 src/transpose.cpp

@@ -13,9 +13,10 @@
  */ 
 
 #include <pc/transpose.hpp>
+#include <pc/benchmarks.hpp>
+#include <pc/mpi.hpp>
 #include <tenno/ranges.hpp>
 #include <immintrin.h>         /* For AVX intrinsics */
-#include <mpi.h>
 
 
 /*============================================*\
@@ -149,15 +150,33 @@ void pc::matTransposeIntrinsic(float **mat_in, float **mat_out, size_t N)
 |                     MPI                      |
 \*============================================*/
 
-void pc::matTransposeMPI(float **M, float **T, tenno::size N)
+void pc::matTransposeMPIInvert2(float **M, float **T, tenno::size N)
 {
-  char message[10] = "Ciaone\0";
-  int err = MPI_Bcast(&message, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
-  if (err != MPI_SUCCESS)
+  (void) T;
+  (void) M;
+
+  /* Create a new datatype */
+  MPI_Datatype row_t;
+  int err = mpi::Type_vector((int) N, (int) N, 0, MPI_FLOAT, &row_t);
+  if (err != mpi::SUCCESS)
     return;
-  
-  for (tenno::size i = 0; i < N; ++i)
-      for (tenno::size j = 0; j < N; ++j)
-            T[i][j] = M[j][i];
+
+  /* Scatter */
+  float *row = new float[N*N];
+  /*
+  err = mpi::Scatter(M, (int) N, row_t, row, 1, row_t, 0, MPI_COMM_WORLD);
+  if (err != mpi::SUCCESS)
+    return;
+  */
+
+  /* Gather */
+
+  /* Create a new datatype */
+
+  /* Scatter */
+
+  /* Gather */
+
+  delete[] row;
   return;
 }

### 2024-12-10 src/workers.cpp

@@ -1,7 +1,9 @@
-#include <mpi.h>
+#include <pc/transpose.hpp>
+#include <pc/mpi.hpp>
 #include <unistd.h>
 #include <stdio.h>
 #include <cstdlib>
+#include <cstring>
 
 /* World rank of this process */
 /* should never be 0          */
@@ -9,28 +11,53 @@ int world_rank;
 
 int main(int argc, char** argv)
 {
-  MPI_Init(&argc, &argv);
-  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+  mpi::Init(&argc, &argv);
+  mpi::Comm_rank(MPI_COMM_WORLD, &world_rank);
   if (world_rank == 0)
     {
       fprintf(stderr, "Error: worker should never be rank 0\n");
-      MPI_Finalize();
+      mpi::Finalize();
       exit(1);
     }
 
   fprintf(stdout, "WORKER %d is listening...\n", world_rank);
   
-  char buff[10] = {};
-  int err = MPI_Bcast(&buff, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
-  if (err != MPI_SUCCESS)
+  char func[10] = {};
+  int err = mpi::Bcast(&func, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
+  if (err != mpi::SUCCESS)
     {
       fprintf(stderr, "Error: MPI_Recv");
-      MPI_Finalize();
+      mpi::Finalize();
       exit(1);
     }
-  fprintf(stdout, "WORKER %d: Received Message: %s\n", world_rank, buff);
+  fprintf(stdout, "WORKER %d: Function: %s\n", world_rank, func);
 
-  // TODO: Receive until STOP message
+  int N = {};
+  do {
+    err = mpi::Bcast(&N, 1, MPI_INT, 0, MPI_COMM_WORLD); /* get N */
+    if (err != mpi::SUCCESS)
+    {
+	fprintf(stderr, "Error: MPI_Bcast");
+	mpi::Finalize();
+	exit(1);
+    }
+
+    fprintf(stdout, "WORKER %d: Function: %s\n", world_rank, func);
+
+    fprintf(stdout, "WORKER %d: N: %d\n", world_rank, N);
+    fprintf(stdout, "WORKER %d: Is \"%s\" equal to \"Invert2\" ??\n", world_rank, func);
+    if (strcmp(func, "Invert2") == 0)
+	{
[…]

### 2024-12-10 tests/mpi/matTransposeInvert2.cpp

@@ -1,0 +1,87 @@
+#include <pc/mpi.hpp>
+#include <iostream>
+
+int world_rank;
+
+/* Executed by root */
+void matTransposeMPIInvert2(float **M, float **T, size_t N)
+{
+
+  /* Create a new datatype */
+  mpi::Datatype row_t;
+  int err = mpi::Type_vector(N, N, 0, MPI_FLOAT, &row_t);
+  if (err != mpi::SUCCESS)
+    return;
+
+  /* MPI Scatter */
+  float row[N];
+  err = mpi::Scatter(&M, N, row_t, &row, 1, row_t, 0, MPI_COMM_WORLD);
+  if (err != mpi::SUCCESS)
+    return;
+
+  /* MPI Gather */
+  
+  return;
+}
+
+/* Executed by workers */
+void matTransposeMPIInvert2Worker(float **M, float **T, size_t N)
+{
+  /* Create a new datatype */
+  mpi::Datatype row_t;
+  int err = mpi::Type_vector(N, N, 0, MPI_FLOAT, &row_t);
+  if (err != mpi::SUCCESS)
+    return;
+
+  float row[N];
+  err = mpi::Scatter(NULL, N, row_t, &row, 1, row_t, 0, MPI_COMM_WORLD);
+  if (err != mpi::SUCCESS)
+    return;
+
+  std::cout << "WORKER " << world_rank << " Received: " << world_rank;
+  for (int i = 0; i < N; ++i)
+    std::cout << row[i] << " ";
+  std::cout << std::endl;
+}
+
+int main(void)
+{
+  mpi::Init(NULL, NULL);
+
+  mpi::Comm_rank(MPI_COMM_WORLD, &world_rank);
+
+  constexpr size_t N = 16;
+  float** mat1;
+  float** mat2;
+  if (world_rank == 0)
+    {
+	/* Initialize matices */
+	mat1 = new float* [N];
+	mat2 = new float* [N];
+	for (size_t i = 0; i < N; ++i)
+	    {
+	    mat1[i] = new float[N];
+	    mat2[i] = new float[N];
+	    for (int j = 0; j < N; ++j)
+	      mat1[i][j] = float(j * i);
+	    }
+
+	matTransposeMPIInvert2(mat1, mat2, N);
+
+	/* Cleanup */
+	for (size_t i = 0; i < N; ++i)
+	{
+	    delete[] mat1[i];
+	    delete[] mat2[i];
+	}
[…]

### 2024-12-10 tests/transpose_test.cpp

@@ -25,10 +25,12 @@
  */
 
 #include <pc/transpose.hpp>
+#include <pc/mpi.hpp>
+#include <pc/benchmarks.hpp>  /* contains definition of matrices and world_rank */
 #include <tenno/ranges.hpp>
 #include <valfuzz/valfuzz.hpp>
 
-TEST(transpose_matrix_test, "Matrix Transpose")
+TEST(transpose_matrix_test, "matTranspose")
 {
     tenno::size N = 10;
     float **M = new float *[N];
@@ -57,7 +59,7 @@ TEST(transpose_matrix_test, "Matrix Transpose")
     }
 }
 
-TEST(transpose_matrix_half_test, "Matrix Transpose Half")
+TEST(transpose_matrix_half_test, "matTransposeHalf")
 {
     tenno::size N = 10;
     float **M = new float *[N];
@@ -82,6 +84,47 @@ TEST(transpose_matrix_half_test, "Matrix Transpose Half")
         for (auto j : tenno::range(N))
         {
             ASSERT(M[i][j] == T[j][i]);
+        }
+    }
+}
+
+TEST(transpose_matrix_mpi_invert2, "matTransposeMPIInvert2")
+{
+    if (pc::world_rank != 0)
+      ASSERT(false);
+
+    constexpr tenno::size N = 10;
+    float **M = new float *[N];
+    float **T = new float *[N];
+    for (auto i : tenno::range(N))
+    {
+        M[i] = new float[N];
+	T[i] = new float[N];
+        for (auto j : tenno::range(N))
+        {
+	  M[i][j] = float(i * N + j);
+        }
+    }
+
+    /* Message the workers */
+    char message[10] = "Invert2\0";
+    int err = MPI_Bcast(&message, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
+    if (err != MPI_SUCCESS)
+      return;
+
+    int n = N; /* -fpermissive gets angry */
+    err = mpi::Bcast(&n, 10, MPI_INT, 0, MPI_COMM_WORLD);
+    if (err != MPI_SUCCESS)
+      return;
+
+    pc::matTransposeMPIInvert2(M, T, N);
+
+    for (auto i : tenno::range(N))
+    {
+        for (auto j : tenno::range(N))
+        {
+            ASSERT(M[i][j] == T[j][i]);
+	    return;
         }
     }
 }

### 2024-12-11 benchmarks/benchmarks.cpp
mpi: working test environment

@@ -47,8 +47,9 @@
 pc::matrix pc::matrix_in;
 pc::matrix pc::matrix_out;
 
-/* The MPI world rank */
+/* The MPI world rank and world_size */
 int pc::world_rank;
+int pc::world_size;
 
 /**
  * Generating two random arrays at compile time for
@@ -117,6 +118,7 @@ BEFORE()
       MPI_Finalize();
       std::exit(1);
     }
+  mpi::Comm_size(MPI_COMM_WORLD, &pc::world_size);
 }
 
 /* Execute this after all benchmarks */
@@ -235,17 +237,20 @@ BENCHMARK(transpose_mpi,
 
     for (size_t N = 2; N <= 12; ++N)
     {
-      err = mpi::Bcast(&N, 1, MPI_INT, 0, MPI_COMM_WORLD);
+      err = mpi::Bcast(&N, 1, MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD);
       if (err != mpi::SUCCESS)
         return;
 
+      /*
       RUN_BENCHMARK((1<<N),
 		    pc::matTransposeMPIInvert2(pc::matrix_in, pc::matrix_out, (1<<N)));
+      */
 
+      break; /* TODO remove */
     }
 
-    int fin = -1;
-    err = mpi::Bcast(&fin, 1, MPI_INT, 0, MPI_COMM_WORLD);
+    size_t fin = 0;  /* terminate */
+    err = mpi::Bcast(&fin, 1, MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD);
     if (err != mpi::SUCCESS)
       return;
 }

### 2024-12-11 include/pc/benchmarks.hpp

@@ -32,6 +32,7 @@ namespace pc
 typedef float** matrix;
 
 extern int world_rank;
+extern int world_size;
 extern matrix matrix_in;
 extern matrix matrix_out;
 

### 2024-12-11 include/pc/mpi.hpp

@@ -1108,7 +1108,8 @@ inline int MPI_Comm_set_attr(MPI_Comm comm, int comm_keyval, void *attribute_val
 inline int MPI_Comm_set_errhandler(MPI_Comm comm, MPI_Errhandler errhandler);
 inline int MPI_Comm_set_info(MPI_Comm comm, MPI_Info info);
 inline int MPI_Comm_set_name(MPI_Comm comm, const char *comm_name);
-inline int MPI_Comm_size(MPI_Comm comm, int *size);
+inline int Comm_size(MPI_Comm comm, int *size)
+{ return MPI_Comm_size(comm, size); }
 inline int MPI_Comm_spawn(const char *command, char *argv[], int maxprocs, MPI_Info info,
                                   int root, MPI_Comm comm, MPI_Comm *intercomm,
                                   int array_of_errcodes[]);
@@ -1235,9 +1236,11 @@ inline int MPI_File_sync(MPI_File fh);
 inline int Finalize(void) { return MPI_Finalize(); }
 inline int MPI_Finalized(int *flag);
 inline int MPI_Free_mem(void *base);
-inline int MPI_Gather(const void *sendbuf, int sendcount, MPI_Datatype sendtype,
+inline int Gather(const void *sendbuf, int sendcount, MPI_Datatype sendtype,
                               void *recvbuf, int recvcount, MPI_Datatype recvtype,
-                              int root, MPI_Comm comm);
+                              int root, MPI_Comm comm)
+{ return MPI_Gather(sendbuf, sendcount, sendtype, recvbuf, recvcount, recvtype,
+		    root, comm); }
 inline int MPI_Igather(const void *sendbuf, int sendcount, MPI_Datatype sendtype,
                                void *recvbuf, int recvcount, MPI_Datatype recvtype,
                                int root, MPI_Comm comm, MPI_Request *request);
@@ -1585,9 +1588,11 @@ inline int MPI_Testsome(int incount, MPI_Request array_of_requests[],
                                 MPI_Status array_of_statuses[]);
 inline int MPI_Topo_test(MPI_Comm comm, int *status);
 inline MPI_Fint MPI_Type_c2f(MPI_Datatype datatype);
-inline int MPI_Type_commit(MPI_Datatype *type);
-inline int MPI_Type_contiguous(int count, MPI_Datatype oldtype,
-                                       MPI_Datatype *newtype);
+inline int Type_commit(Datatype *type)
[…]

### 2024-12-11 include/pc/transpose.hpp

@@ -73,6 +73,6 @@ void matTransposeIntrinsicCyclic(float *mat_in, float *mat_out, size_t N);
 // use MPI_Type_contiguous for rows, scatter, reverse order of all rows, gather,
 // use MPI_Type_vector for columns, scatter, reverse order of all columns, gather.
   
-void matTransposeMPIInvert2(float **M, float **T, tenno::size N);
+void matTransposeMPIInvert2(float *M, float *T, tenno::size N);
 
 } // namespace pc

### 2024-12-11 run-debug.sh

@@ -31,22 +31,22 @@ if [ -d $BUILD_DIR ]; then
     if [ -f "$BUILD_DIR/tests" ]; then
         echo "Running regular tests..."
         mpirun -np 1 ./$BUILD_DIR/tests \
-		: -np 1 gdb ./build/worker
+		: -np 1 xterm -e gdb ./build/worker
     fi
     if [ -f "$BUILD_DIR/tests_opt_o1" ]; then
         echo "Running tests o1..."
         mpirun -np 1 ./$BUILD_DIR/tests_opt_o1 \
-		: -np 1 gdb ./build/worker
+		: -np 1 xterm -e gdb ./build/worker
     fi
     if [ -f "$BUILD_DIR/tests_opt_o2" ]; then
         echo "Running tests o2..."
         mpirun -np 1 ./$BUILD_DIR/tests_opt_o2 \
-		: -np 1 gdb ./build/worker
+		: -np 1 xterm -e gdb ./build/worker
     fi
     if [ -f "$BUILD_DIR/tests_opt_o3" ]; then
         echo "Running tests o3..."
         mpirun -np 1 ./$BUILD_DIR/tests_opt_o3 \
-		: -np 1 gdb ./build/worker
+		: -np 1 xterm -e gdb ./build/worker
     fi
 else
     echo "Please run the build script first"

### 2024-12-11 src/transpose.cpp

@@ -17,6 +17,8 @@
 #include <pc/mpi.hpp>
 #include <tenno/ranges.hpp>
 #include <immintrin.h>         /* For AVX intrinsics */
+#include <algorithm>
+#include <cstdio> /* TODO remove */
 
 
 /*============================================*\
@@ -145,33 +147,98 @@ void pc::matTransposeIntrinsic(float **mat_in, float **mat_out, size_t N)
 |                     MPI                      |
 \*============================================*/
 
-void pc::matTransposeMPIInvert2(float **M, float **T, tenno::size N)
+void pc::matTransposeMPIInvert2(float *M, float *T, tenno::size N)
 {
   (void) T;
-  (void) M;
 
-  /* Create a new datatype */
-  MPI_Datatype row_t;
-  int err = mpi::Type_vector((int) N, (int) N, 0, MPI_FLOAT, &row_t);
-  if (err != mpi::SUCCESS)
-    return;
-
-  /* Scatter */
-  float *row = new float[N*N];
   /*
-  err = mpi::Scatter(M, (int) N, row_t, row, 1, row_t, 0, MPI_COMM_WORLD);
-  if (err != mpi::SUCCESS)
-    return;
+  // DEBUG
+  if (pc::world_rank == 0)
+    {
+      fprintf(stdout, "First row: \n");
+      for (size_t i = 0; i < N; ++i)
+	fprintf(stdout, "%f ", M[i]);
+      fprintf(stdout, "\n");
+    }
   */
 
+  /* Send rows */
+  MPI_Datatype row_t;
+  int err = mpi::Type_contiguous((int) N, MPI_FLOAT, &row_t);
+  if (err != mpi::SUCCESS)
+    return;
+  err = mpi::Type_commit(&row_t);
+  if (err != mpi::SUCCESS)
+    return;
+
+  float *row = new float[N * N / world_size];
+  err = mpi::Scatter(M, (int) (N / world_size), row_t,   /* send */
+		     row, (int) (N / world_size), row_t, /* recv */
+		     0, MPI_COMM_WORLD);
+  if (err != mpi::SUCCESS)
+    return;
+
+  /*
+  // DEBUG
+  fprintf(stdout, "WORKER %d: Got row: ", world_rank);
+  for (size_t i = 0; i < N * N / world_size; ++i)
+    fprintf(stdout, "%f ", row[i]);
[…]

### 2024-12-11 src/workers.cpp

@@ -7,20 +7,27 @@
 
 /* World rank of this process */
 /* should never be 0          */
+namespace pc
+{
+
 int world_rank;
+int world_size;
+
+} // namespace pc
 
 int main(int argc, char** argv)
 {
   mpi::Init(&argc, &argv);
-  mpi::Comm_rank(MPI_COMM_WORLD, &world_rank);
-  if (world_rank == 0)
+  mpi::Comm_rank(MPI_COMM_WORLD, &pc::world_rank);
+  if (pc::world_rank == 0)
     {
       fprintf(stderr, "Error: worker should never be rank 0\n");
       mpi::Finalize();
       exit(1);
     }
+  mpi::Comm_size(MPI_COMM_WORLD, &pc::world_size);
 
-  fprintf(stdout, "WORKER %d is listening...\n", world_rank);
+  fprintf(stdout, "WORKER %d is listening...\n", pc::world_rank);
   
   char func[10] = {};
   int err = mpi::Bcast(&func, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
@@ -30,11 +37,10 @@ int main(int argc, char** argv)
       mpi::Finalize();
       exit(1);
     }
-  fprintf(stdout, "WORKER %d: Function: %s\n", world_rank, func);
 
-  int N = {};
-  do {
-    err = mpi::Bcast(&N, 1, MPI_INT, 0, MPI_COMM_WORLD); /* get N */
+  size_t N = 0;
+  while(true) {
+    err = mpi::Bcast(&N, 1, MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD); /* get N */
     if (err != mpi::SUCCESS)
     {
 	fprintf(stderr, "Error: MPI_Bcast");
@@ -42,21 +48,23 @@ int main(int argc, char** argv)
 	exit(1);
     }
 
-    fprintf(stdout, "WORKER %d: Function: %s\n", world_rank, func);
+    if (N == 0)
+      break;
 
-    fprintf(stdout, "WORKER %d: N: %d\n", world_rank, N);
-    fprintf(stdout, "WORKER %d: Is \"%s\" equal to \"Invert2\" ??\n", world_rank, func);
+    fprintf(stdout, "WORKER %d: Function: %s\n", pc::world_rank, func);
+    fprintf(stdout, "WORKER %d: N: %ld\n", pc::world_rank, N);
+
+
     if (strcmp(func, "Invert2") == 0)
-	{
[…]

### 2024-12-11 tests/mpi/bcast.cpp

@@ -1,0 +1,27 @@
+#include <mpi.h>
+#include <cstdlib>
+#include <cstdio>
+
+int main(int argc, char **argv)
+{
+  MPI_Init(&argc, &argv);
+
+  int world_size;
+  MPI_Comm_size(MPI_COMM_WORLD, &world_size);
+  int world_rank;
+  MPI_Comm_rank(MPI_COMM_WORLD, &world_rank);
+
+  int err;
+  size_t message;
+  if (world_rank == 0)
+    message = 777;
+  err = MPI_Bcast(&message, 1, MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD);
+  if (err != MPI_SUCCESS)
+    {
+      fprintf(stderr, "RANK: %d, Error sending message\n", world_rank);
+    }
+  fprintf(stdout, "RANK: %d, Message %d\n", world_rank, message);
+  
+  MPI_Finalize();
+  return 0;
+}

### 2024-12-11 tests/transpose_test.cpp

@@ -93,18 +93,21 @@ TEST(transpose_matrix_mpi_invert2, "matTransposeMPIInvert2")
     if (pc::world_rank != 0)
       ASSERT(false);
 
-    constexpr tenno::size N = 10;
-    float **M = new float *[N];
-    float **T = new float *[N];
-    for (auto i : tenno::range(N))
-    {
-        M[i] = new float[N];
-	T[i] = new float[N];
-        for (auto j : tenno::range(N))
-        {
-	  M[i][j] = float(i * N + j);
-        }
-    }
+    constexpr tenno::size N = (1<<2);
+    /* stack allocated matrix */
+    float M_cyclic[N*N];
+    float T_cyclic[N*N] = {};
+    for (size_t i = 0; i < N*N; ++i)
+	M_cyclic[i] = float(i);
+
+    fprintf(stdout, "Original: \n");
+    for (size_t i = 0; i < N*N; ++i)
+      {
+	if (i % N == 0 && i != 0)
+	  fprintf(stdout, "\n");
+	fprintf(stdout, "%f ", M_cyclic[i]);
+      }
+    fprintf(stdout, "\n");
 
     /* Message the workers */
     char message[10] = "Invert2\0";
@@ -112,19 +115,34 @@ TEST(transpose_matrix_mpi_invert2, "matTransposeMPIInvert2")
     if (err != MPI_SUCCESS)
       return;
 
-    int n = N; /* -fpermissive gets angry */
-    err = mpi::Bcast(&n, 10, MPI_INT, 0, MPI_COMM_WORLD);
+    size_t n = N; /* -fpermissive gets angry */
+    err = mpi::Bcast(&n, 1, MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD);
     if (err != MPI_SUCCESS)
       return;
 
-    pc::matTransposeMPIInvert2(M, T, N);
+    pc::matTransposeMPIInvert2(M_cyclic, T_cyclic, N);
+
+    size_t fin = 0;
+    err = mpi::Bcast(&fin, 1, MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD);
+    if (err != MPI_SUCCESS)
+      return;
+
+    fprintf(stdout, "Transposed: \n");
+    for (size_t i = 0; i < N*N; ++i)
+      {
+	if (i % N == 0 && i != 0)
+	  fprintf(stdout, "\n");
+	fprintf(stdout, "%f ", T_cyclic[i]);
[…]

### 2024-12-11 include/pc/mpi.hpp
workflow: attempting to make ubuntu work: added proposed repository (attempt 3)

@@ -29,6 +29,7 @@
 namespace mpi
 {
 
+
 /*
  * Typedefs
  */

### 2025-01-02 latex/paper2.tex
matTranspose: non blocking implementation

@@
 6012 Oak Ridge, TN 37831-6367
-
-Notes: The communication is based of the greatest commond divisor (GCD) of the number of rows and columns (P and Q).
+- Notes: The communication is based of the greatest commond divisor (GCD) of the number of rows and columns (P and Q).
 If P and Q are relatively prima, the matrix transpose algorithm involves complete exchange communication.
 A block is LCM/P x LCM/Q if not relativaly prime and only LCM/GCD communications are required.
+- The algorithms make use of non-blocking, point-to-point communication between proces-
+sors. The use of nonblocking communication allows a processor to overlap the messages that it sends to different processors, thereby avoiding unnecessary synchronization.
 
 Jonathan Eckstein, Gyorgy Matyasfalvi. Efficient Distributed-Memory Parallel Matrix-Vector Multiplication with Wide or Tall Unstructured Sparse Matrices.
+- on sparse matrixes, but Its about vector multiplications
 
 John C. Bowman, Malcolm Roberts. Adaptive Matrix Transpose Algorithms for Distributed Multicore Processors.
-
-Notes: "The most popular algorithms for transposing an NxN matrix distributed over P processes are the direct communicatoin (all-to-all, simpler implementation) and recursive binary exchange (butterfly) algorithms."
+- really nice for my problem
+- hybrid approach with both shared and distributed
+- Notes: "The most popular algorithms for transposing an NxN matrix distributed over P processes are the direct communicatoin (all-to-all, simpler implementation) and recursive binary exchange (butterfly) algorithms."
 
@@
 Oak RIdge, TN 37831-6367.
-- they explained (apparently) block cyclic data distribution
+- they explained (apparently) block cyclic data distribution. Actually no
 

### 2025-01-02 benchmarks/benchmarks.cpp

@@ -271,6 +271,50 @@ BENCHMARK(transpose_mpi_benchmark,
     return;
 }
 
+BENCHMARK(transpose_mpi_nonblocking_benchmark,
+	  "matTransposeMPINonblocking")
+{
+    if (pc::world_rank != 0)
+      return;
+
+    float *M_cyclic = new float[PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE];
+    float *T_cyclic = new float[PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE];
+
+    constexpr auto arr1 = random_arr1();
+
+    for (size_t i = 0; i < PC_MATRIX_MAX_SIZE*PC_MATRIX_MAX_SIZE; ++i)
+      M_cyclic[i] = arr1[i % PC_RANDOM_MATRIX_SIZE];
+      
+    int err;
+    char message[10] = "Base\0";
+    long unsigned int num_iterations =
+	valfuzz::get_num_iterations_benchmark() + 2;
+    long unsigned int size;
+    for (size_t N = 2; N <= 12; ++N)
+    {
+      err = MPI_Bcast(&message, 10, MPI_CHAR, 0, MPI_COMM_WORLD);
+      if (err != MPI_SUCCESS)
+        return;
+ 
+      size = (1<<N);
+      err = MPI_Bcast(&size, 1, MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD);
+      if (err != MPI_SUCCESS)
+        return;
+
+      err = MPI_Bcast(&num_iterations, 1,
+		       MPI_UNSIGNED_LONG, 0, MPI_COMM_WORLD);
+      if (err != MPI_SUCCESS)
+        return;
+      
+      RUN_BENCHMARK((1<<N),
+      	    pc::matTransposeMPINonblocking(M_cyclic, T_cyclic, (1<<N)));
+    }
+
+    delete[] M_cyclic;
+    delete[] T_cyclic;
+    return;
+}
+
 BENCHMARK(transpose_mpi_block_benchmark,
 	  "matTransposeMPIBlock")
 {

### 2025-01-02 include/pc/transpose.hpp

@@ -50,6 +50,7 @@ void matTransposeIntrinsicCyclic(float *mat_in, float *mat_out, size_t N);
 \*============================================*/
 
 void matTransposeMPI(float *M, float *T, tenno::size N);
+void matTransposeMPINonblocking(float *M, float *T, tenno::size N);
 void matTransposeMPIBlock(float *M, float *T, tenno::size N);
 
 // Used for debugging purposes

### 2025-01-02 src/transpose.cpp

@@ -218,6 +218,78 @@ void pc::matTransposeMPI(float *M, float *T, tenno::size N)
   return;
 }
 
+void pc::matTransposeMPINonblocking(float *M, float *T, tenno::size N)
+{
+  if (world_size > (int) N) /* fallback */
+  {
+    if (world_rank == 0)
+    {
+      for (tenno::size i = 0; i < N*N; ++i)
+	T[i] = M[N*(i % N) + (i / N)];
+    }
+    return;
+  }
+
+  MPI_Datatype row_t;
+  int err = MPI_Type_contiguous((int) N,   /* count   */
+				 MPI_FLOAT, /* oldtype */
+				 &row_t);   /* newtype */
+  if (err != MPI_SUCCESS)
+    return;
+  err = MPI_Type_commit(&row_t);
+  if (err != MPI_SUCCESS)
+    return;
+
+  MPI_Datatype col_t_tmp, col_t;
+  err = MPI_Type_vector((int) N,        /* count       */
+			 1,              /* blocklength */
+			 (int) N,        /* stride      */
+			 MPI_FLOAT,      /* oldtype     */
+			 &col_t_tmp);    /* newtype     */
+  if (err != MPI_SUCCESS)
+    return;
+  err = MPI_Type_create_resized(col_t_tmp,     /* oldtype */
+				 0,             /* lb      */
+				 sizeof(float), /* extent  */
+				 &col_t);       /* newtype */
+  err = MPI_Type_commit(&col_t);
+  if (err != MPI_SUCCESS)
+    return;
+  MPI_Type_free(&col_t_tmp);
+
+  MPI_Request request;
+  float *row = new float[N * N / world_size];
+  err = MPI_Iscatter(M,                      /* sendbuf   */
+		     (int) (N / world_size), /* sendcount */
+		     row_t,                  /* sendtype  */
+		     row,                    /* recvbuf   */
+		     (int) (N / world_size), /* recvcount */
+		     row_t,                  /* recvtype  */
+		     0,                      /* root      */
+ 		     MPI_COMM_WORLD,         /* comm      */
+		     &request);              /* request   */
+  if (err != MPI_SUCCESS)
+    return;
+  MPI_Wait(&request, NULL);
+
+  err = MPI_Igather(row,                  /* sendbuf   */
+		    (int) N / world_size, /* sendcount */
+		    row_t,                /* sendtype  */
+		    T,                    /* recvbuf   */
[…]

### 2025-01-02 benchmarks/benchmarks.cpp
matTranspose: fixed workers for non blocking implementation

@@ -286,7 +286,7 @@ BENCHMARK(transpose_mpi_nonblocking_benchmark,
       M_cyclic[i] = arr1[i % PC_RANDOM_MATRIX_SIZE];
       
     int err;
-    char message[10] = "Base\0";
+    char message[10] = "NB\0";
     long unsigned int num_iterations =
 	valfuzz::get_num_iterations_benchmark() + 2;
     long unsigned int size;

### 2025-01-02 src/master.cpp

@@ -81,6 +81,11 @@ int main(int argc, char** argv)
 	pc::matTransposeMPI(mat1, mat2, N);
     }
   }
+  else if (strcmp(func, "NB") == 0)
+  {
+    for (unsigned long i = 0; i < num_iterations; ++i)
+      pc::matTransposeMPINonblocking(mat1, mat2, N);
+  }
   else if (strcmp(func, "BlockDbg") == 0)
   {
     for (unsigned long i = 0; i < num_iterations; ++i)

### 2025-01-02 src/workers.cpp

@@ -81,6 +81,11 @@ int main(int argc, char** argv)
 	pc::matTransposeMPI(mat1, mat2, N);
       }
     }
+    else if (strcmp(func, "NB") == 0)
+    {
+      for (unsigned long i = 0; i < num_iterations; ++i)
+	pc::matTransposeMPINonblocking(mat1, mat2, N);
+    }
     else if (strcmp(func, "BlockDbg") == 0)
     {
       for (unsigned long i = 0; i < num_iterations; ++i)

### 2025-01-02 src/transpose.cpp
matTranspose: fixed mpi null status

@@ -270,7 +270,7 @@ void pc::matTransposeMPINonblocking(float *M, float *T, tenno::size N)
 		     &request);              /* request   */
   if (err != MPI_SUCCESS)
     return;
-  MPI_Wait(&request, NULL);
+  MPI_Wait(&request,MPI_STATUS_IGNORE);
 
   err = MPI_Igather(row,                  /* sendbuf   */
 		    (int) N / world_size, /* sendcount */
@@ -283,7 +283,7 @@ void pc::matTransposeMPINonblocking(float *M, float *T, tenno::size N)
 		    &request);
   if (err != MPI_SUCCESS)
     return;
-  MPI_Wait(&request, NULL);
+  MPI_Wait(&request, MPI_STATUS_IGNORE);
 
   delete[] row;
   MPI_Type_free(&row_t);

### 2025-01-09 latex/template.tex
benchmarks: up to 256 processes

--- 2024-11-29 latex/paper.tex
@@
+\iffalse
+
+References draft:
+
+Jaeyoung Choi, Jack J.
+Dongarra, David W.
+Walker.
+Parallel Matrix Transpose Algorithms on Distributed Memory Concurrent Computers.
+Mathematical Sciences Section Oak Ridge National Laboratory P.O.
+Box 2008, Bldg.
+6012 Oak Ridge, TN 37831-6367
+- Notes: The communication is based of the greatest commond divisor (GCD) of the number of rows and columns (P and Q).
+If P and Q are relatively prima, the matrix transpose algorithm involves complete exchange communication.
+A block is LCM/P x LCM/Q if not relativaly prime and only LCM/GCD communications are required.
+- The algorithms make use of non-blocking, point-to-point communication between proces-
+sors. The use of nonblocking communication allows a processor to overlap the messages that it sends to different processors, thereby avoiding unnecessary synchronization.
+
+Jonathan Eckstein, Gyorgy Matyasfalvi. Efficient Distributed-Memory Parallel Matrix-Vector Multiplication with Wide or Tall Unstructured Sparse Matrices.
+- on sparse matrixes, but Its about vector multiplications
+
+John C. Bowman, Malcolm Roberts. Adaptive Matrix Transpose Algorithms for Distributed Multicore Processors.
+- really nice for my problem
+- hybrid approach with both shared and distributed
+- Notes: "The most popular algorithms for transposing an NxN matrix distributed over P processes are the direct communicatoin (all-to-all, simpler implementation) and recursive binary exchange (butterfly) algorithms."
+
+Jaeyoung Choi, Jack Dongarra, David W.
+Walker.
+The design of scalable software libraries for distributed memory concurrent computers.
+Oak Ridge National Laboartory Mathematical Sciences Section.
+Oak RIdge, TN 37831-6367.
+- they explained (apparently) block cyclic data distribution. Actually no
+
+So we learned that block cyclic data distribution (round robin) is efficient for data access. You can mix blocks with cycles to get a few combinations to try out.
+
+Sheng Ma, Libo Huang, Mingche Lai, Wei Shi. Network-on-chip customizations for message passing interface primitives: From Implementations to Programming Paradigms, 2015
+- In conventional MPI implementations, MPI_Bcast is typically implemented using software with a tree-based algorithm.
+Such implementations exploit point-to-point communication operations.
+Thus, the number of hops to reach leaf nodes increases with the total number of nodes (typically in a logarithmic manner); the latency of MPI_Bcast also increases.
+- this is useful in the findings because we find the same outcome
+- but I use scatter and gather... i should look at the implementation
+
+From the source code: coll/coll_base_scatter:
+- There are 3 implementations of scatter:
+  - binomial tree algorithm, as discussed above uses point-to-point communications
+  - linear functions which work the same for small values
+  - linear non blocking
+- The decision is taken in coll/tuned/coll_tuned_scatter_decision.c
+
+\fi
+
 \documentclass[conference]{IEEEtran}
@@
 
-\title{Matrix Transposition and Symmetry: a case study on Superscalar Architectures\\
+\title{Matrix Transposition and Symmetry on Distributed Memory Systems\\
 %{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 \begin{abstract}
-TODO *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+TODO for later
 \end{abstract}
@@
 
-Modern CPUs use superscalar architectures to improve performance, combining multiple cores with advanced Single Instruction Multiple Data (SIMD) technologies.
-Examples include Intel's Streaming SIMD Extensions (SSE) and Advanced Vector Extensions (AVX).
-Although SSE is not a new technology, being first introduced in the Pentium III in 1999 and subsequently expanded in later years \cite{b1}, modern algorithms leveraging SIMD instructions continue to be a topic of active research in recent literature.
-For example, the FreeBSD's SIMD-enhanced libc is actively reworking the standard library's code. \cite{b2}.
-In 2023, most of the string functions were reimplemented by Robert Clausecker et al on many architectures \cite{b3}, exploiting vector instructions and improving the overall performance of the standard library \cite{b4}.
-Similarly, efficient multi-threaded algorithms are becoming more critical as the demand of computation increases worldwide due to AI training \cite{b5} \cite{b6}.
-This case study examines the implementations and performance of two algorithms—out-of-place square matrix transposition (matTranspose) and square matrix symmetry checking (checkSymm)—in the context of a superscalar architecture.
-Both algorithms have been used pervasively in performance critical applications.
-For example, matrix transposition was used in NASA's Apollo 11 Guidance Computer (AGC) to convert vectors in platenary coordinate system to base reference system \cite{b7}.
-Additionally, matrix transposition is used by Fourier Transform in the West (FFTW) \cite{b8}, a popular library to calculate Fast Furier Transofrms, and in signal processing \cite{b9}.
-Furthermore, the current C++26 draft proposes basic linear algebra algorithms in the "linalg" header, including matrix transposition \cite{b10}.
+Introduction here
 
 In section 2, we discuss the state of the art of the two algorithms.
-In section 3, we elaborate on the algorithms analyzed, discuss performance evaluation and we present a new benchmarking framework.
-In section 4, we show experimental data and graphs that is analyzed
-in section 5. The conclusion is in section 6.
+In section 3, we elaborate on the algorithms analyzed.
+In section 4, we explain the benchmarking procedures and system information.
+In section 5, we present the results. The conclusion is in section 6. Section 7 contains
+the project's github link
 
@@
 
-\iffalse
-
-\textit{We define the transpose of a matrix $A$, and we denote it by $A^t$, as the matrix obtained from $A$ by interchanging rows and columns.
-Specifically, if $A$ is an $m$-by-$n$ matrix, then $A^t$ is the $n$-by-$m$ matrix whose entries are given by the equation (1)}.
-\cite{b11}
-
-\begin{equation}
-	(A^t)_{k,j} = A_{j,k}
-\end{equation}
-
-\fi
-
-Computing the transpose of a matrix is a well discussed computational problem. Early
-work focuses on in-situ transporision, that is transposing a matrix without
-relying on additional space. Implementations were often based on the cyclic
-structure of the transposition mapping: values of a matrix of size $N$ can be stored
-in a single contiguous array or size $N$*$N$ and indexed with $row*\#rows+ column$.
-Remarcable contributions include Boothroyd's ACM Algorithm 302 (1966) using cyclic mapping \cite{b13}, Laflin and Brefner ACM Algorithm 380
-adding fixed points arithmetic \cite{b14} and
-a later revision of the two by Esko G. Cate and David W. Twigg in ACM Algorithm 513 \cite{b15} .
-SIMD implementations for specific architectures are also widely adopted
-in performance critical applications like cryptographic libraries, utilizing instructions like Intel's vpunpckldq and vpunpckhdq \cite{b16} and reporting improvements up to 70\% \cite{b17}.
-However, matrix operations are not efficient as vector registers access data only linearly, motivating research on new vector register file (VFS) designs such as diagonal registers \cite{b23}.
-Another area of research includes out-of-core matrix transposition, that
-is computing the transpose of a matrix that exceeds the size of the available in-core memory.
-This problem focuses on optimizing the number of I/O operations and file layout since memory-memory data transfer is notably slower than math operations.
-Work in out-of-core transposition include W.
-0.
-Alltop (1975) \cite{b18}, Suh \& Prasanna (2002) \cite{b19} and Krishnamoorth et al.
-(2004) \cite{b20}.
-Recent work explores parallelization and the use of GPUs to offload algebra operations. Examples are I-Jui Sung et al \cite{b21} and Mark Harris \cite{b22}.
-
-\iffalse
-\textit{A square matrix $A$ is called symmetric if it equals its transpose} \cite{b12}.
-\fi 
-While symmetric matrces are actively used in various areas of research as their
-properties allow for several optimizations in various problems, an in depth case study of symmetry
-checking on modern architectures is yet to be published.
+TODO
 
@@
 
-\iffalse % pseudocode
-    \begin{algorithmic}[1]
-      \Procedure{MyProcedure}{}
-      % use \state and \textit{} for variables
-      % use \text for text
-      % use \gets for assignments
-      % use \If \EndIf \Return \Else
-      \EndProcedure
-    \end{algorithmic}
-\fi
-
-In this report different implementations of matTranspose
-and checkSymm are benchmarked, comparing implicit and explicit parallelism
-techniques. In particular, the benchmarks aim to indicate cache and SIMD performance differences,
-speedup with multiple threads and the impact of compile flags optimizations. Data will be compared through tables
-and visually via graphs such as the roofline model. All benchmarks
-will be tested against every matrix size from $2^2$ to $2^{12}$, we will assume the size value is a multiple of 4 for ease of implementation.  \\
-The algorithms for \textit{matTranspose} will now be explained.
-Each algorithm takes a $N \times N$ input square matrix \textit{input}, a matrix of the same shape \textit{ouput}, and the size of the two matrices \textit{size}.
-The analyzed algorithms are the following:
-\begin{itemize}
-\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix and storing them in the output matrix with indexes of rows and columns swapped.
-See Algorithm~\ref{matTranspose} for the pseudocode.
-
-\begin{algorithm}
-	\caption{matTranspose}\label{matTranspose}
-	\begin{algorithmic}[1]
-		\For{\textit{r} \textbf{ in } \textit{0..size-1}}
-	    \For{\textit{c} \textbf{ in } \textit{0..size-1}}
-		\State $output[c][r] \gets input[r][c]$
-		\EndFor
-		\EndFor
-	\end{algorithmic}
-\end{algorithm}
-
-    \item \textit{matTransposeColumns}: similar to \textit{matTranspose}, but iterating on the input matrix first by columns and then by rows.
-  The purpose of this implementation is to test a different access pattern.
-
-\iffalse
-\begin{algorithm}
-	\caption{matTransposeColumns}\label{matTransposeColumns}
-\begin{algorithmic}[1]
-	\For{\textit{r} \textbf{ in } \textit{0..size-1}}
-	 \For{\textit{c} \textbf{ in } \textit{0..size-1}}
-            \State $output[r][c] \gets input[c][r]$
-            \EndFor
-        \EndFor
-\end{algorithmic}
-\end{algorithm}
-\fi
-
-  \item \textit{matTransposeHalf}: similar to \textit{matTranspose}, but the algorithm
-  iterates only on the upper half of the input matrix. For each
-  pair $(i,j)$ of indices of rows and columns in the upper half of the input matrix $M$, with output matrix $O$, $O[i][j]=M[j][i]$ and $O[j][i] = M[i][j]$.
-\iffalse
-\begin{algorithm}
-	\caption{matTransposeHalf}\label{matTransposeHalf}
-\begin{algorithmic}[1]
-	\For{\textit{r} \textbf{ in } \textit{0..size-1}}
-           \For{\textit{c} \textbf{ in } \textit{r..size-1}}
-            \State $output[r][c] \gets input[c][r]$
-            \State $output[c][r] \gets input[r][c]$
-            \EndFor
-        \EndFor
-\end{algorithmic}
-\end{algorithm}
-\fi
-
-\item \textit{matTransposeCyclic}: implementation of \cite{b13}, storing the matrix as a contiguous one-dimensional array of size $N*N$. See Algorithm~\ref{matTransposeCyclic} for the pseudocode.
-  \begin{algorithm}
-  	\caption{matTransposeCyclic}\label{matTransposeCyclic}
-  \begin{algorithmic}[1]
-
-  	  \For{\textit{n} \textbf{ in } \textit{0..size*size-1}}
-        \State $int\ \textit{i} \gets \frac{n}{size}$
-        \State $int\ \textit{j} \gets n \mod N$
-        \State $T[n] \gets M[size * j + i]$
-        \EndFor
-          \end{algorithmic}
-  \end{algorithm}
-
-\item \textit{matTransposeIntrinsic}: based on \textit{matrix\_transpose4x4} from the linux kernel \cite{b16}, revised to work for out-of-place float matrices on x86 architectures and implemented using intrinsics.
-\textit{transpose\_4x4\_f32\_intrinsic} transposes a 4 by 4 matrix using vector registers of 128 bits by doing the following steps: load 4 rows of 4 floats from the the input matrix, combine pairs of rows in an alternating pattern (interleaving), shuffle to get the transpose and store them in the output matrix.
-\textit{matTransposeIntrinsic} iterates over 4x4 blocks in the input matrix, calling \textit{transpose\_4x4\_f32\_intrinsic} and saving the computed transposed block in the output matrix with rows and columns swapped.
-See Algorithm~\ref{matTransposeIntrinsic} for the pseudocode.
-  \begin{algorithm}
-	\caption{matTransposeIntrinsic}\label{matTransposeIntrinsic}
-	    \begin{algorithmic}[1]
-	\iffalse
-
-      \Procedure{transpose\_4x4\_f32\_intrinsic}{Matrix4x4 in, Matrix4x4 out}
-    \State \textbf{comment} Load rows  
-    \State \textbf{comment} unpack and interleave rows
-    \State \textbf{comment} Shuffle to form transposed rows
-    \State \textbf{comment} Store transposed rows
-      \EndProcedure
-
-      \fi
-
-        \For{$int\ i = 0, i < size, i += 4$}
-        \For{$int\ j = 0, j < size, j += 4$}
-        \State $transpose\_4x4\_f32\_intrinsic(input[i][j],$
-        \State $                               output[j][i])$
-        \EndFor
-        \EndFor
-
-    \end{algorithmic}
-    \end{algorithm}
-
-\item \textit{matTransposeIntrinsicCyclic}: same as \textit{matTransposeIntrinsic} but the matrix is saved as a one dimensional array similarly to \textit{matTransposeCyclic}
-%\item matTranspose vectorization:
-%\item matTranspose half vectorization:
-\item \textit{matTransposeUnrollingOuter}: similar to \textit{matTranspose} but the outer loop is manually unrolled. See Algorithm~\ref{matTransposeUnrollingOuter} for the pseudocode.
-
-  \begin{algorithm}
-    \caption{matTransposeUnrollingOuter}\label{matTransposeUnrollingOuter}
-    \begin{algorithmic}[1]
-        \For{$int\ c = 0, i < size, i += 4$}
-            \For{\textit{r} \textbf{ in } \textit{0..size-1}}
-            \State $output[c][r] \gets input[r][c]$
-            \State $output[c+1][r] \gets input[r][c+1]$
-            \State $output[c+2][r] \gets input[r][c+2]$
-            \State $output[c+3][r] \gets input[r][c+3]$
-            \EndFor
-        \EndFor
-    \end{algorithmic}
-  \end{algorithm}
-
-\item \textit{matTransposeUnrollingInner}: same as \textit{matTransposeUnrollingOuter} but the inner for loop is manually unrolled
-
-\item \textit{matTransposeHalfUnrollingOuter}: same as \textit{matTransposeHalf} but the outer for loop is manually unrolled
-
-\item \textit{matTransposeHalfUnrollingInner}: same as \textit{matTransposeHalf} but the inner for loop is manually unrolled
-
-\item \textit{matTransposeCyclicUnrolled}: same as \textit{matTransposeCyclic} but the for loop in manually unrolled
-\iffalse
-  \begin{algorithm}
-    \caption{matTranspose cyclic unrolled}\label{matTransposeCyclicUnrolled}
-    \begin{algorithmic}[1]
-      \Procedure{matTransposeCyclicUnrolled}{$Matrix input, Matrix output, int size$}
-        \For{\textit{n} \textbf{ in } \textit{0..size*size-1}}
-        \State $T[n] \gets M[size * ((n+1) \mod size ) + (\frac{n}{size})]$
-        \State $T[n+1] \gets M[size * ((n+1+1) \mod size ) + (\frac{n+1}{size})]$
-        \State $T[n+2] \gets M[size * ((n+2+1) \mod size ) + (\frac{n+2}{size})]$
-        \State $T[n+3] \gets M[size * ((n+3+1) \mod size ) + (\frac{n+3}{size})]$
-        \EndFor
-      \EndProcedure
-    \end{algorithmic}
-  \end{algorithm}
-  \fi
-\item \textit{matTransposeOmpX}: this set of algorithms are all derived from \textit{matTranspose} with the outer for loops parallelized using omp with the number of threads \textit{X} being a power of 2 from $2^1$ to $2^6$.
-
-\item \textit{matTransposeOmpXCollapse}: similar to \textit{matTransposeOmpX} but the for loops are parellelized using \textit{collapse(2)}
-
-\item \textit{matTransposeOmp16SchedX}: similar to \textit{matTransposeOmpX} with 16 threads using different schedule algorithms, namely: static, dynamic, guided.
-\end{itemize}
-
-The algorithms for \textit{checkSymm} will now be explained.
-Each algorithm take an $N \times N$ square matrix \textit{input}, a size \textit{size} and return true if the matrix is symmetrix, false otherwise.
-The analyzed algorithms are the following:
-\begin{itemize}
-\item \textit{checkSymm}: naive implementation of checkSymm.
-It Iterates over the upper half of the matrix and testing if every value is equal to the symmetric value.
-See Algorithm~\ref{checkSymm} for the pseudocode.
-
-  \begin{algorithm}
-    \caption{checkSymm}\label{checkSymm}
-    \begin{algorithmic}[1]
-        \State $\textbf{bool}\ isSymm \gets true$
-        \For{\textit{i} \textbf{ in } \textit{0..size-1}}
-            \For{\textit{j} \textbf{ in } \textit{i..size-1}}
-            	\If{$input[i][j] != input[j][i]$}
-            	\State $isSymm \gets \textbf{false}$
-            	\EndIf
-        	\EndFor
-        \EndFor
-        \State \Return $isSymm$
-    \end{algorithmic}
-  \end{algorithm}
-
-% \item \textit{checkSymColumns}:
-
-  % \item \textit{checkSymVectorization}
-\item \textit{checkSymmUnrolledOuter}: similar to \textit{checkSymm} but the outer loop is manually unrolled.
-
-\iffalse
-  \begin{algorithm}
-    \caption{checkSymmUnrolledOuter}\label{checkSymmUnrolledOuter}
-    \begin{algorithmic}[1]
-      \Procedure{checkSymmUnrolledOuter}{$Matrix input, int size$}
-        \State $isSymm = true$
-        \For{\textit{i} \textbf{ in } \textit{0..size-1}}
-            \For{\textit{j} \textbf{ in } \textit{i..size-1}}
-            \If $input[i][j] != input[j][i]$
-            \State $isSymm = false$
-            \EndIf
-            \If $input[i+1][j] != input[j][i+1]$
-            \State $isSymm = false$
-                        \EndIf
-            \If $input[i+2][j] != input[j][i+2]$
-            \State $isSymm = false$
-                        \EndIf
-            \If $input[i+3][j] != input[j][i+3]$
-            \State $isSymm = false$
-                        \EndIf
-        \EndFor
-        \EndFor
-        \Return $isSymm$
-      \EndProcedure
-    \end{algorithmic}
-  \end{algorithm}
-\fi
-
-\item \textit{checkSymmUnrollingInner}: similar to \textit{checkSymmUnrolledOuter} but the inner loop is manually unrolled instead of the outer loop.
-\item \textit{checkSymmUnrollingOuterNoBranch}: similar to \textit{checkSymmUnrollingOuter}, but 4 pairs of values are checked for equality with a single math formula.
-The formula performs a XOR on 4 pairs of values and their transpose, if any of those are different than 0 then the matrix is not symmetric.
-See Arlgoritm~\ref{checkSymmUnrollingOuterNoBranch} for the pseudocode.
-
-\begin{algorithm}
-	\caption{checkSymmUnrollingOuterNoBranch}\label{checkSymmUnrollingOuterNoBranch}
-	\begin{algorithmic}[1]
-		\State $\textbf{bool}\ isSymm \gets true$
-		\For{$int\ i = 0; i < size; i += 4$}
-		\For{$int\ j = i; j < size; j++$}
-		\State $isSymm \gets isSymm \mid (((input[i][j] \oplus input[j][i]) \mid (input[i+1][j] \oplus input[j][i+1]) \mid (input[i+2][j] \oplus input[j][i+2]) \mid (input[i+3][j] \oplus input[j][i+3])) != 0)$
-		\EndFor
-		\EndFor
-		\State \Return $isSymm$
-	\end{algorithmic}
-\end{algorithm}
-
-\item \textit{checkSymmUnrollingInnerNoBranch}: similar to \textit{checkSymmUnrollingOuterNoBranch} but the inner loop is manually unrolled.
-\item \textit{checkSymOmpX}: similar to \textit{matTransposeOmpX} but based on \textit{checkSymm}. After each parallel section, a barrier is needed to synchronize the shared boolean output variable.
-\item \textit{checkSymOmpXCollapse}: similar to \textit{checkSymmOmpX} bur the two loops are parallelized with collapse(2).
-\item \textit{checkSymm16SchedX}: similar to \textit{matTranspose16SchedX} but based on \textit{matTranspose}.
-\end{itemize}
-
-\iffalse
-Remember: write
-- what you did (contribution): high level overview like "hey, i want to test different vectorization ond parallelization implementations"
-- how you did it (methodology): we benchmarked algorithm A ... (all the pseudocodes). Whay we mesured
-
-\fi
+TODO
 
@@
 
-The experiments were conducted on a cluster equipped with 64 Intel(R) Xeon(R) Gold 5220 CPUs at 2.20GHz, 32K of L1d cache, 32K L1i cache, 1025K L2 cache, 25344K L3 cache with cache lines of 64 bits, 1 thread per core, 4GB of DDR4 RAM running Linux 3.10.
-The algorithms were benchmarked using valFuzz \ref{b?}.
-The set of benchmarks were compiled with flags "-fopenmp", "-fopenmp-simd", 4 set of benchmarks are compiles with either no optimizations, -O1, -O2, -O3 -Ofast -march=native.
-To obtain results with statistical significance, each benchmark has been executed 100 times collecting the following data concerning wall-clock execution time: minimum, maximum, median, arithmetic mean, standard deviation, Q1, Q3.
-To prevent interference between the benchmarks, the cache is cleared by initializing an array twice as big as the L3 cache size.
-To ensure fair comparison, the same input matrix is provided to all algorithms.
-The input matrix is initialized with an uniform random distribution of floats ranging from -1 to 1 generated using lcg algorithm \ref{b?} with the parameters a=1664525 and c = 1013904223.
-To save time, generation of random numbers is done at compile time using tenno-tl \ref{b?} with the initial seed "1337".
-Additional algorithms and data structures were used from tenno-tl including ranges and arrays.
-
-All relevant cluster info, compiler, flags, c++ standard. Test flags and data. Benchmarks metrics collected
-Example:
-System Description:
-The experiments were conducted on a system equipped with an Intel Core i9-12900K CPU (16 cores, 24 threads, base clock 3.2 GHz), 32 GB of DDR5 RAM, and an NVIDIA RTX 3080 GPU (10 GB GDDR6X).
-The operating system was Ubuntu 22.04, and all code was compiled using GCC 12.2 with the -O3 optimization flag.
-Benchmarks were run in an isolated environment with minimal background processes to ensure consistency.
-
-Experimental Setup:
-The study benchmarked three sorting algorithms: QuickSort, MergeSort, and HeapSort.
-Input datasets ranged from 1,000 to 1,000,000 elements and included random, sorted, and reverse-sorted orders.
-Each experiment was repeated 10 times, with execution time and memory usage recorded.
-Performance metrics were averaged across runs to account for variability.
-
-Mamma mia che bel framework che ho
-
-Ho runanto tutto sul pbs con questi cores bellissimi con le ottimizzazioni diverse.
-Ho collezionato i dativi in un basino e l'ho grafato con quella mmerda di pitone.
-Ho fatto i test così perchè mi piacevano.
+TODO
 
@@
 - Comparison with the state-of-the-art \\
+\fi
 
-cinquantamila grafici qui
-
-Mamma mia ma quanto e' veloce sto asesmbly,
-buu al compilatore che non ci azzecca per bene. 
-Anche ciclico è nammerda, invece l'unrolling
-è figo.
-In generale abbiamo capito
-cose ed il cluster è pazzesco e le differenze
-sono notevoli, evviva i soldi.
+TODO
 
@@
 
+\iffalse
 - Summary of the key findings and contributions
+\fi
 
-Ciao ciao
+TODO
+
+\section{GIT and Instructions for Reproducibility}
+
+Github link: \href{https://github.com/San7o/parallel-computing-cpp}{https://github.com/San7o/parallel-computing-cpp} \\
+Please read the instructions in the \textit{README.md} for building.
 
 \begin{thebibliography}{00}
-\bibitem{b1} Intel® 64 and IA-32 Architectures Software Developer’s Manual Combined Volumes: 1, 2A, 2B, 2C, 2D, 3A, 3B, 3C, 3D, and 4. Volume 1, Section 2.2.7 "SIMD Instructions"
-\bibitem{b2} SIMD man page "https://man.freebsd.org/cgi/man.cgi?simd(7)"
-\bibitem{b3} Git blame on freeBSD's source code, in "lib/libc/amd64/string", shows recent activity
-\bibitem{b4} https://freebsdfoundation.org/blog/a-sneak-peek-simd-enhanced-string-functions-for-amd64/
-\bibitem{b5} Amodei, D. Hernandez, D. AI and Compute, https://openai.com/blog/ai-and-compute
-\bibitem{b6} Andrew J. Lohn, Micah Musse. AI and Compute How Much Longer Can Computing Power Drive Artificial Intelligence Progress? CSET Issue Brief
-\bibitem{b7} Digitalized Apollo 11 guidance computer by Virtual AGC and MIT Museum.
-Code in "Apollo-11/Comanche055/PLANETERY\_INTERNAL\_ORIENTATION.agc" line 34; "Apollo-11/Luminary099/ATTITUDE\_MANEUVER\_ROUTINE.agc" line 441
-\bibitem{b8} Matteo Frigo and Steven G. Johnson. "The design and implementation of FFTW3". The word "transpose" is mentioned 21 times
-\bibitem{b9} M.
-Henriksson and O.
-Gustafsson, "Streaming Matrix Transposition on FPGAs Using Distributed Memories," 2023 IEEE Nordic Circuits and Systems Conference (NorCAS), Aalborg, Denmark, 2023, pp.
-1-6, doi: 10.1109/NorCAS58970.2023.10305472.
-\bibitem{b10} "https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2023/p1673r12.html" Accessed in november 2024
-\bibitem{b11} Sheldon Axler. Linear Algebra Done Right ourth edition
-16 November 2024. Section 3C Matrices, definition 3.54
-\bibitem{b12} Sheldon Axler. Linear Algebra Done Right ourth edition
-16 November 2024. Section 9A Bilinear Forms and Quadratic Forms, definition 9.11
-\bibitem{b13} BOOTHROYD,Z. Algorithm 302: Transpose vector stored array. Go~r~m.A~M I0, ~ (1967),
-292-293
-\bibitem{b14} LAFLIN, S., AND BREFNER, M.A. Algorithm 380: In-situ transposition of a rectangular
-matrix, Comm. ACM IS, 5 (1970), 324-326
-\bibitem{b15} ESKO G. CATE and DAVID W. TWIGG ALGORITHM 513: Analysis of In-Situ Transposition (1977)
-\bibitem{b16} most of the crypto functions are written in assembly usin SIMD instructions in the
-linux kernel "linux/arch/x86/crypto" for exanple transpose\_4x4 in "serpent-avx2-asm\_64.S", "sm4-aesni-avx-asm\_64.S",
-"cast6-avx-x86\_64-asm\_64.S" and others
-\bibitem{b17} https://github.com/torvalds/linux/commit/d34a460092d857f1616e39eed7eac6f40cea2225
-\bibitem{b18} W. O. Alltop, "A Computer Algorithm for Transposing Nonsquare Matrices," in IEEE Transactions on Computers, vol. C-24, no. 10, pp. 1038-1040, Oct. 1975, doi: 10.1109/T-C.1975.224124.
-\bibitem{b19} Jinwoo Suh and V.
-K.
-Prasanna, "An efficient algorithm for out-of-core matrix transposition," in IEEE Transactions on Computers, vol.
-51, no.
-4, pp.
-420-438, April 2002, doi: 10.1109/12.995452.
-\bibitem{b20} Kandemir, M.
-\& Choudhary, A.
-\& Ramanujam, J.
-\& Arunachalam, Meenakshi.
-(2000).
-A unified framework for optimizing locality, parallelism, and communication in out-of-core computations.
-Parallel and Distributed Systems, IEEE Transactions on.
-11.
-648 - 668.
-10.1109/71.877759.
-\bibitem{b21} I-Jui Sung, Juan G´omez-Luna, Josè Marìa Gonzàlez-Linares
-Nicolàs Gui, Wen-Mei W. Hwu. In-Place Transposition of Rectangular Matrices on Accelerators
-\bibitem{b22} Post: An Efficient Matrix Transpose in CUDA C/C++. Feb 18, 2013 By Mark Harris, https://developer.nvidia.com/blog/efficient-matrix-transpose-cuda-cc/
-\bibitem{b23} Hanounik, Bedros. (2000). Diagonal Registers: Novel Vector Register File Design for High Performance and Multimedia Computing. 
+\bibitem{b1} TODO
 \end{thebibliography}

### 2025-01-11 latex/paper2.tex
report: state of the art done

@@
 
+Sheng Ma, Libo Huang, Mingche Lai, Wei Shi. Network-on-chip customizations for message passing interface primitives: From Implementations to Programming Paradigms, 2015
+- In conventional MPI implementations, MPI_Bcast is typically implemented using software with a tree-based algorithm.
+Such implementations exploit point-to-point communication operations.
+Thus, the number of hops to reach leaf nodes increases with the total number of nodes (typically in a logarithmic manner); the latency of MPI_Bcast also increases.
+- this is useful in the findings because we find the same outcome
+- but I use scatter and gather... i should look at the implementation
+
+From the source code: coll/coll_base_scatter:
+- There are 3 implementations of scatter:
+  - binomial tree algorithm, as discussed above uses point-to-point communications
+  - linear functions which work the same for small values
+  - linear non blocking
+- The decision is taken in coll/tuned/coll_tuned_scatter_decision.c
+
+GPU:
+J.
+Gómez-Luna, I.
+-J.
+Sung, L.
+-W.
+Chang, J.
+M.
+González-Linares, N.
+Guil and W.
+-M.
+W.
+Hwu, "In-Place Matrix Transposition on GPUs," in IEEE Transactions on Parallel and Distributed Systems, vol.
+27, no.
+3, pp.
+776-788, 1 March 2016, doi: 10.1109/TPDS.2015.2412549.
+
+Wonderful paper about cache:
+S.
+Chatterjee and S.
+Sen, "Cache-efficient matrix transposition," Proceedings Sixth International Symposium on High-Performance Computer Architecture.
+HPCA-6 (Cat.
+No.PR00550), Touluse, France, 2000, pp.
+195-205, doi: 10.1109/HPCA.2000.824350.
+
 \fi
+
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{blindtext}  % generate random text
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithm}
+% \usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\usepackage{algpseudocode}
+\usepackage{graphicx}
+\usepackage{hyperref}
+
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\begin{document}
+
+\title{Comparing shared and distributed memory performance: A case study on Matrix Transposition and Symmetry Check\\
+%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
+%should not be used}
+%\thanks{Identify applicable funding agency here. If none, delete this.}
+}
+
+\author{\IEEEauthorblockN{Giovanni Santini}
+\IEEEauthorblockA{\textit{University of Trento}\\
+Trento, Italy \\
+giovanni.santini@studenti.unitn.it \\
+MAT. 235441 \\
+\href{https://github.com/San7o/parallel-computing-cpp}{github.com/San7o/parallel-computing-cpp}}
+}
+
+\maketitle
+
+\begin{abstract}
+TODO for later \end{abstract}
+
+\begin{IEEEkeywords}
+parallel computing, matrix transpose, symmetric matrix, benchmark
+\end{IEEEkeywords}
+
+\section{Introduction}
+
+Shared-memory systems are widely used in performance-critial applications
+to leverage multiple CPU cores on the same machine.
+There are, however, limitations to this approach due to all CPU cores
+competing for access to memory over a shared bus, which creates contention
+and reduces performance. Additionally, scaling the number of cores in
+shared-memory systems is challenging. Consequently, distributed-memory
+models are widely used, as they allow memory operations to be
+distributed across multiple computing nodes, and the system can
+be scaled by adding new nodes to the cluster. While distributed-memory
+systems offer significant advantages, they
+depend on the network layer to exchange information, which can
+introduce network latency and bandwidth saturation as potential
+bottlenecks. To mitigate these challenges, technologies such as
+InfiniBand \cite{b0}, RDMA \cite{b-1} and local caching \cite{b-2}
+aim to minimize communication overhead and
+enhance data transfer efficiency. Modern MPI implementations,
+including OpenMPI and MPICH2, are optimized to utilize these
+technologies \cite{b1} \cite{b2}, enabling improved performance in
+distributed-memory environments. The two memory models each
+have distinct strengths and limitations,
+making a comparison between the two essential for evaluating optimal
+system performance. In this report, we present an implementation of
+matTranspose and checkSymm on distributed-memory systems using MPI.
+This implementation is then compared with a shared-memory counterpart
+developed using OpenMP. Additionally, we discuss various compiler
+optimization strategies and explore implicit parallelism techniques,
+such as SIMD optimizations.
+
+The rest of the report is organized as follows.
+In section 2, we discuss the state of the art of the two algorithms.
+In section 3, we elaborate on the algorithms analyzed.
+In section 4, we explain the benchmarking procedures and system
+information. In section 5, we present the results. The conclusion
+is in section 6. Section 7 contains the project's github link
+
+\section{State of the Art}
+
+\iffalse
+- Parallel Matrix Transpose Algorithms on Distributed Memory Concurrent
+  Computers
+- Adaptive Matrix Transpose Algorithms for Distributed Multicore Processors
+- Efficient Distributed-Memory Parallel Matrix-Vector Multiplication with
+  Wide or Tall Unstructured Sparse Matrices.
+- In-Place Matrix Transposition on GPUs
+
+\fi
+
+Distributed-memory matrix transposition is a well discussed problem.
+Choi, Dongarra and Walker \cite{b3} presented an implementation
+based on the block cyclic data distribution. In their approach, the
+matrix is divided into blocks of arbitrary size. If the numbers of rows and
+columns are relatively prime, then the algorithm consists of a two-dimension
+complete exchange of information where each processor sends a sub block
+to every other processor. Binary exchange was shown to be the faster communication
+scheme \cite{b4} which completes the exchange in $log_2 P$ steps where
+$P$ is the number of processors. If the number of rows and columns
+are not relatively prime, processors are grouped into a number of groups
+equal to the greatest common divisor of the number of rows and columns, and
+communication is overlapped between groups using
+non-blocking point-to-point communication. Bowman and Roberts \cite{b5} later
+presented an hybrid approach leveraging both shared and distributed memory.
+First, block of the matrix are transposed locally. Eckstein and Matyasfalvi
+\cite{b6} studied the distributed problem on wide or tall unstructured sparce matrices.
+Their implementation assignes nonzero elements of the matrix evenly
+between processes by partitioning the matrix by sections of non zero
+elements of the same length. Juan Gómez-Luna et. al. \cite{b7} presented
+several GPU applications and an heuristic to detect the optimal approach.
+
+\section{Contribution and Methodology}
+
+TODO
+In this section, implementations of both \textit{matTranspose}
+and \textit{checkSymm} using MPI are presented.
+
+\section{Experiments and System Description}
+
+\iffalse
+- Detailed description of the computing system and platform. \\
+- Relevant specifications or configurations (e.g., libraries
+and programming toolchains). \\
+- Description of the experimental setup, procedures, and
+methodologies used in the project. \\
+- Discussion on how experiments are designed to test the hypotheses
+or achieve the objectives \\
+\fi
+
+TODO
+
+\section{Results and Discussion}
+
+\iffalse
+- Presentation of results \\
+- Analysis and interpretation in context \\
+- Comparison with the state-of-the-art \\
+\fi
+
+TODO
+
+\section{Conclusions}
+
+\iffalse
+- Summary of the key findings and contributions
+\fi
+
+TODO
+
+\section{GIT and Instructions for Reproducibility}
+
+Github link: \href{https://github.com/San7o/parallel-computing-cpp}{https://github.com/San7o/parallel-computing-cpp} \\
+Please read the instructions in the \textit{README.md} for building.
+
+\begin{thebibliography}{00}
+\bibitem{b0} P. Grun, Introduction to InfiniBand for End Users, InfiniBand
+Trade Assoc., Beaverton, OR, USA, 2010. [Online]. Available:
+cw.infinibandta.org/document/dl/7268
+\bibitem{b-1} Kalia, A., Kaminsky, M., and Andersen, D.
+G.
+(2016).
+Design Guidelines for High Performance RDMA Systems.
+2016 USENIX Annual Technical Conference (USENIX ATC 16), 437–450.
+https://www.usenix.org/conference/atc16/technical-sessions/presentation/kalia
+\bibitem{b-2} Message Passing Interface Forum. (2023). MPI: A Message-Passing Interface Standard Version 4.1. Chapter 8.7 "Caching". https://www.mpi-forum.org/docs/mpi-4.1/mpi41-report.pdf 
+\bibitem{b1} G.
+M.
+Shipman, T.
+S.
+Woodall, R.
+L.
+Graham, A.
+B.
+Maccabe and P.
+G.
+Bridges, "Infiniband scalability in Open MPI," Proceedings 20th IEEE International Parallel and Distributed Processing Symposium, Rhodes, Greece, 2006, pp.
+10 pp.-, doi: 10.1109/IPDPS.2006.1639335.
+\bibitem{b2} J.
+Liu et al., "Design and implementation of MPICH2 over InfiniBand with RDMA support," 18th International Parallel and Distributed Processing Symposium, 2004.
+Proceedings., Santa Fe, NM, USA, 2004, pp.
+16-, doi: 10.1109/IPDPS.2004.1302922.
+\bibitem{b3} Jaeyoung Choi, Jack J.
+Dongarra, David W.
+Walker.
+Parallel Matrix Transpose Algorithms on Distributed Memory Concurrent Computers.
+Mathematical Sciences Section Oak Ridge National Laboratory P.O.
+Box 2008, Bldg.
+6012 Oak Ridge, TN 37831-6367
+\bibitem{b4} S.
+Takkella and S.
+Seidel, "Complete exchange and broadcast algorithms for meshes," Proceedings of IEEE Scalable High Performance Computing Conference, Knoxville, TN, USA, 1994, pp.
+422-428, doi: 10.1109/SHPCC.1994.296674.
+\bibitem{b5} John C. Bowman and Malcolm Roberts. Adaptive Matrix Transpose Algorithms for Distributed Multicore Processors
+\bibitem{b6} Eckstein, J., Matyasfalvi, G. (2018). Efficient Distributed-Memory Parallel Matrix-Vector Multiplication with Wide or Tall Unstructured Sparse Matrices. https://arxiv.org/abs/1812.00904 
+\bibitem{b7} J.
+Gómez-Luna, I.
+-J.
+Sung, L.
+-W.
+Chang, J.
+M.
+González-Linares, N.
+Guil and W.
+-M.
+W.
+Hwu, "In-Place Matrix Transposition on GPUs," in IEEE Transactions on Parallel and Distributed Systems, vol.
+27, no.
+3, pp.
+776-788, 1 March 2016, doi: 10.1109/TPDS.2015.2412549.
+\end{thebibliography}
+\vspace{12pt}
+
+\end{document}

### 2025-01-12 latex/paper2.tex
second-assignment: contribution and methodology

@@
 presented an hybrid approach leveraging both shared and distributed memory.
-First, block of the matrix are transposed locally. Eckstein and Matyasfalvi
+Eckstein and Matyasfalvi
 \cite{b6} studied the distributed problem on wide or tall unstructured sparce matrices.
@@
 
-TODO
 In this section, implementations of both \textit{matTranspose}
-and \textit{checkSymm} using MPI are presented.
+and \textit{checkSymm} using MPI are presented, starting with
+\textit{matTranspose}. We benchmark the bandwidth and speedup
+of different MPI implementations compared to a serial and to a OpenMPI one.
+Different compiler optimizations will also be compared. All benchmarks
+will be tested against every matrix size from $2^2$ to $2^{12}$,
+we will assume the number of processes is divisible by the matrix size
+and the square root is an integer number for ease of implementation,
+however the algorithms can be generalised with few modifications.
+The algorithms for \textit{matTranspose} will now be explained.
+
+\textit{matTranspose} takes a $N \times N$ input square matrix
+\textit{input}, a matrix of the same shape \textit{ouput}, and the
+size of the two matrices \textit{size}. The analyzed algorithms are the following:
+\begin{itemize}
+\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix and storing them in the output matrix with indexes of rows and columns swapped.
+\item \textit{matTransposeMPI}: this implementation takes full advantage
+  of MPI datatypes to test bandwidth. MPI has the possibility to define a
+  custom datatype to access memory, we used \textit{MPI\_Type\_contiguous}
+  to define a row type and \textit{MPI\_Type\_vector} + \textit{MPI\_Type\_create\_resized}
+  for a column type. The root process scatters rows with \textit{MPI\_Scatter} and
+  retrieves columns with \textit{MPI\_Gather}, therefore
+  transposing the matrix. Each process receives a block of rows and
+  sends back the same block to the root but as column type. The algorithm
+  automatically assign the correct number of rows/column to each process based on the
+  number of processes so that the entire matrix is transposed.
+  For example, in a $16x16$ matrix with $4$ processes, each processes
+  receives $\frac{16}{4}=4$ rows and sends the same number back as
+  columns.
+  This implementation fully tests the communication
+  bandwidth and latency by relying only on communication and MPI
+  functionalities.
+\item \textit{matTransposeMPIBlock}: this implementation divides the
+  original matrix into a mesh of square sub-matrixes with side length equal to
+  $\frac{N}{\sqrt{\#processes}}$ using MPI datatypes. Each sub-matrix is assigned
+  to a different processes. The root process sends the sub-matrix to the
+  other processes with \textit{MPI\_Scatterv} and receives the same
+  number of sub-matrices but in transposed order with \textit{MPI\_Gatherv},
+  for example process[0][1] will sand to root[1][0].
+  Each process tranposes its sub-matrix and sends It back to the root.
+  This approach tests a different communication method.
+\end{itemize}
+The algorithm for \textit{checkSymm} will now be explained. The algorithm
+takes an $N \times N$ square matrix \textit{input}, a size \textit{size}
+and returns true if the matrix is symmetrix, false otherwise.
+The algorithm the following:
+\begin{itemize}
+\item \textit{checkSymm}: naive implementation of checkSymm. It Iterates over the upper half of the matrix and testing if every value is equal to the symmetric value.
+\item \textit{checkSymmMPI}: similar to \textit{matTransposeMPIBlock},
+but each process receives both a sub-matrix and the transposed sub-matrix,
+that is, if we treated each sub-matrix as a single element of a matrix,
+the element in the tranposed position will be received. Each process compares the two
+sub-matrices to check if they are symmetric. The result is computed
+via \textit{MPI\_Reduce}.
+\end{itemize}
 
@@
 
-TODO
+Experiments were conducted on a cluster with 96 Intel(R) Xeon(R) Gold
+6252N CPUs (2.30GHz) featuring 32K L1d and L1i caches, 1022K L2 cache,
+36608K L3 cache (64-bit cache lines), 1 thread per core, and 4GB RAM,
+running Linux 3.10. The algorithms were benchmarked using \textit{valFuzz},
+with benchmarks compiled using gcc-9.1.0. Four benchmark sets were
+compiled with different optimization levels: none, -O1, -O2 and "-O3
+-Ofast -march=native".
+
+To ensure statistical significance, each benchmark was executed 100
+times, recording wall-clock execution time statistics: minimum, maximum,
+median, mean, standard deviation, Q1, and Q3. Cache interference was
+minimized by clearing it before each run by writing to an array twice
+the L3 cache size. Input matrices, consistent across algorithms, were
+initialized with uniformly random floats (-1 to 1) generated via a
+Linear Congruential Generator (LCG) \cite{b8} with parameters a=1664525,
+c=1013904223, m=232 \cite{b9}, and seed 1337.
+
+To expedite random number generation, values were precomputed at
+compile time using \textit{tenno-tl}. Additional \textit{tenno-tl}
+features, such as ranges and arrays, were also employed.
 
@@
 776-788, 1 March 2016, doi: 10.1109/TPDS.2015.2412549.
+\bibitem{b8} Rotenberg, A. (1960). "A New Pseudo-Random Number Generator". Journal of the ACM. 7 (1): 75–77
+\bibitem{b9} Numerical Recipes ranqd1, Chapter 7.1, An Even Quicker Generator, Eq. 7.1.6
+parameters from Knuth and H. W. Lewis
 \end{thebibliography}

### 2025-01-19 latex/paper2.tex
second-assignment: report completed

@@
 \usepackage{hyperref}
+\usepackage{booktabs} % table
 
@@
 
-\title{Comparing shared and distributed memory performance: A case study on Matrix Transposition and Symmetry Check\\
+\title{Comparing shared and distributed memory algorithms: A case study on Matrix Transposition and Symmetry Check\\
 %{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 \begin{abstract}
-TODO for later \end{abstract}
+This study evaluates matrix transposition on shared- and distributed-memory
+architectures, focusing on scalability. The shared-memory implementation
+achieves optimal linear strong scaling for matrixes of size $N<2^{10}$,
+while distributed-memory implementations are hindered by communication overhead, with
+\textit{matTransposeMPIBlock} performing better then the seqencial
+implementation only on big matrixes, but scaling well on big matrices.
+Weak scaling is poor for all implementations, though the shared-memory
+approach is slightly more efficient. Shared-memory systems are ideal
+for moderate problem sizes, while distributed-memory systems trade
+efficiency for scalability with larger matrices.
+\end{abstract}
 
@@
 making a comparison between the two essential for evaluating optimal
-system performance. In this report, we present an implementation of
-matTranspose and checkSymm on distributed-memory systems using MPI.
+system performance. In this report, an implementation of \textit{matTranspose}
+and \textit{checkSymm} is presented on distributed-memory systems using MPI.
 This implementation is then compared with a shared-memory counterpart
-developed using OpenMP. Additionally, we discuss various compiler
-optimization strategies and explore implicit parallelism techniques,
-such as SIMD optimizations.
+developed using OpenMP. Additionally, various compiler
+optimization strategies are discussed.
 
@@
 
+Since both \textit{matTranspose} and \textit{checkSymm} are memory bounded
+\cite{b10}, most of the time is expected to be spent on memory movement rather then
+math operation. Implementations with MPI distribute the matrix over the network,
+potentially creating a bottleneck by adding additional latency due to
+synchronyzation, the overhead of the operative system's network stack and
+network congestion. Practically the number of computing nodes
+in a cluster could greately surpass the number of CPU cores of a
+shared-memory computer therefore overcoming the disadvantages and
+possibly increasing the performance by dividing the work in a greater
+number of workers, however the advantages could be shadowed by the
+memory movement memory bounded algorithms. On the other hand, shared-memory
+algorithms do not need to send memory since they share the same address
+space and are limited by the bus bandwidth and scheduling policies.
 In this section, implementations of both \textit{matTranspose}
@@
 \textit{matTranspose}. We benchmark the bandwidth and speedup
-of different MPI implementations compared to a serial and to a OpenMPI one.
-Different compiler optimizations will also be compared. All benchmarks
-will be tested against every matrix size from $2^2$ to $2^{12}$,
-we will assume the number of processes is divisible by the matrix size
-and the square root is an integer number for ease of implementation,
-however the algorithms can be generalised with few modifications.
-The algorithms for \textit{matTranspose} will now be explained.
+of different MPI implementations compared to a serial and to a OpenMPI one,
+discussing weak and strong scaling. Different compiler optimizations
+will also be compared. All benchmarks will be tested against every
+matrix size from $2^2$ to $2^{12}$, we will assume the number of processes
+is divisible by the matrix size and the square root is an integer
+number for ease of implementation, however the algorithms can be
+generalised with few modifications. The algorithms for \textit{matTranspose}
+will now be explained.
 
@@
 \begin{itemize}
-\item \textit{matTranspose}: naive implementation of matTranspose, iterating over values of the input matrix and storing them in the output matrix with indexes of rows and columns swapped.
+\item \textit{matTranspose}: naive implementation of matTranspose,
+iterating over values of the input matrix and storing them in the
+output matrix with indexes of rows and columns swapped.
 \item \textit{matTransposeMPI}: this implementation takes full advantage
-  of MPI datatypes to test bandwidth. MPI has the possibility to define a
-  custom datatype to access memory, we used \textit{MPI\_Type\_contiguous}
+  of MPI datatypes to test bandwidth. Leveraging 
+  custom datatypes in MPI, \textit{MPI\_Type\_contiguous} was used
   to define a row type and \textit{MPI\_Type\_vector} + \textit{MPI\_Type\_create\_resized}
@@
   This approach tests a different communication method.
+\item \textit{matTransposeOMPX}: this set of algorithms are all derived
+  from \textit{matTranspose} with the outer for loops parallelized using
+  omp with the number of threads \textit{X} being a power of 2 from $2^1$ to $2^6$.
 \end{itemize}
@@
 \begin{itemize}
-\item \textit{checkSymm}: naive implementation of checkSymm. It Iterates over the upper half of the matrix and testing if every value is equal to the symmetric value.
+\item \textit{checkSymm}: naive implementation of checkSymm. It
+Iterates over the upper half of the matrix and testing if every value
+is equal to the symmetric value.
 \item \textit{checkSymmMPI}: similar to \textit{matTransposeMPIBlock},
@@
 that is, if we treated each sub-matrix as a single element of a matrix,
-the element in the tranposed position will be received. Each process compares the two
-sub-matrices to check if they are symmetric. The result is computed
+the element in the tranposed position will be received. Each process
+compares the two sub-matrices to check if they are symmetric. The
+result is computed
 via \textit{MPI\_Reduce}.
+\item \textit{checkSymOMPX}: similar to \textit{matTransposeOMPX} but
+based on \textit{checkSymm}. After each parallel section, a barrier
+is needed to synchronize the shared boolean output variable.
 \end{itemize}
+
+\textit{Strong Scaling} or \textit{Speedup} is defined as the reatio
+between the time for running the algorithm using one processor $t(1)$ over
+the time for running the same algorithm on $N$ processors $t(N)$ \ref{eq:1}.
+\begin{equation} \label{eq:1}
+Speedup = \frac{t(1)}{t(N)}
+\end{equation}
+In an ideal scenario, software would achieve a linear increase in
+speed proportional to the number of processors used (speedup = N),
+signifying that each processor is fully utilizing its computational
+capacity.
+In case of \textit{Weak Scaling}, both the number of processors $p$
+and the problem size $N$ are increased \ref{eq:2}. It evaluates wheter a program
+can handle larger problems efficiently when more resources are available.
+\begin{equation} \label{eq:2}
+S_w = \frac{t(N)}{T_p(N*p)}
+\end{equation}
 
@@
 -Ofast -march=native".
-
-To ensure statistical significance, each benchmark was executed 100
-times, recording wall-clock execution time statistics: minimum, maximum,
-median, mean, standard deviation, Q1, and Q3. Cache interference was
+MPI algorithms were benchmarked with number of processes equal to 4, 16, 64.
+MPI benchmarks are composed of a master (root) process and a fleet
+of worker processes. Before each benchmark, the master sends
+a synchronization message to the workers telling them which algorithm
+to execute and how many times, It will then start executing the agreed
+algorithm and partecipate in the work. Only the master process will be
+timed, recording wall-clock execution time statistics: minimum, maximum,
+median, mean, standard deviation, Q1, and Q3. To ensure statistical
+significance, each benchmark was executed 10 times. Cache interference was
 minimized by clearing it before each run by writing to an array twice
@@
 c=1013904223, m=232 \cite{b9}, and seed 1337.
-
 To expedite random number generation, values were precomputed at
@@
 
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_base.pdf"}
+	\caption{matTransposeMPI average time over matrix size}
+	\label{fig:transpose_mpi_base}
+\end{figure}
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_base.pdf"}
+	\caption{matTransposeMPIBlock average time over matrix size}
+	\label{fig:transpose_mpi_block}
+\end{figure}
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_speedup.pdf"}
+	\caption{matTransposeMPI speedup measurements}
+	\label{fig:transpose_mpi_speedup}
+\end{figure}
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_speedup.pdf"}
+	\caption{matTransposeMPIBlock speedup measurements}
+	\label{fig:transpose_mpi_block_speedup}
+\end{figure}
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/transpose_explicit_speedup.pdf"}
+	\caption{matTransposeOMP speedup}
+	\label{fig:transpose_mp_speedup}
+\end{figure}
+
 \iffalse
-- Presentation of results \\
-- Analysis and interpretation in context \\
-- Comparison with the state-of-the-art \\
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_weak_scaling.pdf"}
+	\caption{matTransposeMPI weak scaling measurements}
+	\label{fig:transpose_mpi_weak_scaling}
+\end{figure}
 \fi
 
-TODO
+\iffalse
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_bandwidth.pdf"}
+	\caption{matTransposeMPIBlock bandwidth measurements}
+	\label{fig:transpose_mpi_block_bandwidth}
+\end{figure}
+\fi
+
+\iffalse
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_weak_scaling.pdf"}
+	\caption{matTransposeMPIBlock weak scaling measurements}
+	\label{fig:transpose_mpi_block_weak_scaling}
+\end{figure}
+\fi
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_weak_scaling_final.pdf"}
+	\caption{Comparison of weak scaling between all the matTranspose implementations, averaging all the compiler optimization flags}
+	\label{fig:transpose_mpi_block_weak_scaling_final}
+\end{figure}
+
+\begin{table}[h!]
+\centering
+\caption{Average time \% spent on each part of \\
+\textit{matTransposeMPIBlock}}
+\label{tab:normalized_averages}
+\begin{tabular}{lc}
+\toprule
+\textbf{Column} & \textbf{Average Time \%} \\
+\midrule
+Displacement    & 0.000088 \\
+Setup           & 0.000246 \\
+Scatter         & 0.372926 \\
+Transpose       & 0.149277 \\
+Gather          & 0.477462 \\
+\midrule
+\end{tabular}
+\end{table}
+
+The results of the study are presented here. Shared-memory performance
+data referenced in this analysis are derived from previous work \cite{b10}.
+
+As discussed earlier, the memory-bound nature of both \textit{matTranspose}
+and \textit{checkSymm} inherently disadvantages the distributed-memory
+implementation when compared to the shared-memory approach due to its
+higher memory movement requirements. Experimental results confirm this
+observation. In the case of \textit{matTransposeMPIBlock}, a significant
+portion of the execution time is consumed by data transmission
+(approximately 37\%) and data reception (approximately 47\%), as shown
+in Table~\ref{tab:normalized_averages}. This behavior accounts for the
+slightly slower performance of \textit{matTransposeMPI} relative to the
+sequential implementation (Figure~\ref{fig:transpose_mpi_base}) and the
+considerably slower performance compared to the shared-memory OpenMP
+implementation for $N > 2^7$. However, as the matrix size increases,
+the performance of \textit{matTransposeMPI} approaches that of the
+sequential version. Optimization of data transmission patterns further
+improves performance. Notably, \textit{matTransposeMPIBlock} outperforms
+\textit{matTransposeMPI} and even surpasses the sequential implementation
+(Figure~\ref{fig:transpose_mpi_block}), although it remains inferior
+to the shared-memory approach. The superiority of shared-memory
+implementations is further underscored by the use of a loopback interface
+for MPI communication within the same node: if data packets were
+transmitted over slower mediums, such as Ethernet, the performance gap
+between shared-memory and distributed-memory implementations would
+likely widen, leading to even slower execution for the latter.
+\textit{MatTransposeMPIBlock} exhibits better strong scaling than
+\textit{matTransposeMPI} (Figure~\ref{fig:transpose_mpi_block_speedup}
+and Figure~\ref{fig:transpose_mpi_block_speedup}) due to its optimized data
+transfer patterns. In contrast, \textit{matTransposeOMP}
+demonstrates linear scaling (Figure~\ref{fig:transpose_mp_speedup}), achieving
+optimal strong scaling performance until the scaling becomes sublinear
+for large matrix sizes ($N>2^{9}$), while \textit{matTransposeMPIBlock}
+appears to not slow down scaling. This result underscores the
+advantage of shared-memory systems for matrix transposition on small matrixes,
+where the high memory bandwidth and low-latency data movement within
+a single node allow efficient scaling with additional processors. The
+overall weak scaling results indicate that all implementations
+struggle to effectively manage the increased workload as system size
+expands (Figure~\ref{fig:transpose_mpi_block_weak_scaling_final}).
+
+Equivalent results are found in implementations of \textit{checkSymm}.
 
@@
 \fi 
-
-TODO
-
-\section{GIT and Instructions for Reproducibility}
-
-Github link: \href{https://github.com/San7o/parallel-computing-cpp}{https://github.com/San7o/parallel-computing-cpp} \\
-Please read the instructions in the \textit{README.md} for building.
+This study compared the performance of matrix transposition on
+shared- and distributed-memory systems. The shared-memory implementation
+achieved optimal linear strong scaling up to matrix of size $N<2^{10}$, while distributed-memory
+implementations were hindered by communication overhead, with
+\textit{matTransposeMPIBlock} showing some improvement. Weak scaling
+was poor for all implementations, though shared-memory performed
+slightly better. Overall, shared-memory systems are more efficient
+for moderate problem sizes, while distributed-memory systems are
+necessary for larger matrices despite reduced efficiency.
 
@@
 parameters from Knuth and H. W. Lewis
+\bibitem{b10} G. Santini. Matrix Transposition and checkSymm: A case study on superscalar architectures.
 \end{thebibliography}

### 2025-01-19 latex/paper2.tex
latex: updated names in figures

@@
 	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_base.pdf"}
-	\caption{matTransposeMPI average time over matrix size}
+	\caption{matTransposeMPI average time over matrix size of multiple number of processes}
 	\label{fig:transpose_mpi_base}
@@
 	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_base.pdf"}
-	\caption{matTransposeMPIBlock average time over matrix size}
+	\caption{matTransposeMPIBlock average time over matrix size of multiple number of processes}
 	\label{fig:transpose_mpi_block}
@@
 	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_speedup.pdf"}
-	\caption{matTransposeMPI speedup measurements}
+	\caption{matTransposeMPI speedup measurements of multiple number of processes}
 	\label{fig:transpose_mpi_speedup}
@@
 	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_speedup.pdf"}
-	\caption{matTransposeMPIBlock speedup measurements}
+	\caption{matTransposeMPIBlock speedup measurements of multiple number of processes}
 	\label{fig:transpose_mpi_block_speedup}
@@
 	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_weak_scaling.pdf"}
-	\caption{matTransposeMPI weak scaling measurements}
+	\caption{matTransposeMPI weak scaling measurements of multiple number of processes}
 	\label{fig:transpose_mpi_weak_scaling}
@@
 	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_bandwidth.pdf"}
-	\caption{matTransposeMPIBlock bandwidth measurements}
+	\caption{matTransposeMPIBlock bandwidth measurements of multiple number of processes}
 	\label{fig:transpose_mpi_block_bandwidth}
@@
 	\includegraphics[width=0.44\textwidth]{"../benchmarks/plotting/images/matTransposeMPI_block_weak_scaling.pdf"}
-	\caption{matTransposeMPIBlock weak scaling measurements}
+	\caption{matTransposeMPIBlock weak scaling measurements of multiple number of processes}
 	\label{fig:transpose_mpi_block_weak_scaling}
